---
title: "Vita9b and Vita9c: is the genotype effect sex-dependent?"
subtitle: "ITP genotyped cohorts, control + NDE-treated"
author: "Joshua Preston"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: default
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(nnet)
library(qtl)

# echo is on here: the point of this document is to show the code alongside
# the results, not just the results
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = FALSE)
```

trajectory.R tests each locus with an LRT, which shows whether genotype predicts
BW class but not in which direction. Vita9b and Vita9c are significant within
each sex but not when the sexes are pooled. This analysis tests directly if 
this can be attributed to opposing directional effects cancelling on pooling.

# Load genotypes, phenotypes, metadata, and class assignments

```{r source trajectory}
# Define the two loci in question
loci <- c(Vita9b = "9_104091597", Vita9c = "9_124056586")

# trajectory.R (Danny) builds mcross / gtsp, defines computeLog10, and writes
# output/census_mapping.txt -- the input to Figures 5L and S9D. Sourcing it
# regenerates that file, so the values should be unchanged afterward.
source("trajectory.R")
```

```{r pheno}
# Covariates -------------------------------------------------------------------
# mcross carries the phenotype columns from um-het3-rqtl.csvr alongside the
# genotypes, keyed by GenoID. That key is what lets census rownames index into
# gtsp further down.
pheno <- data.frame(
  sex = as.factor(pull.pheno(mcross)[, "sex"]),
  site = as.factor(pull.pheno(mcross)[, "site"]),
  cohort = as.factor(pull.pheno(mcross)[, "cohort"]),
  treatment = as.factor(pull.pheno(mcross)[, "treatment"])
)
rownames(pheno) <- pull.pheno(mcross)[, "GenoID"]


# Census loader ----------------------------------------------------------------
# trajectory.R indexes the id column by position (15 / 11 / 12 / 17 / 13 / 13,
# depending on the file). Naming it is equivalent and does not depend on column
# order holding.
loadCensus <- function(file) {
  census <- read.csv(file, row.names = "id")
  cdata <- cbind(
    new_class_bw = as.factor(census[, "new_class_bw"]),
    pheno[rownames(census), ]
  )
  list(cdata = cdata, gts = gtsp[rownames(census), ])
}
```

# Reported associations

Reproduce the pattern being explained: significant in each sex, not when pooled.

```{r lod table}
lodM[names(loci), ]
```

Both loci clear the Bonferroni threshold of 3.85 (0.05 / 354 tests) in the
sex-stratified control+treated analyses (`census_tx_f`, `census_tx_m`) and
neither clears it when the sexes are pooled (`census_tx`). The control-only
columns are all well below threshold, which is why Figures 5L and S9D display
the `census_tx_*` columns.

# Class structure of the three models

The pooled model and the two sex-stratified models are separate LCM fits, so
their classes are not the same variable.

```{r class counts}
tx <- loadCensus("output/itp_geno_tx_census.csv")
txf <- loadCensus("output/itp_geno_tx_f_census.csv")
txm <- loadCensus("output/itp_geno_tx_m_census.csv")

# pooled model: three classes, two of which are strongly sex-segregated.
# sex comes from the cross object, where R/qtl codes 0 = female, 1 = male --
# relabelled here so the table matches the M/F used in the census files.
cls <- table(sex = factor(tx$cdata$sex, labels = c("F", "M")),
             class = tx$cdata$new_class_bw)
cls

# share of each class that is female / male
round(100 * prop.table(cls, 2), 1)

# each sex-stratified model: two classes, fit within that sex
table(txf$cdata$new_class_bw)
table(txm$cdata$new_class_bw)
```

Class 2 is 94% M, class 3 is 99% F... so in the pooled model the class 2 vs 1
contrast is driven almost entirely by males and class 3 vs 1 almost entirely by
females. Each stratified model instead fits two classes within that sex. The
three analyses are not the same test run on different subsets of animals.

# Does the genotype effect depend on sex?

Fit on the pooled census so both sexes share one class definition. The first
model (`lm.add`) gives genotype a single effect applied to every mouse; the
second (`lm.int`) adds a sex-by-genotype interaction, letting the effect differ
between M/F. Opposing effects in the two sexes can only exist in the second
model, so if that explanation is right, it must fit better.

```{r interaction}
cdata <- tx$cdata

lods.int <- c()
for (marker in loci) {
  mp <- tx$gts[, grep(marker, colnames(tx$gts))]
  lm.add <- multinom(new_class_bw ~ sex + site + cohort + treatment + mp + 0,
                     data = cdata, trace = FALSE)
  lm.int <- multinom(new_class_bw ~ site + cohort + treatment + sex * mp + 0,
                     data = cdata, trace = FALSE)
  lods.int <- c(lods.int, computeLog10(lm.add, lm.int))
}
names(lods.int) <- names(loci)

round(lods.int, 2)
10^-lods.int
```

Neither interaction comes close to significance, so letting the genotype effect
differ between the sexes does not improve the fit. The opposing-effects
explanation requires this test to be positive, and it is not.

To be fair, this test has limited power: class 2 is 94% M and class 3 is 99% F,
so the sex-specific effects are estimated from few animals on the contrasts that
matter. Thus, a null interaction is weak evidence of absence, which means that
we need to examine the direction of the effects before making conclusions.

# Direction of the effect in each sex

Same models trajectory.R fits, reading out the genotype coefficients instead of
a likelihood ratio.

```{r direction}
# Helper function --------------------------------------------------------------

genoCoef <- function(x, marker) {
  mp <- x$gts[, grep(marker, colnames(x$gts))]
  cdata <- x$cdata
  fit <- multinom(new_class_bw ~ site + cohort + treatment + mp + 0,
                  data = cdata, trace = FALSE)
  co <- coef(fit)
  co[grep("^mp", names(co))]
}


# Coefficients by sex ----------------------------------------------------------

out <- c()
for (locus in names(loci)) {
  f <- genoCoef(txf, loci[locus])
  m <- genoCoef(txm, loci[locus])
  out <- rbind(out, data.frame(
    locus = locus,
    allele = sub("^mp", "", names(f)),
    female = round(f, 3),
    male = round(m, 3)
  ))
}
rownames(out) <- NULL

out
```

The four haplotype probabilities sum to 1, so adding a constant to all four
coefficients gives the same fit. Only their relative ordering is interpretable,
which is what this correlation compares:

```{r direction summary}
for (locus in names(loci)) {
  s <- out[out$locus == locus, ]
  cat(locus, ": r =", round(cor(s$female, s$male), 2),
      " highest =", s$allele[which.max(s$female)], "(F)",
      s$allele[which.max(s$male)], "(M)",
      " lowest =", s$allele[which.min(s$female)], "(F)",
      s$allele[which.min(s$male)], "(M)\n")
}
```

# Summary

```{r summary table}
data.frame(
  locus = names(loci),
  lod_female = lodM[names(loci), "census_tx_f"],
  lod_male = lodM[names(loci), "census_tx_m"],
  lod_pooled = lodM[names(loci), "census_tx"],
  interaction_p = round(10^-lods.int, 2),
  cor_female_male = sapply(names(loci), function(locus) {
    s <- out[out$locus == locus, ]
    round(cor(s$female, s$male), 2)
  }),
  row.names = NULL
)
```

In summary:

1. Each locus is significant within each sex and not when the sexes are pooled.
2. The sex-by-genotype interaction is not significant.
3. The F/M haplotype effects are positively correlated rather than opposed.

Together, this all fails to support the opposing-effects explanation. Therefore,
the pooled result is not a "cancelled" version of the sex-stratified results.
Moreover, since the pooled model generates three classes, two of which are
strongly sex-segregated, but each stratified model has two classes fit within
M/F, the two analyses in reality test genotype associations against
fundamentally different outcomes, thus their p-values were not directly
comparable in the first place.


