---
title: "SLAM Metabolic Class Supplmental Figures - `r config$out_tag` Mice"
subtitle: "External Validation: `r config$external_validate`"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../06_create_figures/output/slam_c1-c10_age_all_bwfatgluc.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)

window_legend <- TRUE
window_legend_palette <- magma(3, begin = .2, end = .7, direction = -1)
window_legend_labels <- c("2 measures", "3 measures", "4 measures")

cumulative_legend <- TRUE
cumulative_legend_palette <- viridis(3, begin = .2, end = .7, direction = -1)

max_age <- 150
max_age_km <- NA
```

```{r read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  # params$input_path <- "../06_create_figures/output/test_local.yaml"
  # params$input_path <- "../06_create_figures/output/all_only_bw.yaml"
  # params$input_path <- "../06_create_figures/output/20250306_itp.yaml"
  # params$input_path <- "../06_create_figures/output/all.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_itp_C2010_C2011_C2013_C2016_controls_cohort.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_all_cohort.yaml"
  params$input_path <- "../06_create_figures/output/slam_c1-c10_age_fb6_bwfatgluc.yaml"

  # params$input_path <- "../06_create_figures/output/all_lean.yaml"
  # params$input_path <- "../06_create_figures/output/fb6.yaml"
  # params$input_path <- "../06_create_figures/output/external_validation.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
```

```{r load models}
load(input$datasets_path)
load(input$final_models_path)
config <- yaml::read_yaml(input$config_path)


# load in main_cat_surv
main_cat_surv <- read.csv(file = config$survival_dataset$path)
main_cat_surv <- as.data.frame(main_cat_surv)

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

```{r how they work, fig.height=5, fig.width=10}
p1 <- final_models[, "spag_plot"][[1]]
p3 <- final_models[, "traj_obs"][[1]]

how_it_works <- cowplot::plot_grid(p1, p3, nrow = 1)
```

```{r validation data}
plot_validation <- function(data1,
                            data2,
                            age_var,
                            outcome,
                            age_name,
                            outcome_name) {
  ggplot2::ggplot(
    data = data1,
    mapping = ggplot2::aes(
      x = eval(as.symbol((age_var))),
      y = eval(as.symbol(outcome))
    )
  ) +
    ggplot2::geom_point(alpha = .03, color = "navy") +
    ggplot2::geom_point(
      data = data2,
      mapping = ggplot2::aes(
        x = eval(as.symbol((age_var))),
        y = eval(as.symbol(outcome))
      ),
      alpha = .5,
      color = "navy"
    ) +
    ggplot2::labs(
      x = age_name,
      y = outcome_name
    )
}


datasets <- lapply(datasets, function(dataset) {
  data1 <- dataset$data
  outcome <- dataset$outcome
  age_var <- dataset$age_var[[1]]
  age_var <- paste0(age_var, "_ns")
  age_var_name <- paste(
    dataset$labels$age_var_name,
    dataset$labels$age_var_units
  )
  outcome_name <- paste(
    dataset$labels$oc_name,
    dataset$labels$oc_units
  )

  prediction_data <- dataset$prediction_data$data

  dataset$validation_plots <- lapply(prediction_data, function(data) {
    lapply(data, plot_validation,
      data1 = data1,
      age = age_var, outcome = outcome,
      age_name = age_var_name,
      outcome_name = outcome_name
    )
  })

  dataset
})
```

```{r hist outcome, echo = FALSE, include = FALSE}
hist_outcome_plots <- lapply(datasets, function(dataset) {
  data <- dataset$data
  outcome <- dataset$outcome
  data_mod <- dataset$data_mod
  ggplot2::ggplot(data = data, aes_string(x = outcome)) +
    ggplot2::geom_histogram() +
    ggplot2::labs(title = data_mod)
})

hist_outcome_plot_panel <- cowplot::plot_grid(
  plotlist = hist_outcome_plots,
  nrow = round(sqrt(length(datasets))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of outcomes for complete and training data",
    sep = " "
  ))

hist_outcome_plot_panel <- cowplot::plot_grid(title_theme,
  hist_outcome_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

```{r hist log outcome, echo = FALSE, include = FALSE}
hist_log_outcome_plots <- lapply(datasets, function(dataset) {
  data <- dataset$data
  outcome <- dataset$outcome
  outcome <- paste0("log(", outcome, ")")
  data_mod <- dataset$data_mod
  ggplot2::ggplot(data = data, aes_string(x = outcome)) +
    ggplot2::geom_histogram() +
    ggplot2::labs(title = data_mod)
})

hist_log_outcome_plot_panel <- cowplot::plot_grid(
  plotlist = hist_log_outcome_plots,
  nrow = round(sqrt(length(datasets))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of log(outcomes) for complete and training data",
    sep = " "
  ))

hist_log_outcome_plot_panel <- cowplot::plot_grid(title_theme,
  hist_log_outcome_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

```{r hist prob, echo = FALSE, include = FALSE}
rest_index <- final_models$data_mod == "og"
census <- final_models[rest_index, "combined_census"][[1]]

prob_cols <- names(census)[grepl("^prob\\d+$", names(census))]


hist_plots <- lapply(prob_cols, function(prob_col) {
  ggplot2::ggplot(data = census, aes_string(x = prob_col)) +
    ggplot2::geom_histogram()
})

hist_plot_panel <- cowplot::plot_grid(
  plotlist = hist_plots,
  nrow = round(sqrt(length(prob_cols))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of probabilites of class membership for all subjects",
    sep = " "
  ))

hist_plot_panel <- cowplot::plot_grid(title_theme, hist_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

```{r initiate figure and table number}
fig_no <- 1
tab_no <- 1
figs <- list()
```

### Figure `r as.character(fig_no)`: How it works

```{r how it works, fig.width = 12, fig.height = 6}
figs$how_it_works <- how_it_works
how_it_works
fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Kaplan Meier by Harmonizer

This may explain our trouble externally validating with cohorts 9 and 10, since
both have lower levels of survival

```{r kap plot by harmonize}
figs$harmonize_km <- lapply(
  config$datasets[[1]]$harmonize$variable,
  function(var) {
    fig <- SLAM::surv_plot_km(census, var,
      age_death = config$survival_dataset$age_death,
      event = config$survival_dataset$event
    )
    table <- kable(fig$hr$hr_table[, "final", drop = FALSE])
    plot <- fig$plot
    list(plot, table)
  }
)
figs$harmonize_km

fig_no <- fig_no + 1
```

\newpage

```{r heights and widths}
pred_obs_height <- nrow(final_models) / 2 * 6
dose_height_width <- sqrt(length(unique(final_models$census[[1]]$class))) * 6
```

### Figure `r as.character(fig_no)`: KM plots of covariates - complete data

```{r plot km covs}
census <- final_models[rest_index, "combined_census"][[1]]

cov_dummy <- datasets[[1]]$covariates_dummy
figs$cov_km <- lapply(cov_dummy, function(cov) {
  census[[cov]] <- factor(census[[cov]])
  fig <- SLAM::surv_plot_km(census, cov,
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )
  table <- kable(fig$hr$hr_table[, "final", drop = FALSE])
  plot <- fig$plot
  list(plot, table)
})
figs$cov_km

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Histogram of Outcome Variables - Complete and Training Data

```{r disply hist outcome panel, fig.height=12, fig.width=12}
figs$hist_outcome_plot_panel <- hist_outcome_plot_panel
figs$hist_outcome_plot_panel

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Histogram of log transform of outcome Variables - Complete and Training Data

```{r disply hist log outcome panel, fig.height=12, fig.width=12}
figs$hist_log_outcome_plot_panel <- hist_log_outcome_plot_panel
figs$hist_log_outcome_plot_panel

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Histogram of probability of class membership - Complete Data

```{r disply hist class prob panel, fig.height=12, fig.width=12}
figs$hist_plot_panel <- hist_plot_panel
figs$hist_plot_panel

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Display Testing Data - Cumulative

```{r cumulative validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[1]]
figs$cumulative_validation <- plot_grid(plotlist = plots)
figs$cumulative_validation
```

\newpage

### Figure `r as.character(fig_no)`: Display Testing Data - Window

```{r window validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[2]]
figs$window_validation <- plot_grid(plotlist = plots)
figs$window_validation
```

\newpage

### Figure `r as.character(fig_no)`: Display Testing Data - Interval

```{r interval validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[3]]
figs$interval_validation <- plot_grid(plotlist = plots)
figs$interval_validation
```

\newpage

### Figure `r as.character(fig_no)`: Display Testing Data - Proportion Resampled


```{r resampled validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[4]]
figs$proportion_validation <- plot_grid(plotlist = plots)
figs$proportion_validation
fig_no <- fig_no + 1
```

```{r save figs}
output_path <- file.path("output", config$out_tag)
dir.create(output_path)
output_path <- normalizePath(output_path)

save(figs, file = file.path(output_path, "figs.RDATA"))
```

