---
title: "SLAM Metabolic Class Figures"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../04_prediction_data/output/test_local.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE, eval = TRUE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(include = TRUE)

mega_pal <- rep(palette(), 40)
```

## Analysis

### Load Input File

```{r read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  params$input_path <- "../04_prediction_data/output/test_local.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
```

### Load Models

```{r load models}
load(input$cf_path)
load(input$datasets_path)
load(input$final_models_path)
config <- yaml::read_yaml(input$config_path)


# load in main_cat_surv
main_cat_surv <- read.csv(file = config$survival_dataset$path)
main_cat_surv <- as.data.frame(main_cat_surv)

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

### Class Analysis

#### Create Census Dataframe
```{r census data frame}
# will number classes so that all classes of list given to function
# are labelled uniquely
make_census <- function(cf_row, dataset) {
  model <- unlist(cf_row$final_model, recursive = FALSE)
  class_df <- model$pprob

  data <- dataset$data
  id <- dataset$id

  # add class_df to main
  census <- data[!duplicated(data[[id]]), ]
  census_cols <- dataset[c("id", "covariates", "factor")]
  census_cols <- unlist(census_cols, use.names = FALSE)
  census <- census[, census_cols]
  census <- merge(x = census, y = class_df, by = id)

  # add survival information to census
  surv_id <- config$survival_dataset$id

  census <- merge(x = census, y = main_cat_surv, by.x = id, by.y = surv_id)

  census
}


final_models$census <- lapply(seq_len(nrow(final_models)), function(i) {
  cf_row <- final_models[i, ]
  dataset <- datasets[[cf_row$dataset_index]]
  make_census(cf_row, dataset)
})
```

#### Create new class numbering

```{r class numbering}
reorder_class <- function(final_models, mo_names) {
  x <- final_models
  all_names <- x$model_name
  missing_names <- all_names[!(all_names %in% mo_names)]
  i <- 0
  for (name in mo_names) {
    # generate new_class
    census <- x[x$model_name == name, ]$census
    census <- unlist(census, recursive = FALSE)
    census$new_class <- as.character(census$class + i)

    # Create new prob based off of new_class
    probs_index <- grepl("^prob", names(census))
    probs <- names(census)[grepl("^prob", names(census))]
    probs <- stringr::str_split(probs, pattern = "prob", simplify = TRUE)[, 2]
    new_probs <- as.numeric(probs) + i
    new_probs <- paste0("prob", new_probs)
    names(census)[probs_index] <- new_probs

    uni_class <- length(unique(census$class))
    i <- uni_class + i
    x[x$model_name == name, ]$census[[1]] <- as.data.frame(census)
  }

  x
}

# get model_names that are not train_test
train_test_index <- (final_models$data_mod == "train_test")
renumber_train_test <- final_models[train_test_index, "model_name"]
renumber_rest <- final_models[!train_test_index, "model_name"]

final_models <- reorder_class(final_models, renumber_train_test)
final_models <- reorder_class(final_models, renumber_rest)

look <- final_models[, c("model_name", "census")]
```

#### Create Table 1
```{r source create table 1, file = 'R/t1.R'}
source("R/t1.R")
```

```{r create table 1}
final_models$t1_raw <- lapply(seq_along(final_models$census), function(i) {
  t1 <- t1(
    df = final_models$census[[i]],
    columns = final_models$covariates[[i]],
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )

  t1
})
```

#### Perform Chi-squared Test
```{r source chi squared test, file ='R/chi_test.R'}
source("R/chi_test.R")
```

```{r chi squred test}
chi_results <- lapply(seq_along(final_models$census), function(i) {
  chi <- chi_test(final_models$census[[i]],
    columns = final_models$covariates[[i]],
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )
  chi
})

final_models$chi_df <- chi_results$dfs
final_models$chi_test <- chi_results$tests
```

#### Create clean t1
```{r source clean t1, file = 'R/clean_t1.R'}
source("R/clean_t1.R")
```

```{r clean t1}
final_models$clean_t1 <- lapply(seq_along(final_models$census), function(i) {
  clean_t1 <- clean_t1(final_models$census[[i]],
    columns = final_models$covariates[[i]],
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )

  clean_t1
})
```

### Surival Analysis 

```{r source surv analysis, file = 'R/surv.R'}
source("R/surv.R")
```

```{r surv analysis}
final_models$census <- lapply(final_models$census, function(df) {
  df$Class <- df$new_class
  df
})

final_models$kap_plot <- lapply(seq_len(nrow(final_models)), function(i) {
  kap_plot <- kap_plot_all(
    df = final_models$census[[i]],
    var = "Class",
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event,
    ptitle = paste(final_models$oc_name[[i]],
      final_models$age_var_name[[i]],
      "KM",
      sep = " "
    )
  )

  kap_plot
})
```

### Create Cox Models

- Model 1: $ Surv ~ Class $
- Model 2: $ Surv ~ Class + Sex + Strain$
- Model 3: $ Surv ~ Class + Sex * Strain$
```{r source model_cox, file = 'R/model_cox.R'}
source("R/model_cox.R")
```

```{r cox models}
final_models$cox_models <- lapply(seq_len(nrow(final_models)), function(i) {
  model_cox(
    census = final_models$census[[i]],
    var = "Class",
    covariates = final_models$covariates[[i]],
    age_death = config$survival_dataset$age_death,
    censor = config$survival_dataset$event
  )
})
```

### Trajectory Fixed Effects Plotted

```{r source fixed effects, include = FALSE, file = 'R/traj_plot.R'}
source("R/traj_plot.R")
```

```{r fixed effects}
final_models$traj_pred <- lapply(seq_len(nrow(final_models)), function(i) {
  fixcov_names <- final_models$fixcov[[i]]
  fixcov <- rep(0, length(final_models$fixcov[[i]]))
  names(fixcov) <- fixcov_names
  traj_plot(
    df = final_models$dfs[[i]],
    mo = final_models$final_model[[i]],
    t1 = final_models$t1_raw[[i]],
    age_var = final_models$age_var[[i]],
    y_var = final_models$oc[[i]],
    title = paste("Predicted",
      final_models$oc_name[[i]],
      "Trajectories across",
      final_models$age_var_name[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    ),
    fixcov = fixcov
  )
  # can render this plot with ggdraw or plot_grid
  plot <- recordPlot()

  plot
})
```

### Create Observed Trajectory Plots

```{r source observed plots, include = FALSE, file = 'R/plot_obs.R'}
source("R/plot_obs.R")
```

```{r observed plots}
final_models$traj_obs <- lapply(seq_len(nrow(final_models)), function(i) {
  plot <- traj_obs(
    main = final_models$census[[i]],
    df = final_models$dfs[[i]],
    t1 = final_models$t1_raw[[i]],
    age_var = final_models$age_var[[i]],
    y_var = final_models$oc[[i]],
    id_var = final_models$subject[[i]],
    title = paste("Observed",
      final_models$oc_name[[i]],
      "Trajectory with",
      final_models$age_var_name[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    )
  )

  plot
})


lapply(final_models$traj_obs, print)
```

### Create Plots for Other Metabolic Outcomes by Class

```{r source other plot, file = 'R/plot_other.R'}
source("R/plot_combined.R")
```

```{r other metabolic outcomes}
for (model_name in final_models$model_name) {
  col_name <- paste0("plot_", model_name, "_across")
  a <- plot_combined_across(
    final_models,
    model_name
  )
  final_models[[col_name]] <- a
  print(a)
}
```

### Fit other metabolic outcomes with LME

- Model0 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model0.5 : $outcome = Age + Age^2 + Class + sex * Strain + (Age + Age^2 | Subject)$
- Model1 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model2 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Sex * Strain + (Age + Age^2 | Subject)$
- Model3 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Sex * Strain + (Age + Age^2 | Subject)$

```{r source mixed effects, file = 'R/model_other.R'}
source("R/model_combined.R")
```

```{r mixed effects}
for (model_name in final_models$model_name) {
  col_name <- paste0("model_", model_name, "_across")
  a <- model_combined_across(
    final_models,
    datasets,
    model_name,
    class = "new_class"
  )
  final_models[[col_name]] <- a
}
```

### Create Spaghettie Plots

```{r source spag plot, file = "R/plot_spag.R"}
source("R/plot_spag.R")
```

```{r spag plot}
final_models$spag_plot <- lapply(seq_len(nrow(final_models)), function(i) {
  plot_spag(
    data = final_models$dfs[[i]],
    x = final_models$age_var[[i]],
    y = final_models$oc[[i]],
    id = final_models$subject[[i]],
    title = paste(final_models$oc_name[[i]],
      final_models$age_var_name[[i]],
      "Spaghetti Plot",
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    )
  )
})
```

### Class Prediction

#### Class Prediction Raw Results
```{r source create prediction results, file = 'R/pred_class.R'}
source("R/pred_class.R")
```

```{r create prediction results}
# Get the names of the filtered data
prediction_data_names <- lapply(datasets, function(dataset) {
  names(dataset$prediction_data$data)
})
prediction_data_names <- unlist(prediction_data_names, use.names = FALSE)
prediction_columns <- unique(prediction_data_names)

prediction_names <- lapply(prediction_columns, strsplit, split = "_data")
prediction_names <- unlist(prediction_names)

# create a list of lists, where parent list is the filter applied, and the
# daughter lists provides the filtered dataframes for the corresponding
# row of final_models. This code does not take into account the case where
# certain filters are applied to one row, but not another
filter_list <- lapply(prediction_columns, function(filter) {
  row_list <- lapply(seq_len(nrow(final_models)), function(row_number) {
    data_name <- final_models[row_number, "data"]
    datasets[[data_name]]$prediction_data$data[[filter]]
  })
  row_list
})
names(filter_list) <- prediction_columns

for (filter in names(filter_list)) {
  final_models[[filter]] <- filter_list[[filter]]
}

prediction_results_names <- paste(prediction_names, "results", sep = "_")

prediction_results <- lapply(prediction_columns, function(col) {
  print(col)
  predict_class_model_list(
    nested_newdata_list = final_models[[col]],
    model_list = final_models$final_model,
    vars_list = final_models$data_age_vars,
    center = config$center,
    scale = config$scale,
    ref_data_list = final_models$dfs,
    names = final_models$model_name
  )
})

final_models[, prediction_results_names] <- prediction_results
```

### Sensititviy, Specificity, PPV, NPV, Tables

|  | Gold Yes | Gold No | total |
| --- | --- | --- | ---- |
| pred Yes | $TP = tp$ | $FP = pred\_class - tp$ | pred_class |
| pred No | $FN =og\_class - tp$ | $TN = n - (og\_class + pred\_class - tp)$ | n - pred_class | 
| total | og_class | n - og_class | n

- Sensitivity = 
$\frac{TP}{TP + FN} = 
\frac{tp}{tp + og\_class - tp} =
\frac{tp}{og\_class}$
- Specificity =
$\frac{TN}{FP + TN} = 
\frac{n - (og\_class + pred\_class - tp)}{pred\_class - tp + n - og\_class - pred\_class + tp)} = 
\frac{n - (og\_class + pred\_class - tp)}{n - og\_class}$
- Positive Predictive Value = 
$\frac{TP}{TP + FP} = 
\frac{tp}{tp + pred\_class - tp} = 
\frac{tp}{pred\_classs}$
- Negative Predictive Value = 
$\frac{TN}{FN + TN} = 
\frac{n - (og\_class + pred\_class -tp)}{og\_class - tp + n - og\_class - pred\_class + tp) } =
\frac{n - (og\_class + pred\_class - tp)}{n - pred\_class}$
- Accuracy = 
$\frac{TN + TP}{TN + TP + FN + FP} = 
\frac{2tp + n - og\_class - pred\_class}{n}$


```{r source prediciton tables, file = 'R/pred_table.R'}
source("R/pred_table.R")
```

```{r prediciton tables}
prediction_table_names <- paste(prediction_names,
  "table",
  sep = "_"
)

prediction_results_tables <- lapply(prediction_results_names, function(subset) {
  create_pred_table_subset(final_models, subset)
})

final_models[, prediction_table_names] <- prediction_results_tables
```

#### Dataframe for each subset type with prediction statistics

```{r source create prediciton dataframe for plotting, file = 'R/combine_pred_df.R'}
source("R/combine_pred_df.R")
```

```{r create prediciton dataframe for plotting}
prediction_df_names <- paste(prediction_names, "df", sep = "_")
pred_results <- lapply(
  prediction_table_names, combine_pred_df_subset,
  final_models = final_models
)

names(pred_results) <- prediction_df_names
final_models[, prediction_df_names] <- pred_results
```

#### Prepare prediciton dataframe for plotting

```{r source prepare prediction dataframe, file = 'R/rename_pred.R'}
source("R/rename_pred.R")
```

```{r prepare prediction dataframe}
final_models$filter_interval_df <- lapply(
  final_models$filter_interval_df,
  name_interval
)

final_models$filter_cumulative_df <- lapply(
  final_models$filter_cumulative_df,
  name_threshold
)

final_models$resampled_df <- lapply(
  final_models$resampled_df,
  name_sample
)

final_models$filter_window_df <- lapply(
  final_models$filter_window_df,
  name_threshold
)
```

#### Plot Prediction Results

```{r source plot pred_results, file = 'R/plot_pred.R'}
source("R/plot_pred.R")
```

```{r plot pred_results, fig.height=10, fig.width=5}
final_models$interval_plots <- lapply(
  seq_along(final_models$filter_interval_df),
  function(i) {
    plot_interval(final_models$filter_interval_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "Interval",
        sep = " "
      )
    )
  }
)

final_models$interval_plots

final_models$cumulative_plots <- lapply(
  seq_along(final_models$filter_cumulative_df),
  function(i) {
    plot_threshold(final_models$filter_cumulative_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "threshold",
        sep = " "
      )
    )
  }
)

final_models$cumulative_plots

final_models$resampled_plots <- lapply(
  seq_along(final_models$resampled_df),
  function(i) {
    plot_sample(final_models$resampled_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "sample",
        sep = " "
      )
    )
  }
)

final_models$resampled_plots

final_models$window_plots <- lapply(
  seq_along(final_models$filter_window_df),
  function(i) {
    plot_window(final_models$filter_window_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "window",
        sep = " "
      )
    )
  }
)

final_models$window_plots
```

#### Plot Accuracy

```{r plot accuracy}
final_models$interval_accuracy_plots <- lapply(
  seq_along(final_models$filter_interval_df),
  function(i) {
    plot_interval(final_models$filter_interval_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "Interval",
        sep = " "
      )
    )
  }
)

final_models$interval_accuracy_plots

final_models$cumulative_accuracy_plots <- lapply(
  seq_along(final_models$filter_cumulative_df),
  function(i) {
    plot_threshold(final_models$filter_cumulative_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "threshold",
        sep = " "
      )
    )
  }
)

final_models$cumulative_accuracy_plots

final_models$resampled_accuracy_plots <- lapply(
  seq_along(final_models$resampled_df),
  function(i) {
    plot_sample(final_models$resampled_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "sample",
        sep = " "
      )
    )
  }
)

final_models$resampled_accuracy_plots

final_models$window_accuracy_plots <- lapply(
  seq_along(final_models$filter_window_df),
  function(i) {
    plot_window(final_models$filter_window_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "window",
        sep = " "
      )
    )
  }
)

final_models$window_accuracy_plots
```

### Compute Accuracy Again as Double Check and Bootstrap

```{r source boot accuracy, file = 'R/boot_accuracy.R'}
source("R/boot_accuracy.R")
```

```{r compute accuracy}
final_models$interval_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_interval_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$cumulative_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_cumulative_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$window_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_window_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$resampled_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$resampled_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)
```

#### Combine bootstraped accuracy for each simulation type

```{r source combine accuracy, file = 'R/combine_boot_accuracy.R'}
source("R/combine_boot_accuracy.R")
```

```{r combine accuracy}
x <- final_models
x$interval_accuracy_df <- combine_boot_accuracies(x$interval_accuracy)
x$cumulative_accuracy_df <- combine_boot_accuracies(x$cumulative_accuracy)
x$resampled_accuracy_df <- combine_boot_accuracies(x$resampled_accuracy)
x$window_accuracy_df <- combine_boot_accuracies(x$window_accuracy)
final_models <- x
```

#### Prepare Accuracy Dataframe for pLotting

```{r prepare accuracy for plotting}
final_models$interval_accuracy_df <- lapply(
  final_models$interval_accuracy_df,
  name_interval
)


final_models$cumulative_accuracy_df <- lapply(
  final_models$cumulative_accuracy_df,
  name_threshold
)

final_models$resampled_accuracy_df <- lapply(
  final_models$resampled_accuracy_df,
  name_sample
)

final_models$window_accuracy_df <- lapply(
  final_models$window_accuracy_df,
  name_new_interval
)
```

#### Extract Mean and CI from Bootstrapped Accuracy

```{r source extraction, file = 'R/extract_accuracy_ci.R'}
source("R/extract_accuracy_ci.R")
```

```{r mean and ci}
final_models$interval_accuracy_df <- lapply(
  final_models$interval_accuracy_df,
  extract_accuracy_ci
)

final_models$cumulative_accuracy_df <- lapply(
  final_models$cumulative_accuracy_df,
  extract_accuracy_ci
)

final_models$resampled_accuracy_df <- lapply(
  final_models$resampled_accuracy_df,
  extract_accuracy_ci
)

final_models$window_accuracy_df <- lapply(
  final_models$window_accuracy_df,
  extract_accuracy_ci
)
```

#### Plot Accuracy Again With confidence Intervals

```{r source boot plot, file = 'R/plot_boot_accuracy.R'}
source("R/plot_boot_accuracy.R")
```

```{r plot accuracy with bootstrapped confidence intervals}
final_models$interval_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_interval(final_models$interval_accuracy_df[[i]],
      xlab = "Life Stage",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$interval_boot_accuracy_plot

final_models$window_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_window(final_models$window_accuracy_df[[i]],
      xlab = "Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$window_boot_accuracy_plot

final_models$cumulative_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_threshold(final_models$cumulative_accuracy_df[[i]],
      xlab = "Upper Bound Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Cumulative Classification Accuracy using",
        final_models$oc_name[[i]]
      ),
      legend_title = "Lower Bound"
    )
  }
)

final_models$cumulative_boot_accuracy_plot

final_models$resampled_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_sample(final_models$resampled_accuracy_df[[i]],
      xlab = "Lower Bound Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$resampled_boot_accuracy_plot
```

## Combined Metabolic Class

### Create Combined Cox 

```{r source combined_class_cox, file = 'R/combined_class_cox.R'}
source("R/combined_class_cox.R")
```

```{r combined survival}
# get model_names that are not train_test
train_test_index <- (final_models$data_mod == "train_test")
train_test_names <- final_models[train_test_index, "model_name"]
rest_names <- final_models[!train_test_index, "model_name"]

model_name_list <- list(rest_names, train_test_names)

combined_cox_models <- lapply(
  model_name_list,
  cox_combine_class,
  final_model_object = final_models,
  covariates = config$covariates,
  age_death = config$survival_dataset$age_death,
  censor = config$survival_dataset$event
)

names(combined_cox_models) <- lapply(model_name_list, paste, collapse = "-")

for (name in names(combined_cox_models)) {
  name_vector <- strsplit(x = name, split = "-")[[1]]
  model_index <- final_models$model_name %in% name_vector
  model_index <- match(name_vector, final_models$model_name)
  for (i in model_index) {
    final_models[i, "cox_combined_models"] <- list(combined_cox_models[name])
  }
}
```

### Determine get Hazard Predictions from test set

#### Predict Class Assignement for Test Dataframe

```{r source create prediction results, file = 'R/pred_class.R'}
source("R/pred_class.R")
```

```{r create prediction results}
# Get the names of the filtered data
prediction_data_names <- lapply(datasets, function(dataset) {
  names(dataset$prediction_test_data$data)
})
prediction_data_names <- unlist(prediction_data_names, use.names = FALSE)
prediction_columns <- unique(prediction_data_names)

prediction_names <- lapply(prediction_columns, strsplit, split = "_data")
prediction_names <- unlist(prediction_names)

# create a list of lists, where parent list is the filter applied, and the
# daughter lists provides the filtered dataframes for the corresponding
# row of final_models. This code does not take into account the case where
# certain filters are applied to one row, but not another
filter_list <- lapply(prediction_columns, function(filter) {
  row_list <- lapply(seq_len(nrow(final_models)), function(row_number) {
    data_name <- final_models[row_number, "data"]
    datasets[[data_name]]$prediction_test_data$data[[filter]]
  })
  row_list
})
names(filter_list) <- prediction_columns

for (filter in names(filter_list)) {
  final_models[[filter]] <- filter_list[[filter]]
}

prediction_names <- paste0(prediction_names, "_test")
prediction_results_names <- paste(prediction_names, "results", sep = "_")

prediction_results <- lapply(prediction_columns, function(col) {
  print(col)
  predict_class_model_list(
    nested_newdata_list = final_models[[col]],
    model_list = final_models$final_model,
    vars_list = final_models$data_age_vars,
    center = config$center,
    scale = config$scale,
    ref_data_list = final_models$dfs,
    names = final_models$model_name
  )
})

final_models[, prediction_results_names] <- prediction_results
```

#### Create new class numbering

```{r class numbering}
increment_class <- function(filtered_df, i = 0) {
  filtered_df$new_class <- as.character(filtered_df$class + i)

  # Create new prob based off of new_class
  probs_index <- grepl("^prob", names(filtered_df))
  probs <- names(filtered_df)[grepl("^prob", names(filtered_df))]
  probs <- unlist(lapply(strsplit(probs, split = "prob"), `[`, 2))
  new_probs <- as.numeric(probs) + i
  new_probs <- paste0("prob", new_probs)
  names(filtered_df)[probs_index] <- new_probs

  filtered_df
}

reorder_prediction_class <- function(final_models,
                                     model_names,
                                     filtered_df_name) {
  print(filtered_df_name) # delete
  i <- 0
  for (model_name in model_names) {
    print(model_name) # delete
    index <- final_models$model_name == model_name
    row <- final_models[index, ]
    filtered_dfs <- row[, filtered_df_name][[1]]
    if (is.list(filtered_dfs)) {
      filtered_dfs_inced <- lapply(
        filtered_dfs,
        increment_class,
        i
      )
    } else {
      filtered_dfs_inced <- NA
    }

    column <- final_models[, filtered_df_name]
    column[[which(index)]] <- filtered_dfs_inced
    final_models[[filtered_df_name]] <- column
    i <- as.numeric(row$ng) + i
  }
  final_models
}

# get model_names that are not train_test
train_test_index <- (final_models$data_mod == "train_test")
train_test_names <- final_models[train_test_index, "model_name"]
rest_names <- final_models[!train_test_index, "model_name"]

for (filtered_df_name in prediction_results_names) {
  final_models <- reorder_prediction_class(
    final_models = final_models,
    model_names = train_test_names,
    filtered_df_name = filtered_df_name
  )
}
```

#### Merge test_data prediction dataframes

```{r merge prediction dataframes}
merge_prediction_data <- function(x, y) {
  print(y$suffix)
  if (length(x$data_list) != length(y$data_list)) {
    stop("Data lists are not of the same length, cannot merge them")
  }
  z <- list()
  z$data_list <- mapply(merge,
    x = x$data_list, y = y$data_list,
    MoreArgs = list(by.x = x$id, by.y = y$id, suffixes = c(x$suffix, y$suffix)),
    SIMPLIFY = FALSE
  )
  z$id <- x$id
  z$suffix <- ""

  z
}


merge_prediction_data_group <- function(
    final_models,
    model_names,
    filtered_df_name) {
  index <- final_models$model_name %in% model_names
  desired_dfs <- final_models[index, filtered_df_name]
  desired_ids <- final_models[index, "subject"]
  desired_suffixes <- paste0("_", final_models[index, "oc"])

  if (all(is.na(desired_dfs))) {
    merged_data <- NA
  } else {
    reduce_list <- lapply(seq_len(length(desired_dfs)), function(i) {
      list(
        data_list = desired_dfs[[i]],
        id = desired_ids[[i]],
        suffix = desired_suffixes[[i]]
      )
    })
    merged_data <- Reduce(f = merge_prediction_data, x = reduce_list)$data_list

    # the last element name improperly following reduce function, thus:
    merged_data <- lapply(merged_data, function(data) {
      last_class_name <- paste0("class", tail(desired_suffixes, n = 1))
      last_new_class_name <- paste0("new_class", tail(desired_suffixes, n = 1))
      colnames(data)[colnames(data) == "class"] <- last_class_name
      colnames(data)[colnames(data) == "new_class"] <- last_new_class_name
      data
    })
  }

  # create new column to store data
  column <- final_models[, filtered_df_name]
  for (row in which(index)) {
    column[[row]] <- merged_data
  }
  new_column_name <- gsub("_results$", "", filtered_df_name)
  new_column_name <- paste0(new_column_name, "_merged")
  final_models[[new_column_name]] <- column
  final_models
}

train_test_index <- (final_models$data_mod == "train_test")
train_test_names <- final_models[train_test_index, "model_name"]

for (filtered_df_name in prediction_results_names) {
  final_models <- merge_prediction_data_group(
    final_models = final_models,
    model_names = train_test_names,
    filtered_df_name = filtered_df_name
  )
}
```

#### Create test census dataframe

```{r create test census dataframe}
make_test_census <- function(cf_row, dataset) {
  if (is.null(dataset$test_data)) {
    return(NA)
  }
  data <- dataset$test_data
  id <- dataset$id

  # create census from data
  census <- data[!duplicated(data[[id]]), ]
  census_cols <- dataset[c("id", "covariates", "factor")]
  census_cols <- unlist(census_cols, use.names = FALSE)

  census <- census[, census_cols]
  # add survival information to census
  surv_id <- config$survival_dataset$id

  census <- merge(x = census, y = main_cat_surv, by.x = id, by.y = surv_id)

  census
}


final_models$test_census <- lapply(seq_len(nrow(final_models)), function(i) {
  cf_row <- final_models[i, ]
  dataset <- datasets[[cf_row$dataset_index]]
  make_test_census(cf_row, dataset)
})
```

```{r merge test census}
merge_test_census_apply <- function(data_list, census, id) {
  if (all(is.na(data_list))) {
    data_list <- NA
    warning("data_list is all NA, returning NA")
  } else {
    data_list <- lapply(data_list, merge, y = census, by = id)
  }
  data_list
}

merge_test_census_column <- function(final_models, filtered_df_name) {
  final_models_test <- final_models[final_models$data_mod == "train_test", ]
  censuses <- final_models_test[, "test_census"]
  data_lol <- final_models_test[, filtered_df_name]
  ids <- final_models_test[, "subject"]
  data_lol <- mapply(merge_test_census_apply,
    data_list = data_lol, census = censuses, id = ids,
    SIMPLIFY = FALSE
  )

  final_models[[filtered_df_name]] <- data_lol
  final_models
}

merged_prediction_names <- paste0(prediction_names, "_merged")

for (filtered_df_name in merged_prediction_names) {
  final_models <- merge_test_census_column(
    final_models = final_models,
    filtered_df_name = filtered_df_name
  )
}
```

#### We need to predict hazard ratios for our test sets

```{r test mortaility prediction using train test}
predict_cox <- function(test_data,
                        model,
                        type = c("lp", "risk"),
                        suffix = "") {
  print(nrow(test_data))
  if (nrow(test_data) == 0) {
    warning("test_data has 0 obs, cannot predict; returning itself")
    return(test_data)
  }
  for (t in type) {
    name <- paste0(t, suffix)
    test_data[[name]] <- predict(model, test_data, type = t)
  }

  test_data
}

predict_cox_apply <- function(test_data, model_list, type = c("lp", "risk")) {
  for (i in seq_len(length(model_list))) {
    test_data <- predict_cox(
      test_data = test_data,
      model = model_list[[i]],
      type = type,
      suffix = i
    )
  }

  test_data
}

predict_cox_apply2 <- function(test_data_list,
                               model_list,
                               type = c("lp", "risk")) {
  if (is.list(test_data_list)) {
    test_data_list <- lapply(
      X = test_data_list,
      FUN = predict_cox_apply,
      model_list = model_list,
      type = type
    )
  } else {
    test_data_list <- NA
    warning("test_data_list is not a list, returning NA")
  }

  test_data_list
}

predict_cox_final_model <- function(final_models,
                                    filtered_df_name,
                                    type = c("lp", "risk")) {
  test_data_lol <- final_models[, filtered_df_name]
  model_lol <- final_models[, "cox_combined_models"]
  data_mods <- final_models[, "data_mod"]

  test_data_lol <- lapply(seq_len(nrow(final_models)), function(row_no) {
    print(row_no)
    data_mod <- data_mods[[row_no]]
    if (data_mod == "train_test") {
      test_data_list <- predict_cox_apply2(
        test_data_list = test_data_lol[[row_no]],
        model_list = model_lol[[row_no]],
        type = type
      )
    } else {
      test_data_list <- test_data_lol[[row_no]]
    }

    test_data_list
  })

  final_models[[filtered_df_name]] <- test_data_lol
  final_models
}

final_models_test <- final_models[final_models$data_mod == "train_test", ]

for (filtered_df_name in merged_prediction_names) {
  print(filtered_df_name)
  final_models <- predict_cox_final_model(
    final_models = final_models,
    filtered_df_name = filtered_df_name,
    type = c("lp", "risk")
  )
}
```

#### Determine Calibration by Regressing PI on Test outcomes

```{r create new dataframes from the predictions}
get_slope_concord <- function(test_data, age_death, censor) {
  surv_obj <- survival::Surv(
    time = test_data[[age_death]], event = test_data[[censor]]
  )

  col_names <- names(test_data)
  # get columns that have lp in them (the linear predictor)
  index <- grepl(pattern = "lp\\d", x = col_names)
  lp_cols <- col_names[index]

  sum_concord_list <- lapply(lp_cols, function(lp) {
    form <- as.formula(paste0("surv_obj ~ ", lp))
    cox <- survival::coxph(formula = form, data = test_data)
    sum <- as.data.frame(summary(cox)$coefficients)
    names(sum) <- c("coef", "exp_coef", "se_coef", "z", "p")
    names(sum) <- paste(lp, names(sum), sep = "_")

    concord <- as.data.frame(t(summary(cox)$concordance))
    names(concord) <- c("concord", "se_concord")
    names(concord) <- paste(lp, names(concord), sep = "_")
    sum_concord <- cbind(sum, concord)
    sum_concord
  })

  sum_df_row <- do.call(what = cbind, args = sum_concord_list)

  sum_df_row
}

get_slope_concord_apply <- function(test_data_list, age_death, censor) {
  if (is.list(test_data_list)) {
    test_data_df <- lapply(
      X = test_data_list,
      FUN = get_slope_concord,
      age_death = age_death,
      censor = censor
    )

    # for when there are intervals with no observations
    null_index <- unlist(lapply(test_data_df, is.null))

    test_data_df <- do.call(
      what = rbind,
      args = test_data_df
    )

    test_data_df$subset_name <- names(test_data_list)[!null_index]
  } else {
    test_data_df <- NA
    warning("test_data_list is not a list, returning NA")
  }

  test_data_df
}

get_slope_concord_final_models <- function(final_models,
                                           filtered_df_name,
                                           age_death,
                                           censor) {
  column <- final_models[, filtered_df_name]
  train_test_index <- final_models$data_mod == "train_test"
  column[train_test_index] <- lapply(column[train_test_index],
    get_slope_concord_apply,
    age_death = age_death, censor = censor
  )
  column[!train_test_index] <- NA

  new_column_name <- gsub("_merged$", "", filtered_df_name)
  new_column_name <- paste0(new_column_name, "_summary")

  final_models[[new_column_name]] <- column

  final_models
}

filtered_df_name <- merged_prediction_names[[1]]

for (filtered_df_name in merged_prediction_names) {
  print(filtered_df_name)
  final_models <- get_slope_concord_final_models(
    final_models = final_models,
    filtered_df_name = filtered_df_name,
    age_death = config$survival_dataset$age_death,
    censor = config$survival_dataset$event
  )
}
```

#### Generated columns for prediction data plotting

```{r prediction data renaming}
name_window <- function(data, col_name) {
  if (all(is.na(data))) {
    warning("data is NA, returning data")
    return(data)
  }
  col <- data[[col_name]]
  col_split <- strsplit(x = col, split = "\\[|,| |\\]")
  lower_bound <- lapply(col_split, function(i) i[[2]])
  lower_bound <- as.numeric(unlist(lower_bound))
  upper_bound <- lapply(col_split, function(i) i[[4]])
  upper_bound <- as.numeric(unlist(upper_bound))
  window_size <- upper_bound - lower_bound

  data[["upper_bound"]] <- upper_bound
  data[["lower_bound"]] <- lower_bound
  data[["window_size"]] <- window_size
  data[["midpoint"]] <- window_size / 2 + lower_bound

  data
}

name_cumulative <- function(data, col_name) {
  if (all(is.na(data))) {
    warning("data is NA, returning data")
    return(data)
  }
  col <- data[[col_name]]
  col_split <- strsplit(x = col, split = "\\[|,| |\\]")
  lower_bound <- lapply(col_split, function(i) i[[2]])
  lower_bound <- as.numeric(unlist(lower_bound))
  upper_bound <- lapply(col_split, function(i) i[[4]])
  upper_bound <- as.numeric(unlist(upper_bound))

  data[["upper_bound"]] <- upper_bound
  data[["lower_bound"]] <- lower_bound
  data[["midpoint"]] <- (upper_bound - lower_bound) / 2 + lower_bound

  data
}

final_models$filter_cumulative_test_summary <- lapply(
  final_models$filter_cumulative_test_summary,
  name_cumulative,
  col_name = "subset_name"
)

final_models$filter_window_test_summary <- lapply(
  final_models$filter_window_test_summary,
  name_window,
  col_name = "subset_name"
)
```

#### Create plots of concordances and regression on PI

```{r create concordance plots}
# TODO: Create plots of one combined cox model.
# The final goal is to have a 2x3 figure where the type
# of data filter determines the columns and the concordance
# and correlation of PI form the two rows

# TODO: We need to adjust the our combined_cox_models to account for comparing
# absolute vs. class vs. class + absolute

data <- final_models$filter_window_test_summary[[4]]

outcome <- "lp1_exp_coef"
window_size <- "window_size"
midpoint <- "midpoint"
se <- "lp1_se_coef"

plot_window_concord_lp <- function(data, outcome, window_size, midpoint, se) {
  ggplot2::ggplot(
    data = data,
    mapping = ggplot2::aes(
      x = as.numeric(.data[[midpoint]]),
      y = as.numeric(.data[[outcome]]),
      color = factor(.data[[window_size]])
    )
  ) +
    ggplot2::geom_line(stat = "identity") +
    ggplot2::geom_errorbar(
      ggplot2::aes(
        x = .data[[midpoint]],
        ymin = .data[[outcome]] - 1.96 * .data[[se]],
        ymax = .data[[outcome]] + 1.96 * .data[[se]],
        stat = "identity",
        width = 0.4
      )
    )
}
```

### Output CSV with idno's and class

```{r idnos and class}
# we want to output class, idno, tag to a csv so that collaborators can easily
# work with the data

setwd(file.path("output", params$model_dir))
dir.create("csv")
setwd("csv")

for (model in seq_len(nrow(final_models))) {
  census <- final_models$census[[model]]
  model_name <- final_models$model_name[[model]]
  census <- census[, c("class", "idno", "tag", "sex", "strain", "cohort.y")]
  file_name <- paste0(model_name, ".csv")
  write.csv(census, file = file_name)
}
```

### Save R Objects

#### Prune Final Models Before Saving 

```{r prune}
final_models <- final_models %>%
  select(
    -c(
      interval_accuracy_plots,
      threshold_accuracy_plots,
      sample_accuracy_plots
    ),
    -c(
      ends_with("results_table"),
      ends_with("accuracy")
    ),
    -c(final_model)
  )

final_models <- final_models %>%
  select(-c(
    ends_with("_df"),
    ends_with("_results")
  ))

final_models <- final_models %>%
  select(-interval, -subset, -threshold, -sample)

final_models <- final_models %>%
  filter(model_name == "bw_main|bw_age_wk" |
    model_name == "fat_main|fat_age_wk" |
    model_name == "gluc_main|gluc_age_wk")
```

#### Save 

```{r save R objects}
setwd(file.path("output", params$model_dir))
dir.create("data")
setwd("data")

save(cf, file = "cf.RDATA")
save(final_models, file = "final_models.RDATA")
save(config, file = "config.RDATA")
save(df_list, file = "df_list.RDATA")
save(dfs_prediction, file = "dfs_prediction.RDATA")
```
