---
title: "SLAM Metabolic Class Figures"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  model_dir: "20220617_142335_test"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE, eval = TRUE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(include = TRUE)

mega_pal <- rep(palette(), 40)

```

## Analysis

### Load Models

```{r load models}
path <- normalizePath(file.path("..", "04_model_select", "output", params$model_dir))

load_output <- function(file_name) {
  file_path <- normalizePath(file.path(path, paste0(file_name, ".RDATA")))
  load(file = file_path, envir = .GlobalEnv)
}

load_output("cf")
load_output("df_list")
load_output("final_models")
load_output("models")
load_output("config")

# load in main_cat_surv
load(file = "../04b_healthcard_cod/output/main_cat_surv.RDATA")
# load in health
load(file = "../04b_healthcard_cod/output/health.RDATA")
# load in dvr
load(file = "../04b_healthcard_cod/output/dvr.RDATA")

# load in prediction data
load(file = "../04c_prediction_data/output/dfs_prediction.RDATA")

final_models$dfs <- lapply(final_models$data, function(name) {
  df_list[[name]]
})
```

#### Name Model Variables

```{r source naming, file = 'R/name_model_args.R'}
source("R/name_model_args.R")
```

```{r naming}
final_models_info <- final_models %>%
  select(-final_model)
```

#### Healthcard Cleaning
```{r healthcard}

# now we have to add the health card information. lets convert it to wide format
# so that we can add it to the subject summary
# create wave for health
health <- health %>%
  group_by(tag) %>%
  arrange(date_created) %>%
  mutate(wave = row_number()) %>%
  ungroup()
health_wide <- pivot_wider(health,
  id_cols = c("idno", "tag", "dod"),
  names_from = "wave",
  names_prefix = "condition_",
  values_from = "condition_clean"
)

health_wide_heat <- pivot_wider(health,
  id_cols = c("idno", "tag", "dod"),
  names_from = "condition_clean",
  values_from = "condition_clean"
)

health_wide_heat <- health_wide_heat

col_of_interest <- 4:length(health_wide_heat)

health_wide_heat[, col_of_interest] <- lapply(
  health_wide_heat[, col_of_interest],
  function(col) {
    ifelse(col == "NULL", 0, 1)
  }
)

count <- as.data.frame(apply(health_wide_heat[, col_of_interest], 2, sum))

count

names(count) <- "n"
count$condition <- rownames(count)
rownames(count) <- NULL

# select interesting healthcard
count <- arrange(count, -n)

# save healthcard count
write.csv(count, file = "output/conditions_and_counts.csv")
```

### Class Analysis

#### Create Summary Dataframe
```{r create summary dataframe}
merge_dvr <- dvr %>%
  select(idno, cod, dod, cod_coded, cod_1:cod_4_parens)

merge_health <- health_wide_heat %>%
  select(-dod, -idno)

subject_sum <- main_cat_surv %>%
  select(
    idno,
    tag,
    sex,
    strain,
    dead_censor,
    fu_age_wk,
    le_wk,
    cod,
    lastdate,
    maxdate,
    tod
  ) %>%
  mutate(idno = as.numeric(idno)) %>%
  left_join(merge_dvr, by = "idno")

# check na's after merging
apply(apply(subject_sum, 2, is.na), 2, sum)

# obs goe from 1304 to 1310, so their are duplicates, lets find these duplicates
# and choose the correct one
id <- subject_sum$idno
subject_sum[duplicated(id) | duplicated(id, fromLast = TRUE), ]

# if there a duplicates, we are going to take the observation with the cod.y
# that is not NA or Undetermined. If there is no NA or undetermined, we are
# going to take the observation with the later dod
subject_sum <- subject_sum %>%
  group_by(idno) %>%
  # cod_effective is 0 if "undetermined" or is.na
  mutate(cod_effective = ifelse(cod.y == "Undetermined" | is.na(cod.y), 0, 1)) %>%
  arrange(desc(cod_effective), desc(dod)) %>%
  distinct(idno, .keep_all = TRUE) %>%
  ungroup() %>%
  as.data.frame() %>%
  rename(
    cod_census = cod.x,
    cod_dvr = cod.y
  )

subject_sum <- subject_sum %>%
  left_join(merge_health, by = "tag")
```

#### Create Census Dataframe
```{r census data frame}
# will number classes so that all classes of list given to function
# are labelled uniquely
make_census <- function(cf_row) {
  model <- unlist(cf_row$final_model, recursive = FALSE)
  class_df <- model$pprob

  main <- df_list[[cf_row$data]]

  # add class_df to main
  census <- main %>%
    mutate(idno = as.numeric(idno)) %>%
    left_join(class_df, by = "idno") %>%
    distinct(idno, .keep_all = T) %>%
    select(class, idno, cohort, starts_with("prob")) %>%
    filter(!is.na(class)) %>%
    mutate(idno = as.character(idno))

  # add survival information to census
  census <- census %>%
    mutate(idno = as.numeric(idno)) %>%
    left_join(subject_sum, by = "idno") %>%
    mutate(
      age_wk_death = le_wk,
      dead_nat = dead_censor
    )

  # return census
  census
}


final_models$census <- lapply(seq_len(nrow(final_models)), function(i) {
  cf_row <- final_models[i, ]
  make_census(cf_row)
})
```


#### Create Table 1
```{r source create table 1, file = 'R/t1.R'}
source("R/t1.R")
```

```{r create table 1}
final_models$t1_raw <- lapply(1:length(final_models$census), function(i) {
  cat(i, "- ")
  t1 <- t1(final_models$census[[i]], columns = c("sex", "strain"))
  cat("\n")

  t1
})
```

#### Perform Chi-squared Test
```{r source chi squared test, file ='R/chi_test.R'}
source("R/chi_test.R")
```

```{r chi squred test}
chi_results <- lapply(final_models$census,
  chi_test,
  columns = c("sex", "strain")
)

final_models$chi_df <- chi_results$dfs
final_models$chi_test <- chi_results$tests
```

#### Create clean t1
```{r source clean t1, file = 'R/clean_t1.R'}
source('R/clean_t1.R')
```

```{r clean t1}
final_models$clean_t1 <- lapply(final_models$census,
  clean_t1,
  columns = c("sex", "strain")
)
```

### Surival Analysis 
```{r source surv analysis, file = 'R/surv.R'}
source('R/surv.R')
```

```{r surv analysis}
final_models$kap_plot <- lapply(1:nrow(final_models), function(i) {
  if (!is.na(final_models$t1[[i]])) {
    kap_plot <- kap_plot_all(final_models$census[[i]],
      var = "class",
      ptitle = paste(final_models$oc_name[[i]],
        final_models$age_var_name[[i]],
        "KM",
        sep = " "
      )
    )
  } else {
    kap_plot <- NA
  }

  kap_plot
})

```

### Trajectory Fixed Effects Plotted

```{r source fixed effects, include = FALSE, file = 'R/traj_plot.R'}
source('R/traj_plot.R')
```

```{r fixed effects}
if (is.null(config$female) & is.null(config$het3)) {
  fixcov <- c(female = 0, het3 = 0)
} else if (is.null(config$female)) {
  fixcov <- c(het3 = 0)
} else if (is.null(config$het3)) {
  fixcov <- c(female = 0)
} else {
  fixcov <- NULL
}

final_models$traj_pred <- lapply(1:nrow(final_models), function(i) {
  if (!is.na(final_models$t1[[i]])) {
    traj_plot(
      df = final_models$dfs[[i]],
      mo = final_models$final_model[[i]],
      t1 = final_models$t1_raw[[i]],
      age_var = final_models$age_var[[i]],
      y_var = final_models$oc[[i]],
      title = paste("Predicted",
        final_models$oc_name[[i]],
        "Trajectories across",
        final_models$age_var_name[[i]],
        sep = " "
      ),
      ylab = paste(final_models$oc_name[[i]],
        final_models$oc_units[[i]],
        sep = " "
      ),
      xlab = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        sep = " "
      ),
      fixcov = fixcov
    )
    # can render this plot with ggdraw or plot_grid
    plot <- recordPlot()
  } else {
    plot <- NA
  }

  plot
})
```

### Create Observed Trajectory Plots

```{r source observed plots, include = FALSE, file = 'R/traj_plot.R'}
source('R/traj_plot.R')
```

```{r observed plots}
final_models$traj_obs <- lapply(1:nrow(final_models), function(i) {
  if (final_models$ng[[i]] != 1) {
    plot <- traj_obs(
      main = final_models$census[[i]],
      df = final_models$dfs[[i]],
      t1 = final_models$t1_raw[[i]],
      age_var = final_models$age_var[[i]],
      y_var = final_models$oc[[i]],
      title = paste("Observed",
        final_models$oc_name[[i]],
        "Trajectory with",
        final_models$age_var_name[[i]],
        sep = " "
      ),
      ylab = paste(final_models$oc_name[[i]],
        final_models$oc_units[[i]],
        sep = " "
      ),
      xlab = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        sep = " "
      )
    )
  } else {
    plot <- NA
  }

  plot
})


lapply(final_models$traj_obs, print)

```

### Create Spaghettie Plots

```{r source spag plot, file = "R/plot_spag.R"}
source("R/plot_spag.R")
```

```{r spag plot}
lapply(1:nrow(final_models), function(i) {
  plot_spag(
    data = final_models$dfs[[i]],
    x = final_models$age_var[[i]],
    y = final_models$oc[[i]],
    id = final_models$subject[[i]],
    title = paste(final_models$oc_name[[i]],
      final_models$age_var_name[[i]],
      "Spaghetti Plot",
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]], final_models$age_var_units[[i]], sep = " "),
    ylab = paste(final_models$oc_name[[i]], final_models$oc_units[[i]], sep = " ")
  )
})
```

### Class Prediction

#### Class Prediction Raw Results
```{r source create prediction results, file = 'R/pred_class.R'}
source('R/pred_class.R')
```

```{r create prediction results}
# add prediction dataframes to final_models
for (subset_name in names(dfs_prediction)) {
  subset <- list()
  for (i in seq_len(nrow(final_models))) {
    data_name <- final_models$data[[i]]
    dfs_pred_list <- dfs_prediction[[subset_name]][[data_name]]
    subset[i] <- list(dfs_pred_list)
  }
  final_models[[subset_name]] <- subset
}

prediction_columns <- names(dfs_prediction)[-1]
prediction_results_names <- paste(prediction_columns, "results", sep = "_")

prediction_results <- lapply(prediction_columns, function(col) {
  predict_class_model_list(
    nested_newdata_list = final_models[[col]],
    model_list = final_models$final_model,
    vars = c("age_wk", "age_wk2", "per_age_wk", "per_age_wk2"),
    center = TRUE,
    scale = FALSE,
    ref_data = final_models$dfs,
    names = final_models$model_name
  )
})

final_models[, prediction_results_names] <- prediction_results
```

#### Sensititviy, Specificity, PPV, NPV, Tables

|  | Gold Yes | Gold No | total |
| --- | --- | --- | ---- |
| pred Yes | $TP = tp$ | $FP = pred\_class - tp$ | pred_class |
| pred No | $FN =og\_class - tp$ | $TN = n - (og\_class + pred\_class - tp)$ | n - pred_class | 
| total | og_class | n - og_class | n

- Sensitivity = 
$\frac{TP}{TP + FN} = 
\frac{tp}{tp + og\_class - tp} =
\frac{tp}{og\_class}$
- Specificity =
$\frac{TN}{FP + TN} = 
\frac{n - (og\_class + pred\_class - tp)}{pred\_class - tp + n - og\_class - pred\_class + tp)} = 
\frac{n - (og\_class + pred\_class - tp)}{n - og\_class}$
- Positive Predictive Value = 
$\frac{TP}{TP + FP} = 
\frac{tp}{tp + pred\_class - tp} = 
\frac{tp}{pred\_classs}$
- Negative Predictive Value = 
$\frac{TN}{FN + TN} = 
\frac{n - (og\_class + pred\_class -tp)}{og\_class - tp + n - og\_class - pred\_class + tp) } =
\frac{n - (og\_class + pred\_class - tp)}{n - pred\_class}$


```{r source prediciton tables, file = 'R/pred_table.R'}
source('R/pred_table.R')
```

```{r prediciton tables}
prediction_results_table_names <- paste(prediction_results_names,
  "table",
  sep = "_"
)

prediction_results_tables <- lapply(
  prediction_results_names,
  function(subset) {
    create_pred_table_subset(subset,
      subject = "idno"
    )
  }
)

final_models[, prediction_results_table_names] <- prediction_results_tables
```

#### Dataframe for each subset type with prediction statistics

```{r source create prediciton dataframe for plotting, file = 'R/combine_pred_df.R'}
source('R/combine_pred_df.R')
```

```{r create prediciton dataframe for plotting}
prediction_results_df_names <- paste(prediction_results_names,
  "df",
  sep = "_"
)
pred_results <- lapply(
  prediction_results_table_names,
  combine_pred_df_subset
)

names(pred_results) <- prediction_results_df_names
final_models[, prediction_results_df_names] <- pred_results
```

#### Prepare prediciton dataframe for plotting

```{r prepare prediction dataframe, file = 'R/rename_pred.R'}
source('R/rename_pred.R')
```

```{r prepare prediction dataframe}
final_models$interval_results_df <- lapply(
  final_models$interval_results_df,
  name_interval
)


final_models$threshold_results_df <- lapply(
  final_models$threshold_results_df,
  name_threshold
)

final_models$sample_results_df <- lapply(
  final_models$sample_results_df,
  name_sample
)
```

#### Plot Prediction Results

```{r source plot pred_results, file = 'R/plot_pred.R'}
source('R/plot_pred.R')
```

```{r plot pred_results, fig.height=10, fig.width=5}
final_models$interval_plots <- lapply(
  seq_along(final_models$interval_results_df),
  function(i) {
    plot_interval(final_models$interval_results_df[[i]],
      title = paste(final_models$data_name[[i]],
                    final_models$age_var_name[[i]],
                    sep = " "),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
                final_models$age_var_units[[i]],
                "Interval",
                sep = " ")
    )
  }
)

final_models$interval_plots

final_models$threshold_plots <- lapply(
  seq_along(final_models$threshold_results_df),
  function(i) {
    plot_threshold(final_models$threshold_results_df[[i]],
      title = paste(final_models$data_name[[i]],
                    final_models$age_var_name[[i]],
                    sep = " "),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
                final_models$age_var_units[[i]],
                "threshold",
                sep = " ")
    )
  }
)

final_models$threshold_plots

final_models$sample_plots <- lapply(
  seq_along(final_models$sample_results_df),
  function(i) {
    plot_sample(final_models$sample_results_df[[i]],
      title = paste(final_models$data_name[[i]],
                    final_models$age_var_name[[i]],
                    sep = " "),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
                final_models$age_var_units[[i]],
                "sample",
                sep = " ")
    )
  }
)

final_models$sample_plots

```

### Save R Objects

```{r save R objects}
dir.create("output_data")
setwd("output_data")

save(cf, file = "cf.RDATA")
save(final_models, file = "final_models.RDATA")
save(config, file = "config.RDATA")
save(df_list, file = "df_list.RDATA")
save(dfs_prediction, file = "dfs_prediction.RDATA")
```