---
title: "SLAM Metabolic Class Figures"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  model_dir: "test_local"
  yaml: "yaml/test.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE, eval = TRUE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(include = TRUE)

mega_pal <- rep(palette(), 40)
```

## Analysis

### Load Models

```{r load models}
load_output <- function(path, file_name) {
  file_path <- normalizePath(file.path(path, paste0(file_name, ".RDATA")))
  load(file = file_path, envir = .GlobalEnv)
}

path_04 <- normalizePath(file.path(
  "..",
  "04_model_select",
  "output",
  params$model_dir
))

path_04c <- normalizePath(file.path(
  "..",
  "04c_prediction_data",
  "output",
  params$model_dir
))

load_output(path_04, "cf")
load_output(path_04c, "datasets")
load_output(path_04, "final_models")
load_output(path_04, "config")


# load in main_cat_surv
load(file = "../04b_healthcard_cod/output/main_cat_surv.RDATA")
main_cat_surv <- as.data.frame(main_cat_surv)
# load(file = "../01_dataset/output/main_cat_surv.RDATA")
# load in health
load(file = "../04b_healthcard_cod/output/health.RDATA")
# load in dvr
load(file = "../04b_healthcard_cod/output/dvr.RDATA")

# load in prediction data
# load(file = "../04c_prediction_data/output/dfs_prediction.RDATA")

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

### Read Yaml
```{r load yaml}
yaml_models <- yaml::read_yaml(params$yaml)
```

### Drop all models that will not be used
The models that are used for figure generation are defined in the yaml file
```{r drop useless}
model_keep <- character()
for (group in yaml_models) {
  model_keep <- c(model_keep, group$model_names)
}
final_models <- final_models[final_models$model_name %in% model_keep, ]

# Make rownames null so that the rownumbers are correct
rownames(final_models) <- NULL
```

```{r read in census}
census <- read.csv(file = "../01_dataset/data/census.csv") %>%
  select(cohort, tag, sex, strain, idno) %>%
  filter(cohort <= 10)
```
#### Name Model Variables

```{r source naming, file = 'R/name_model_args.R'}
source("R/name_model_args.R")
```

```{r naming}
final_models_info <- final_models %>%
  select(-final_model)
```

#### Healthcard Cleaning
```{r healthcard}
# now we have to add the health card information. lets convert it to wide format
# so that we can add it to the subject summary
# create wave for health
health <- health %>%
  group_by(tag) %>%
  arrange(date_created) %>%
  mutate(wave = row_number()) %>%
  ungroup()
health <- as.data.frame(health)

health_wide <- pivot_wider(health,
  id_cols = c("idno", "tag", "dod"),
  names_from = "wave",
  names_prefix = "condition_",
  values_from = "condition_clean"
)
health_wide <- as.data.frame(health_wide)

health_wide_heat <- pivot_wider(health,
  id_cols = c("idno", "tag", "dod"),
  names_from = "condition_clean",
  values_from = "condition_clean"
)
health_wide <- as.data.frame(health_wide)

health_wide_heat <- health_wide_heat

col_of_interest <- 4:length(health_wide_heat)

health_wide_heat[, col_of_interest] <- lapply(
  health_wide_heat[, col_of_interest],
  function(col) {
    ifelse(col == "NULL", 0, 1)
  }
)

count <- as.data.frame(apply(health_wide_heat[, col_of_interest], 2, sum))

count

names(count) <- "n"
count$condition <- rownames(count)
rownames(count) <- NULL

# select interesting healthcard
count <- arrange(count, -n)

# save healthcard count
write.csv(count, file = "output/conditions_and_counts.csv")
```

### Class Analysis

#### Create Summary Dataframe
```{r create summary dataframe}
merge_dvr <- dvr %>%
  select(idno, cod, dod, cod_coded, cod_1:cod_4_parens)

merge_health <- health_wide_heat %>%
  select(-dod, -idno)

main_cat_surv_prep <- main_cat_surv %>%
  select(
    -tag,
    -sex,
    -strain,
    dead_censor,
    fu_age_wk,
    le_wk,
    cod,
    lastdate,
    maxdate,
    tod
  )
main_cat_surv_prep <- as.data.frame(main_cat_surv_prep)

subject_sum <- census %>%
  left_join(main_cat_surv_prep, by = "idno") %>%
  mutate(idno = as.numeric(idno)) %>%
  left_join(merge_dvr, by = "idno")


# check na's after merging
apply(apply(subject_sum, 2, is.na), 2, sum)

# obs goe from 1304 to 1310, so their are duplicates, lets find these duplicates
# and choose the correct one
id <- subject_sum$idno
subject_sum[duplicated(id) | duplicated(id, fromLast = TRUE), ]

# if there a duplicates, we are going to take the observation with the cod.y
# that is not NA or Undetermined. If there is no NA or undetermined, we are
# going to take the observation with the later dod
subject_sum <- subject_sum %>%
  group_by(idno) %>%
  # cod_effective is 0 if "undetermined" or is.na
  mutate(cod_effective = ifelse(cod.y == "Undetermined" | is.na(cod.y), 0, 1)) %>%
  arrange(desc(cod_effective), desc(dod)) %>%
  distinct(idno, .keep_all = TRUE) %>%
  ungroup() %>%
  as.data.frame() %>%
  rename(
    cod_census = cod.x,
    cod_dvr = cod.y
  ) %>%
  mutate(
    female = ifelse(sex == "F", 1, 0),
    het3 = ifelse(strain == "HET3", 1, 0)
  )

subject_sum <- subject_sum %>%
  left_join(merge_health, by = "tag")
```

#### Create Census Dataframe
```{r census data frame}
# will number classes so that all classes of list given to function
# are labelled uniquely
make_census <- function(cf_row) {
  model <- unlist(cf_row$final_model, recursive = FALSE)
  class_df <- model$pprob

  main <- datasets[[cf_row$data]]$data

  # add class_df to main
  census <- main %>%
    mutate(idno = as.numeric(idno)) %>%
    left_join(class_df, by = "idno") %>%
    distinct(idno, .keep_all = T) %>%
    select(class, idno, cohort, starts_with("prob")) %>%
    filter(!is.na(class)) %>%
    mutate(idno = as.character(idno))

  # add survival information to census
  census <- census %>%
    mutate(idno = as.numeric(idno)) %>%
    left_join(subject_sum, by = "idno") %>%
    mutate(
      age_wk_death = le_wk,
      dead_nat = dead_censor
    )

  # return census
  census
}


final_models$census <- lapply(seq_len(nrow(final_models)), function(i) {
  cf_row <- final_models[i, ]
  make_census(cf_row)
})
```

#### Create new class numbering

```{r class numbering}
reorder_class <- function(final_models, mo_names) {
  x <- final_models
  all_names <- x$model_name
  missing_names <- all_names[!(all_names %in% mo_names)]
  i <- 0
  for (name in mo_names) {
    # generate new_class
    census <- x[x$model_name == name, ]$census
    census <- unlist(census, recursive = FALSE)
    census$new_class <- as.character(census$class + i)

    # Create new prob based off of new_class
    probs_index <- grepl("^prob", names(census))
    probs <- names(census)[grepl("^prob", names(census))]
    probs <- stringr::str_split(probs, pattern = "prob", simplify = TRUE)[, 2]
    new_probs <- as.numeric(probs) + i
    new_probs <- paste0("prob", new_probs)
    names(census)[probs_index] <- new_probs

    uni_class <- length(unique(census$class))
    i <- uni_class + i
    x[x$model_name == name, ]$census[[1]] <- as.data.frame(census)
  }

  x
}

for (group in yaml_models) {
  final_models <- reorder_class(final_models, group$model_names)
}

look <- final_models[, c("model_name", "census")]
```

#### Create Table 1
```{r source create table 1, file = 'R/t1.R'}
source("R/t1.R")
```

```{r create table 1}
final_models$t1_raw <- lapply(seq_along(final_models$census), function(i) {
  cat(i, "- ")
  t1 <- t1(final_models$census[[i]], columns = final_models$dummy[[i]])
  cat("\n")

  t1
})
```

#### Perform Chi-squared Test
```{r source chi squared test, file ='R/chi_test.R'}
source("R/chi_test.R")
```

```{r chi squred test}
chi_results <- lapply(seq_along(final_models$census), function(i) {
  cat(i, "- ")
  chi <- chi_test(final_models$census[[i]], columns = final_models$dummy[[i]])
  cat("\n")

  chi
})

final_models$chi_df <- chi_results$dfs
final_models$chi_test <- chi_results$tests
```

#### Create clean t1
```{r source clean t1, file = 'R/clean_t1.R'}
source("R/clean_t1.R")
```

```{r clean t1}
final_models$clean_t1 <- lapply(seq_along(final_models$census), function(i) {
  cat(i, "- ")
  clean_t1 <- clean_t1(final_models$census[[i]],
    columns = final_models$dummy[[i]]
  )
  cat("\n")

  clean_t1
})
```

### Surival Analysis 

```{r source surv analysis, file = 'R/surv.R'}
source("R/surv.R")
```

```{r surv analysis}
final_models$census <- lapply(final_models$census, function(df) {
  df <- df %>%
    mutate(Class = new_class)
  df
})

final_models$kap_plot <- lapply(seq_len(nrow(final_models)), function(i) {
  kap_plot <- kap_plot_all(final_models$census[[i]],
    var = "Class",
    ptitle = paste(final_models$oc_name[[i]],
      final_models$age_var_name[[i]],
      "KM",
      sep = " "
    )
  )

  kap_plot
})
```

### Create Cox Models

- Model 1: $ Surv ~ Class $
- Model 2: $ Surv ~ Class + Sex + Strain$
- Model 3: $ Surv ~ Class + Sex * Strain$
```{r source model_cox, file = 'R/model_cox.R'}
source("R/model_cox.R")
```

```{r cox models}
x <- final_models
final_models$cox_models <- lapply(seq_len(nrow(x)), function(i) {
  model_cox(
    census = x$census[[i]],
    var = "Class",
    covariates = c("sex", "strain"),
    age_death = "age_wk_death",
    censor = "dead_nat"
  )
})
```

### Trajectory Fixed Effects Plotted

```{r source fixed effects, include = FALSE, file = 'R/traj_plot.R'}
source("R/traj_plot.R")
```

```{r fixed effects}
final_models$traj_pred <- lapply(seq_len(nrow(final_models)), function(i) {
  fixcov_names <- final_models$fixcov[[i]]
  fixcov <- rep(0, length(final_models$fixcov[[i]]))
  names(fixcov) <- fixcov_names
  traj_plot(
    df = final_models$dfs[[i]],
    mo = final_models$final_model[[i]],
    t1 = final_models$t1_raw[[i]],
    age_var = final_models$age_var[[i]],
    y_var = final_models$oc[[i]],
    title = paste("Predicted",
      final_models$oc_name[[i]],
      "Trajectories across",
      final_models$age_var_name[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    ),
    fixcov = fixcov
  )
  # can render this plot with ggdraw or plot_grid
  plot <- recordPlot()

  plot
})
```

### Create Observed Trajectory Plots

```{r source observed plots, include = FALSE, file = 'R/plot_obs.R'}
source("R/plot_obs.R")
```

```{r observed plots}
final_models$traj_obs <- lapply(seq_len(nrow(final_models)), function(i) {
  plot <- traj_obs(
    main = final_models$census[[i]],
    df = final_models$dfs[[i]],
    t1 = final_models$t1_raw[[i]],
    age_var = final_models$age_var[[i]],
    y_var = final_models$oc[[i]],
    title = paste("Observed",
      final_models$oc_name[[i]],
      "Trajectory with",
      final_models$age_var_name[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    )
  )

  plot
})


lapply(final_models$traj_obs, print)
```

### Create Plots for Other Metabolic Outcomes by Class

```{r source other plot, file = 'R/plot_other.R'}
source("R/plot_other.R")
```

```{r other metabolic outcomes}
for (model in names(yaml_models$models)) {
  col_name <- paste0("plot_", model)
  a <- plot_other_apply(
    final_models,
    yaml_models$models[[model]]
  )
  final_models[[col_name]] <- a
}
```

### Fit other metabolic outcomes with LME

- Model0 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model0.5 : $outcome = Age + Age^2 + Class + sex * Strain + (Age + Age^2 | Subject)$
- Model1 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model2 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Sex * Strain + (Age + Age^2 | Subject)$
- Model3 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Sex * Strain + (Age + Age^2 | Subject)$

```{r source mixed effects, file = 'R/model_other.R'}
source("R/model_other.R")
```

```{r mixed effects}
for (model in names(yaml_models$models)) {
  col_name <- paste0("model_", model)
  model_name <- yaml_models$models[[model]]
  fixcovs <- final_models[final_models$model_name == model_name, ]$fixcov[[1]]
  a <- model_other_apply(
    final_models,
    model_name,
    fixcovs = fixcovs,
    class = "new_class"
  )
  final_models[[col_name]] <- a
}
```

### Create Spaghettie Plots

```{r source spag plot, file = "R/plot_spag.R"}
source("R/plot_spag.R")
```

```{r spag plot}
final_models$spag_plot <- lapply(seq_len(nrow(final_models)), function(i) {
  plot_spag(
    data = final_models$dfs[[i]],
    x = final_models$age_var[[i]],
    y = final_models$oc[[i]],
    id = final_models$subject[[i]],
    title = paste(final_models$oc_name[[i]],
      final_models$age_var_name[[i]],
      "Spaghetti Plot",
      sep = " "
    ),
    xlab = paste(final_models$age_var_name[[i]],
      final_models$age_var_units[[i]],
      sep = " "
    ),
    ylab = paste(final_models$oc_name[[i]],
      final_models$oc_units[[i]],
      sep = " "
    )
  )
})
```

### Class Prediction

#### Class Prediction Raw Results
```{r source create prediction results, file = 'R/pred_class.R'}
source("R/pred_class.R")
```

```{r create prediction results}
# Get the names of the filtered data
prediction_data_names <- NULL
for (dataset in datasets) {
  new_prediction_data_names <- names(dataset$prediction_data$data)
  prediction_data_names <- c(prediction_data_names, new_prediction_data_names)
}
prediction_columns <- unique(prediction_data_names)

# create a list of lists, where parent list is the filter applied, and the
# daughter lists provides the filtered dataframes for the corresponding
# row of final_models. This code does not take into account the case where
# certain filters are applied to one row, but not another
filter_list <- list()
for (filter in prediction_columns) {
  row_list <- list()
  for (row_number in seq_len(nrow(final_models))) {
    data_name <- final_models[row_number, "data"]
    dataset <- datasets[[data_name]]
    prediction_data <- dataset$prediction_data
    row_list[[row_number]] <- prediction_data$data[[filter]]
  }
  filter_list[[filter]] <- row_list
}

for (filter in names(filter_list)) {
  final_models[[filter]] <- filter_list[[filter]]
}

prediction_results_names <- paste(prediction_columns, "results", sep = "_")

prediction_results <- lapply(prediction_columns, function(col) {
  print(col)
  predict_class_model_list(
    nested_newdata_list = final_models[[col]],
    model_list = final_models$final_model,
    vars_list = final_models$data_age_vars,
    center = TRUE,
    scale = FALSE,
    ref_data_list = final_models$dfs,
    names = final_models$model_name
  )
})

final_models[, prediction_results_names] <- prediction_results
```

### Sensititviy, Specificity, PPV, NPV, Tables

|  | Gold Yes | Gold No | total |
| --- | --- | --- | ---- |
| pred Yes | $TP = tp$ | $FP = pred\_class - tp$ | pred_class |
| pred No | $FN =og\_class - tp$ | $TN = n - (og\_class + pred\_class - tp)$ | n - pred_class | 
| total | og_class | n - og_class | n

- Sensitivity = 
$\frac{TP}{TP + FN} = 
\frac{tp}{tp + og\_class - tp} =
\frac{tp}{og\_class}$
- Specificity =
$\frac{TN}{FP + TN} = 
\frac{n - (og\_class + pred\_class - tp)}{pred\_class - tp + n - og\_class - pred\_class + tp)} = 
\frac{n - (og\_class + pred\_class - tp)}{n - og\_class}$
- Positive Predictive Value = 
$\frac{TP}{TP + FP} = 
\frac{tp}{tp + pred\_class - tp} = 
\frac{tp}{pred\_classs}$
- Negative Predictive Value = 
$\frac{TN}{FN + TN} = 
\frac{n - (og\_class + pred\_class -tp)}{og\_class - tp + n - og\_class - pred\_class + tp) } =
\frac{n - (og\_class + pred\_class - tp)}{n - pred\_class}$
- Accuracy = 
$\frac{TN + TP}{TN + TP + FN + FP} = 
\frac{2tp + n - og\_class - pred\_class}{n}$


```{r source prediciton tables, file = 'R/pred_table.R'}
source("R/pred_table.R")
```

```{r prediciton tables}
prediction_results_table_names <- paste(prediction_results_names,
  "table",
  sep = "_"
)

prediction_results_tables <- lapply(
  prediction_results_names,
  function(subset) {
    create_pred_table_subset(subset,
      subject = "idno"
    )
  }
)

final_models[, prediction_results_table_names] <- prediction_results_tables
```

#### Dataframe for each subset type with prediction statistics

```{r source create prediciton dataframe for plotting, file = 'R/combine_pred_df.R'}
source("R/combine_pred_df.R")
```

```{r create prediciton dataframe for plotting}
prediction_results_df_names <- paste(prediction_results_names,
  "df",
  sep = "_"
)
pred_results <- lapply(
  prediction_results_table_names,
  combine_pred_df_subset
)

names(pred_results) <- prediction_results_df_names
final_models[, prediction_results_df_names] <- pred_results
```

#### Prepare prediciton dataframe for plotting

```{r source prepare prediction dataframe, file = 'R/rename_pred.R'}
source("R/rename_pred.R")
```

```{r prepare prediction dataframe}
final_models$filter_interval_data_results_df <- lapply(
  final_models$filter_interval_data_results_df,
  name_interval
)

final_models$filter_cumulative_data_results_df <- lapply(
  final_models$filter_cumulative_data_results_df,
  name_threshold
)

final_models$resampled_data_results_df <- lapply(
  final_models$resampled_data_results_df,
  name_sample
)

final_models$filter_window_data_results_df <- lapply(
  final_models$filter_window_data_results_df,
  name_threshold
)
```

#### Plot Prediction Results

```{r source plot pred_results, file = 'R/plot_pred.R'}
source("R/plot_pred.R")
```

```{r plot pred_results, fig.height=10, fig.width=5}
final_models$interval_plots <- lapply(
  seq_along(final_models$filter_interval_data_results_df),
  function(i) {
    plot_interval(final_models$filter_interval_data_results_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "Interval",
        sep = " "
      )
    )
  }
)

final_models$interval_plots

final_models$cumulative_plots <- lapply(
  seq_along(final_models$filter_cumulative_data_results_df),
  function(i) {
    plot_threshold(final_models$filter_cumulative_data_results_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "threshold",
        sep = " "
      )
    )
  }
)

final_models$cumulative_plots

final_models$resampled_plots <- lapply(
  seq_along(final_models$resampled_data_results_df),
  function(i) {
    plot_sample(final_models$resampled_data_results_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "sample",
        sep = " "
      )
    )
  }
)

final_models$resampled_plots

final_models$window_plots <- lapply(
  seq_along(final_models$filter_window_data_results_df),
  function(i) {
    plot_window(final_models$filter_window_data_results_df[[i]],
      cols = c("Sensitivity", "Specificity", "PPV", "NPV"),
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "window",
        sep = " "
      )
    )
  }
)

final_models$window_plots
```

#### Plot Accuracy

```{r plot accuracy}
final_models$interval_accuracy_plots <- lapply(
  seq_along(final_models$filter_interval_data_results_df),
  function(i) {
    plot_interval(final_models$filter_interval_data_results_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "Interval",
        sep = " "
      )
    )
  }
)

final_models$interval_accuracy_plots

final_models$cumulative_accuracy_plots <- lapply(
  seq_along(final_models$filter_cumulative_data_results_df),
  function(i) {
    plot_threshold(final_models$filter_cumulative_data_results_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "threshold",
        sep = " "
      )
    )
  }
)

final_models$cumulative_accuracy_plots

final_models$resampled_accuracy_plots <- lapply(
  seq_along(final_models$resampled_data_results_df),
  function(i) {
    plot_sample(final_models$resampled_data_results_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "sample",
        sep = " "
      )
    )
  }
)

final_models$resampled_accuracy_plots

final_models$window_accuracy_plots <- lapply(
  seq_along(final_models$filter_window_data_results_df),
  function(i) {
    plot_window(final_models$filter_window_data_results_df[[i]],
      cols = "Accuracy",
      title = paste(final_models$data_name[[i]],
        final_models$age_var_name[[i]],
        sep = " "
      ),
      y = "Value",
      x = paste(final_models$age_var_name[[i]],
        final_models$age_var_units[[i]],
        "window",
        sep = " "
      )
    )
  }
)

final_models$window_accuracy_plots
```

### Compute Accuracy Again as Double Check and Bootstrap

```{r source boot accuracy, file = 'R/boot_accuracy.R'}
source("R/boot_accuracy.R")
```

```{r compute accuracy}
final_models$interval_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_interval_data_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$cumulative_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_cumulative_data_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$window_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$filter_window_data_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)

final_models$resampled_accuracy <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    boot_accuracy_list(
      pred_df_list = final_models$resampled_data_results[[i]],
      og_df = final_models$census[[i]],
      subject = final_models$subject[[i]]
    )
  }
)
```

#### Combine bootstraped accuracy for each simulation type

```{r source combine accuracy, file = 'R/combine_boot_accuracy.R'}
source("R/combine_boot_accuracy.R")
```

```{r combine accuracy}
x <- final_models
x$interval_accuracy_df <- combine_boot_accuracy_list(x$interval_accuracy)
x$cumulative_accuracy_df <- combine_boot_accuracy_list(x$cumulative_accuracy)
x$resampled_accuracy_df <- combine_boot_accuracy_list(x$resampled_accuracy)
x$window_accuracy_df <- combine_boot_accuracy_list(x$window_accuracy)
final_models <- x
```

#### Prepare Accuracy Dataframe for pLotting

```{r prepare accuracy for plotting}
final_models$interval_accuracy_df <- lapply(
  final_models$interval_accuracy_df,
  name_interval
)


final_models$cumulative_accuracy_df <- lapply(
  final_models$cumulative_accuracy_df,
  name_threshold
)

final_models$resampled_accuracy_df <- lapply(
  final_models$resampled_accuracy_df,
  name_sample
)

final_models$window_accuracy_df <- lapply(
  final_models$window_accuracy_df,
  name_new_interval
)
```

#### Extract Mean and CI from Bootstrapped Accuracy

```{r source extraction, file = 'R/extract_accuracy_ci.R'}
source("R/extract_accuracy_ci.R")
```

```{r mean and ci}
final_models$interval_accuracy_df <- lapply(
  final_models$interval_accuracy_df,
  extract_accuracy_ci
)

final_models$cumulative_accuracy_df <- lapply(
  final_models$cumulative_accuracy_df,
  extract_accuracy_ci
)

final_models$resampled_accuracy_df <- lapply(
  final_models$resampled_accuracy_df,
  extract_accuracy_ci
)

final_models$window_accuracy_df <- lapply(
  final_models$window_accuracy_df,
  extract_accuracy_ci
)
```

#### Plot Accuracy Again With confidence Intervals

```{r source boot plot, file = 'R/plot_boot_accuracy.R'}
source("R/plot_boot_accuracy.R")
```

```{r plot accuracy with bootstrapped confidence intervals}
final_models$interval_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_interval(final_models$interval_accuracy_df[[i]],
      xlab = "Life Stage",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$interval_boot_accuracy_plot

final_models$window_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_window(final_models$window_accuracy_df[[i]],
      xlab = "Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$window_boot_accuracy_plot

final_models$cumulative_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_threshold(final_models$cumulative_accuracy_df[[i]],
      xlab = "Upper Bound Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Cumulative Classification Accuracy using",
        final_models$oc_name[[i]]
      ),
      legend_title = "Lower Bound"
    )
  }
)

final_models$cumulative_boot_accuracy_plot

final_models$resampled_boot_accuracy_plot <- lapply(
  seq_len(nrow(final_models)),
  function(i) {
    plot_boot_accuracy_sample(final_models$resampled_accuracy_df[[i]],
      xlab = "Lower Bound Age (weeks)",
      ylab = "Accuracy",
      title = paste(
        "Classification Accuracy using",
        final_models$oc_name[[i]]
      )
    )
  }
)

final_models$resampled_boot_accuracy_plot
```

## Combined Metabolic Class

### Create Combined Cox 

```{r source combined_class_cox, file = 'R/combined_class_cox.R'}
source("R/combined_class_cox.R")
```

```{r combined survival}
# We can parameterize this later
covariates <- c("sex", "strain")
age_death <- "age_wk_death"
censor <- "dead_nat"


model_name_list <- unlist(yaml_models, recursive = FALSE)

combined_cox_models <- lapply(
  model_name_list,
  cox_combine_class,
  final_models,
  datasets,
  covariates,
  age_death,
  censor
)

names(combined_cox_models) <- lapply(model_name_list, paste, collapse = "-")

for (name in names(combined_cox_models)) {
  name_vector <- stringr::str_split(name, pattern = "-")[[1]]
  model_index <- final_models$model_name %in% name_vector
  model_index <- match(name_vector, final_models$model_name)
  for (i in model_index) {
    final_models[i, "cox_combined_models"] <- list(combined_cox_models[name])
  }
}
```

### Determine get Hazard Predictions from test set

#### Predict Class Assignement for Test Dataframe

```{r source create prediction results, file = 'R/pred_class.R'}
source("R/pred_class.R")
```

```{r create prediction results}
# Get the names of the filtered data
prediction_data_names <- NULL
for (dataset in datasets) {
  new_prediction_data_names <- names(dataset$prediction_test_data$data)
  prediction_data_names <- c(prediction_data_names, new_prediction_data_names)
}
prediction_columns <- unique(prediction_data_names)

# create a list of lists, where parent list is the filter applied, and the
# daughter lists provides the filtered dataframes for the corresponding
# row of final_models. This code does not take into account the case where
# certain filters are applied to one row, but not another
filter_list <- list()
for (filter in prediction_columns) {
  row_list <- list()
  for (row_number in seq_len(nrow(final_models))) {
    data_name <- final_models[row_number, "data"]
    dataset <- datasets[[data_name]]
    prediction_data <- dataset$prediction_test_data
    row_list[[row_number]] <- prediction_data$data[[filter]]
  }
  filter_list[[filter]] <- row_list
}

for (filter in names(filter_list)) {
  # we need a new name for test_data prediction data so as to not overlap
  # with the nontest data prediction data
  filter_name <- paste0(filter, "_test_data")
  final_models[[filter_name]] <- filter_list[[filter]]
}
# add _test_data to prediction columns names
prediction_columns <- paste0(prediction_columns, "_test_data")

prediction_results_names <- paste(prediction_columns, "results", sep = "_")

prediction_results <- lapply(prediction_columns, function(col) {
  print(col)
  predict_class_model_list(
    nested_newdata_list = final_models[[col]],
    model_list = final_models$final_model,
    vars_list = final_models$data_age_vars,
    center = TRUE,
    scale = FALSE,
    ref_data_list = final_models$dfs,
    names = final_models$model_name
  )
})

final_models[, prediction_results_names] <- prediction_results
```

#### Create new class numbering

```{r class numbering}
increment_class <- function(filtered_df, i = 0) {
  filtered_df$new_class <- as.character(filtered_df$class + i)

  # Create new prob based off of new_class
  probs_index <- grepl("^prob", names(filtered_df))
  probs <- names(filtered_df)[grepl("^prob", names(filtered_df))]
  probs <- stringr::str_split(probs,
    pattern = "prob",
    simplify = TRUE
  )
  probs <- probs[, 2]
  new_probs <- as.numeric(probs) + i
  new_probs <- paste0("prob", new_probs)
  names(filtered_df)[probs_index] <- new_probs

  filtered_df
}

reorder_prediction_class <- function(final_models,
                                     model_names,
                                     filtered_df_name) {
  i <- 0
  for (model_name in model_names) {
    index <- final_models$model_name == model_name
    row <- final_models[index, ]
    filtered_dfs <- row[, filtered_df_name]
    filtered_dfs <- filtered_dfs[[1]]
    if (is.list(filtered_dfs)) {
      filtered_dfs_inced <- lapply(
        filtered_dfs,
        increment_class,
        i
      )
    } else {
      filtered_dfs_inced <- NA
    }

    column <- final_models[, filtered_df_name]
    column[[which(index)]] <- filtered_dfs_inced
    final_models[[filtered_df_name]] <- column
    i <- as.numeric(row$ng) + i
  }
  final_models
}

for (group in yaml_models) {
  for (filtered_df_name in prediction_results_names) {
    final_models <- reorder_prediction_class(
      final_models,
      group$model_names,
      filtered_df_name
    )
  }
}
```
#### Merge test_data prediction dataframes based upon input yaml

```{r merge prediction dataframes}
merge_prediction_data <- function(data_list_x,
                                  data_list_y,
                                  id_x,
                                  id_y,
                                  suffixes) {
  # remake suffixes
  suffixes <- paste0("_", suffixes)

  # check that data_list_x and data_list_y are the same length
  l_x <- length(data_list_x)
  l_y <- length(data_list_y)
  if (l_x != l_y) stop("Data lists are not of the same length")

  merged_datasets <- list()
  for (i in seq_len(l_x)) {
    merged_dataset <- base::merge(
      x = data_list_x[[i]],
      y = data_list_y[[i]],
      by.x = id_x,
      by.y = id_y,
      suffixes = suffixes
    )
    merged_datasets <- c(merged_datasets, list(merged_dataset))
  }

  merged_datasets
}

docall_merged_prediction_data <- function(
    data_list_list,
    id_list,
    suffixes_list,
    names) {
  merged_census <- data_list_list[[1]]
  for (i in seq_len(length(data_list_list) - 1)) {
    merged_census <- merge_prediction_data(
      data_list_x = merged_census,
      data_list_y = data_list_list[[i + 1]],
      id_x = id_list[[i]],
      id_y = id_list[[i + 1]],
      suffixes = suffixes_list[c(i, i + 1)]
    )
  }
  names(merged_census) <- names
  merged_census
}

merge_prediction_data_group <- function(
    final_models,
    model_names,
    filtered_df_name) {
  index <- final_models$model_name %in% model_names
  desired_dfs <- final_models[index, filtered_df_name]
  desired_ids <- final_models[index, "subject"]
  desired_suffixes <- final_models[index, "oc"]

  if (all(is.na(desired_dfs))) {
    merged_data <- NA
  } else {
    merged_data <- docall_merged_prediction_data(
      data_list_list = desired_dfs,
      id_list = desired_ids,
      suffixes_list = desired_suffixes,
      names = names(desired_dfs[[1]])
    )
  }

  # create new column to store data
  column <- final_models[, filtered_df_name]
  for (row in which(index)) {
    column[[row]] <- merged_data
  }
  new_column_name <- paste0(filtered_df_name, "_merged")
  final_models[[new_column_name]] <- column
  final_models
}

for (group in yaml_models) {
  for (filtered_df_name in prediction_results_names) {
    final_models <- merge_prediction_data_group(
      final_models,
      group$model_names,
      filtered_df_name
    )
  }
}
```
#### Create test census dataframe

```{r create test census}
# TODO: Refactor merge census, and code such that it is not dependent on dplyr
# TODO: Remove nonsense entries from census
# TODO: Refactor all the datasets such that all the outliers ect. would be
# part of their own run
merge_census <- function(data, census, id) {
  if (is.data.frame(data)) {
    data <- merge(
      x = data,
      y = census,
      by = id
    )

    colnames(data)[colnames(data) == "le_wk"] <- "age_wk_death"
    colnames(data)[colnames(data) == "dead_censor"] <- "dead_nat"
  } else {
    data <- NA
    warning("data is not a data.frame, returning NA")
  }

  data
}

merge_census_apply <- function(data_list, census, id) {
  if (all(is.na(data_list))) {
    data_list <- NA
    warning("data_list is all NA, returning NA")
  } else {
    data_list <- lapply(data_list, merge_census, census = census, id = id)
  }
  data_list
}

filtered_df_name <- merged_prediction_result_names[[1]]

merge_census_apply_column <- function(final_models, filtered_df_name, census) {
  data_lol <- final_models[, filtered_df_name]
  ids <- final_models[, "subject"]
  data_lol <- lapply(data_lol, merge_census_apply, census = census, id = ids)

  final_models[[filtered_df_name]] <- data_lol
  final_models
}

for (filtered_df_name in merged_prediction_result_names) {
  final_models <- merge_census_apply_column(
    final_models = final_models,
    filtered_df_name = filtered_df_name,
    census = subject_sum
  )
}
```

#### We need to predict hazard ratios for our test sets

```{r test mortaility prediction using train test}
# delete
test_data <- 
```

### Output CSV with idno's and class

```{r idnos and class}
# we want to output class, idno, tag to a csv so that collaborators can easily
# work with the data

setwd(file.path("output", params$model_dir))
dir.create("csv")
setwd("csv")

for (model in seq_len(nrow(final_models))) {
  census <- final_models$census[[model]]
  model_name <- final_models$model_name[[model]]
  census <- census[, c("class", "idno", "tag", "sex", "strain", "cohort.y")]
  file_name <- paste0(model_name, ".csv")
  write.csv(census, file = file_name)
}
```

### Save R Objects

#### Prune Final Models Before Saving 

```{r prune}
final_models <- final_models %>%
  select(
    -c(
      interval_accuracy_plots,
      threshold_accuracy_plots,
      sample_accuracy_plots
    ),
    -c(
      ends_with("results_table"),
      ends_with("accuracy")
    ),
    -c(final_model)
  )

final_models <- final_models %>%
  select(-c(
    ends_with("_df"),
    ends_with("_results")
  ))

final_models <- final_models %>%
  select(-interval, -subset, -threshold, -sample)

final_models <- final_models %>%
  filter(model_name == "bw_main|bw_age_wk" |
    model_name == "fat_main|fat_age_wk" |
    model_name == "gluc_main|gluc_age_wk")
```

#### Save 

```{r save R objects}
setwd(file.path("output", params$model_dir))
dir.create("data")
setwd("data")

save(cf, file = "cf.RDATA")
save(final_models, file = "final_models.RDATA")
save(config, file = "config.RDATA")
save(df_list, file = "df_list.RDATA")
save(dfs_prediction, file = "dfs_prediction.RDATA")
```
