---
title: "SLAM Metabolic Class COD and healthcard"
author: "William Mueller, Joshua Preston"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
--- 

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
  library(consoler)
  library(SLAM)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(broom)
  library(scales)
  library(pheatmap)
  library(RColorBrewer)
  library(survminer)
  library(survival)
  library(gridtext)
  library(gridExtra)
  library(grid)
  library(ggplot2)
  library(svglite)
  set.seed(365)

  # setting global knitr options for r markdown output
  knitr::opts_chunk$set(autodep = TRUE)
  knitr::opts_chunk$set(warning = FALSE)
  knitr::opts_chunk$set(message = FALSE)
  knitr::opts_chunk$set(echo = FALSE)
  knitr::opts_chunk$set(cache = FALSE)

  # Setting global class and bin colors for plots
    class_colors <- c(
      "1" = "#f2605d", # red
      "2" = "#36b22b", # green
      "3" = "#5087fe" # blue
    )
    bin_colors <- c(
      "0" = "#5be996", # green
      "1" = "#f2d14c", # yellow
      "2" = "#efa52e", # darker yellow-orange
      "3" = "#f06151", # red
      "4" = "#982215", # dark red
      "5+" = "#350202" # dark dark red
    )
    bin_colors_per_yr <- c(
      "[0]" = "#5be996", # green
      "(0–0.5)" = "#f2d14c", # yellow
      "[0.5–1)" = "#efa52e", # darker yellow-orange
      "[1–2)" = "#f06151", # red
      "[2–4)" = "#982215", # dark red
      "≥4" = "#350202" # dark dark red
    )
```
# Load in input data
```{r read in censuses}
  census_path <- "/Users/jdp2019/Library/CloudStorage/Box-Box/metabolic-class-ITP/98_healthcard_cod/data/census.csv"
  class_census_path <- "/Users/jdp2019/Library/CloudStorage/Box-Box/metabolic-class-ITP/98_healthcard_cod/data/20250426_all_cohort/complete_census.csv"
  healthcard_path <- "/Users/jdp2019/Library/CloudStorage/Box-Box/metabolic-class-ITP/98_healthcard_cod/data/SLAM Healthcard reconciled.xlsx"

  class_census <- read.csv(class_census_path)
  census <- read.csv(census_path)
  healthcard_ii <- as.data.frame(readxl::read_xlsx(healthcard_path, sheet = "Healthcard"))
  census <- consoler::rename(census, c(animal_id = "Animal_ID", tag_history = "taghistory"))
```

## Clean Healthcard
```{r clean healthcard}
  #+ Clean Healthcard
    new_names <- c(
      tag = "Tag", dod = "Date of Death",
      created_date = "Created Date",
      recovery_date = "Recovery Date",
      condition = "Condition",
      idno_rec = "idno"
    )
    healthcard_i <- consoler::rename(healthcard_ii, new_names) %>% # trim whitespace
      mutate(condition = str_trim(condition))
    condition_mappings <- tibble::tribble(
      ~canonical_name, ~aliases,
      "fighting", c("Fighting", "fighting"),
      "dermatitis", c("Dermatitis", "Skin Reaction", "Red Skin"),
      "alopecia", c("Alopecia/Barbering"),
      "trauma", c("traumatic injury", "Tramatic Injury", "Traumatic Injury"),
      "dehydration", c("Dehydrated", "Dehydration"),
      "seizure", c("Seizure"),
      "edema", c("Swelling", "Swollen Feet", "edema", "Edema", "Swollen Hindfoot"),
      "flood", c("Flooded Cage", "cage flood", "Cage Flood"),
      "mass", c("Mass", "mass", "Mass (size in mm)", "Mass "),
      "prolapse", c("Prolapse", "prolapse", "Prolapse (rectal/genital)", "Prolapsed Penis", "Rectal Prolapse", "Prolapse ", "Penile Prolapse", "prolpase"),
      "lesion", c("Lesion", "lesion", "Lesions"),
      "weight_loss", c("Weight Loss", "weight loss", "Losing Weight"),
      "bleed", c("Bleeding", "Rectal Bleeding", "Nose Bleeding", "bleeding", "Vaginal Bleeding"),
      "irritation", c("Irritation"),
      "lethargy", c("Lethargy", "lethargy", "Lethargic", "Slow Moving"),
      "hunch", c("Hunched Posture", "Thin/Hunched"),
      "dyspnea", c("Dyspnea", "dyspnea", "Dyspnea ", "Increased RR", "Dyspnea (Labored Breathing)"),
      "distented_abdomen", c("Distended Abdomen"),
      "abscess", c("Abscess"),
      "paralysis", c("Paralysis", "paralysis"),
      "eye_problem", c("Eye Problem", "eye problem", "Eye Issue"),
      "ear_problem", c("Ear Problem", "Ear Problems"),
      "head_tilt", c("Head Tilt", "head tilt"),
      "discharge", c("Discharge", "Vaginal Discharge"),
      "moribund", c("Moribund", "moribund", "Moribund ", "Moribund (near death)"),
      "urinary", c("Urinary Distress", "Urine Staining", "Excessive Urination", "Unable to Urinate", "Discolored Urine", "Cloudy/Tinted Urine"),
      "gait", c("Abnormal Gait", "Unsteady Gait", "Hindlimb Weakness", "Favoring Hindlimb"),
      "malocclusion", c("Malocclusion"),
      "hernia", c("Hernia"),
      "swollen_testicle", c("Swollen Testicle", "Enlarged testicle"),
      "thin", c("Thin"),
      "skin_tag", c("Skin Tag"),
      "hypoglycemic", c("Low Glucose", "hypoglycemic"),
      "cyanotic", c("Cyanotic"),
      "organomegaly", c("Enlarged Organs"),
      "permanent_defect", c("Permanent Defect"),
      "hydrocehpalus", c("Hydrocephalus"),
      "diarrhea", c("Diarrhea"),
      "nutritional_support", c("Nutritional Support"),
      "low_temp", c("Low Temperature", "Low body temperature"),
      "blood_in_cage", c("Blood in Cage"),
      "red_paws", c("Red Paws"),
      "behavior", c("odd behavior"),
      "other", c("Other"),
      "enlarged_bladder", c("enlarged bladder"),
      "tail_issue", c("Tail Issue")
    )
  #+ Rename per above
    for (i in seq_len(nrow(condition_mappings))) {
      row <- condition_mappings[i, ]
      aliases <- unlist(row$aliases)
      canonical_name <- row$canonical_name
      healthcard_i$condition[healthcard_i$condition %in% aliases] <- canonical_name
    }
  #+ Verify the remapping
    mapped_conditions <- unlist(condition_mappings$aliases, use.names = FALSE) %>% trimws()
    unique_conditions <- unique(healthcard_i$condition) %>% trimws()
    unmapped <- setdiff(unique_conditions, condition_mappings$canonical_name)
    if (length(unmapped) > 0) {
      message("⚠️ Unmapped conditions found:")
      print(unmapped)
    } else {
      message("✅ All conditions are mapped correctly.")
    }
  #+ Convert to dummy columns
    healthcard <- fastDummies::dummy_cols(healthcard_i, "condition",
      omit_colname_prefix = TRUE
    ) %>%
    filter(idno_rec != 154) # Filter out this mouse because in the HC reconcile this was report was created by accident on wrong mouse
    #+ Merge with census
    healthcard_final <- as_tibble(merge(census, healthcard, by = "tag", all.y = TRUE)) %>%
      select(idno, idno_rec, everything(), -c(Comments, name, cage, eartag, "Created By", "Entered By"))
```

## Process healthcard to prep for graphing- EOL
```{r process healthcard}
  # Clean up and pare HC
  HC_relevant <- healthcard_final %>%
    select(idno_rec, created_date, recovery_date, abscess:weight_loss) %>%
    dplyr::rename(idno = idno_rec) %>%
    mutate(recovery_date = as.Date(recovery_date)) %>%
    mutate(created_date = as.Date(as.numeric(created_date), origin = "1899-12-30"))

  # Get a count of each mouses number of HCs
  hc_counts <- HC_relevant %>%
    count(idno, name = "num_HC")
    
  # Now assign the counts and HCs to the full class census
  classes_i <- as_tibble(class_census) %>%
    mutate(any_HC = if_else(idno %in% HC_relevant$idno, 1, 0)) %>%
    left_join(HC_relevant, by = "idno") %>%
    left_join(hc_counts, by = "idno") %>%
    mutate(num_HC = replace_na(num_HC, 0)) %>%
    arrange(desc(num_HC)) %>%
    mutate(
      tod = as.Date(tod), # ensure `tod` is Date class
      condition_length = as.integer(coalesce(recovery_date, tod) - created_date),
      condition_to_death = as.integer(tod - created_date)
    )

  # Get a count of each condition, and pull those with only 5 issues or less
  condition_counts <- classes_i %>%
    summarise(across(abscess:weight_loss, ~ sum(.x, na.rm = TRUE))) %>%
    pivot_longer(cols = everything(), names_to = "condition", values_to = "count") %>%
    arrange(count) %>%
    filter(count <= 5) %>%
    pull(condition)

  # Now remove those conditions from the dataset
  classes <- classes_i %>%
    select(condition_to_death, condition_length, any_HC, num_HC, idno, new_class_bw, cod, tod, created_date:weight_loss, -c(all_of(condition_counts))) %>%
    arrange(desc(condition_to_death))

  # Create versions which is only EOL (within 4 months of death) healthcards
  valid_hc_rows_EOL <- classes %>%
    filter(condition_to_death > 122)
    
  hc_long_EOL <- valid_hc_rows_EOL %>%
    pivot_longer(cols = abscess:weight_loss, names_to = "condition", values_to = "has_condition") %>%
    filter(has_condition == 1) %>% 
    distinct(idno, condition, new_class_bw) # Only counts condition once
  animals_with_conditions_EOL <- hc_long_EOL %>%
    distinct(idno)

  all_animals_EOL <- classes %>%
    distinct(idno, new_class_bw)

  no_eol_hc <- all_animals_EOL %>% # Anti-join to get animals with no qualifying conditions
    anti_join(animals_with_conditions_EOL, by = "idno") %>%
    mutate(condition = "no_EOL_HCs")

  EOL_hc_final <- bind_rows(hc_long_EOL, no_eol_hc) # Bind to condition table
```

## Calculate chi-square or fisher's p-values for EOL
```{r run chi square and fisher's}
  # Chi square and fisher's for EOL conditions
  EOL_results <- EOL_hc_final %>%
  mutate(present = 1) %>%
  complete(idno, condition, new_class_bw, fill = list(present = 0)) %>% # ensure all combinations
  pivot_wider(names_from = condition, values_from = present, values_fill = 0) %>%
  pivot_longer(cols = -c(idno, new_class_bw), names_to = "condition", values_to = "present") %>%
  group_by(condition) %>%
  nest() %>%
  mutate(
    test = map(data, ~ {
      tab <- table(.x$present, .x$new_class_bw)
      if (any(tab < 5)) {
        fisher.test(tab)
      } else {
        chisq.test(tab)
      }
    }),
    tidy = map(test, broom::tidy)
  ) %>%
  unnest(tidy) %>%
  select(condition, method, p.value, statistic) %>%
  arrange(p.value)
```

## Calculate prevalence of conditions for graphing for EOL, join p-values
```{r calculate prevalence, assign p-values}
  # Calculate prevalence for EOL, join p-values from above
  class_totals_EOL <- EOL_hc_final %>%
    distinct(idno, new_class_bw) %>%
    count(new_class_bw, name = "denominator")

  condition_counts_EOL <- EOL_hc_final %>%
    distinct(idno, new_class_bw, condition) %>%
    count(condition, new_class_bw, name = "numerator")

  plot_data_EOL_i <- condition_counts_EOL %>%
    left_join(class_totals_EOL, by = "new_class_bw") %>%
    mutate(prevalence = (numerator / denominator) * 100) %>%
    left_join(EOL_results %>% select(condition, p.value), by = "condition") %>%
    mutate(sig_label = case_when(
      p.value <= 0.001 ~ "***",
      p.value <= 0.01 ~ "**",
      p.value <= 0.05 ~ "*",
      TRUE ~ ""
    ))

  # Calculate prevalence for EOL, join p-values from above
  class_totals_LL <- LL_hc_final %>%
    distinct(idno, new_class_bw) %>%
    count(new_class_bw, name = "denominator")
  condition_counts_LL <- LL_hc_final %>%
    distinct(idno, new_class_bw, condition) %>%
    count(condition, new_class_bw, name = "numerator")
  plot_data_LL_i <- condition_counts_LL %>%
    left_join(class_totals_LL, by = "new_class_bw") %>%
    mutate(prevalence = (numerator / denominator) * 100) %>%
    left_join(LL_results %>% select(condition, p.value), by = "condition") %>%
    mutate(sig_label = case_when(
      p.value <= 0.001 ~ "***",
      p.value <= 0.01 ~ "**",
      p.value <= 0.05 ~ "*",
      TRUE ~ ""
    ))
```

## Clean the names of the conditions in both for plotting
```{r clean the conditions, assign p-values}
  # Create a vector to clean all names in the dataset
  conditions_subset <- classes %>% # Find the unique column names for conditions
    select(abscess:weight_loss) %>%
    colnames()
  condition_renames <- c(
    abscess = "Abscess",
    hypoglycemic = "Hypoglycemic",
    bleed = "Bleeding",
    dehydration = "Dehydration",
    dermatitis = "Dermatitis",
    distented_abdomen = "Distended Abdomen",
    dyspnea = "Dyspnea",
    ear_problem = "Ear Problem",
    edema = "Edema",
    eye_problem = "Eye Problem",
    fighting = "Fighting",
    flood = "Cage Flooding",
    gait = "Gait Abnormality",
    head_tilt = "Head Tilt",
    hunch = "Hunched Posture",
    lesion = "Lesion",
    lethargy = "Lethargy",
    low_glucose = "Low Glucose",
    low_temp = "Low Temperature",
    malocclusion = "Malocclusion",
    mass = "Mass",
    moribund = "Moribund",
    nutritional_support = "Nutritional Support",
    paralysis = "Paralysis",
    permanent_defect = "Permanent Defect",
    prolapse = "Prolapse",
    seizure = "Seizure",
    swollen_testicle = "Swollen Testicle",
    trauma = "Trauma",
    urinary = "Urinary Problem",
    weight_loss = "Weight Loss",
    no_EOL_HCs = "No EOL Healthcards",
    no_lifelong_HCs = "No Lifelong Healthcards"
  )

  # Now rename conditions in both datasets according to the above
  plot_data_EOL <- plot_data_EOL_i %>%
    mutate(condition = recode(condition, !!!condition_renames))
  unique(plot_data_EOL$condition)
  ```

## Divide EOL datasets into significant and non-significant, plot
```{r divide into sig and NS, assign p-values}
  # Divide into sig and NS
  EOL_sig <- plot_data_EOL %>%
    filter(p.value < 0.05) %>%
    filter(condition != "No EOL Healthcards") %>%
    bind_rows(tibble(
      condition = "Lesion",
      new_class_bw = 1,
      numerator = 0,
      denominator = 124, # match class 1’s denominator
      prevalence = 0,
      p.value = 5.38e-6, # use same p-value as others for Lesion
      sig_label = "***"
    )) %>%
      arrange(condition, new_class_bw)
  EOL_NS <- plot_data_EOL %>%
    filter(is.na(p.value) | p.value >= 0.05)
  no_EOL_prevalence <- plot_data_EOL %>%
    filter(condition == "No EOL Healthcards") %>%
    mutate(label = paste0("Class ", new_class_bw, ": ", round(prevalence, 1), "%")) %>%
    pull(label) %>%
    paste(collapse = "\n")
  # Plot EOL significant
  EOL_sig_plot_unified <- ggplot(EOL_sig, aes(x = condition, y = prevalence, fill = factor(new_class_bw))) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
    # geom_text(aes(label = sig_label, y = prevalence * 1.02),
    #   position = position_dodge(width = 0.8),
    #   vjust = 0, size = 4
    # ) +
    scale_y_continuous(
      labels = scales::percent_format(scale = 1),
      expand = expansion(mult = c(0, 0.15))
    ) +
    scale_fill_manual(values = class_colors, name = "Class") +
    labs(
      x = "Condition",
      y = "Prevalence (%)",
      fill = "Class",
      title = "Health Issues Differing by Class in Final 4 Months of Life"
    ) +
    theme_minimal(base_size = 16) +
    theme(
      axis.text.x = element_text(face = "bold",angle = 45, hjust = 1, size = 15),
      plot.title = element_text(face = "bold", size = 18),
      axis.title = element_text(face = "bold", size = 16),
      legend.title = element_text(face = "bold", size = 16)
    ) + 
  annotate(
    "text",
    x = Inf, y = Inf,
    hjust = 1, vjust = 1.5,
    label = paste("No EOL HCs prevalence:\n", no_EOL_prevalence),
    size = 4.5, fontface = "bold"
  )
  # Now count EOL conditions per mouse, bin, count mice per bin
  eol_bin_counts <- EOL_hc_final %>%
    mutate(n_conditions = if_else(condition == "no_EOL_HCs", 0L, 1L)) %>%
    group_by(idno, new_class_bw) %>%
    summarise(n_conditions = sum(n_conditions), .groups = "drop") %>%
    mutate(
      condition_bin = case_when(
        n_conditions == 0 ~ "0",
        n_conditions == 1 ~ "1",
        n_conditions == 2 ~ "2",
        n_conditions == 3 ~ "3",
        n_conditions == 4 ~ "4",
        n_conditions >= 5 ~ "5+"
      ),
      condition_bin = factor(condition_bin, levels = c("0", "1", "2", "3", "4", "5+"))
    ) %>%
    count(new_class_bw, condition_bin, name = "count")
  eol_bin_props <- eol_bin_counts %>%
    group_by(new_class_bw) %>%
    mutate(prop = count / sum(count)) %>%
    ungroup()
  # Plot the bins
  ggplot(eol_bin_props, aes(x = factor(new_class_bw), y = prop, fill = condition_bin)) +
    geom_bar(stat = "identity", position = "stack") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_fill_manual(values = bin_colors, name = "HCs at EOL") +
    labs(
      title = "Proportion of Mice in Each Class by EOL Condition Bin",
      x = "Class",
      y = "Proportion of Mice"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.title = element_text(face = "bold", size = 14),
      axis.text = element_text(size = 12),
      legend.title = element_text(face = "bold"),
      legend.text = element_text(size = 12)
    )
  ```


## Examine lifelong healthcards accounting for incidence (non-condition specific)
```{r Incidence study}
  #* Now, create a lifelong healthcard
    HC_relevant_LL <- classes_i %>%
      select(idno,le_wk,new_class_bw,created_date,recovery_date, tod,abscess:condition_to_death,-na) 
    total_HCs_LL <- HC_relevant_LL %>%
      rowwise() %>%
      mutate(total_conditions = sum(c_across(abscess:weight_loss), na.rm = TRUE)) %>%
      ungroup() %>%
      group_by(idno) %>%
      summarise(total_conditions = sum(total_conditions, na.rm = TRUE))  
  #* Calculate incidence, raw and as proportion, and as binned
    hc_lifespan_data <- total_HCs_LL %>%
      left_join(class_census %>% select(idno, new_class_bw, le_wk), by = "idno") %>%
      mutate(
        hc_per_yr = total_conditions / (le_wk/52)
      ) %>%
      mutate(
        condition_bin = case_when(
          total_conditions == 0 ~ "0",
          total_conditions == 1 ~ "1",
          total_conditions == 2 ~ "2",
          total_conditions == 3 ~ "3",
          total_conditions == 4 ~ "4",
          total_conditions >= 5 ~ "5+"
        )
      ) %>%
      mutate(condition_bin = factor(condition_bin, levels = c("0", "1", "2", "3", "4", "5+")
      )) %>%
      mutate(hc_yr_bin = case_when(
        hc_per_yr == 0 ~ "[0]",
        hc_per_yr < 0.5 ~ "(0–0.5)",
        hc_per_yr < 1.0 ~ "[0.5–1)",
        hc_per_yr < 2 ~ "[1–2)",
        hc_per_yr < 4 ~ "[2–4)",
        TRUE ~ "≥4"
      )) %>%
        mutate(hc_yr_bin = factor(hc_yr_bin, levels = c("[0]", "(0–0.5)", "[0.5–1)", "[1–2)", "[2–4)", "≥4")))
  #* Plot raw counts over lifetime
    raw_counts_LL_HCs <- ggplot(hc_lifespan_data, aes(x = condition_bin, fill = factor(new_class_bw))) +
    geom_bar(position = "stack") +
    scale_fill_manual(values = c("#f2605d", "#36b22b", "#5087fe"), name = "Class") +
    labs(
      title = "Bins of Mice by Number of Conditions (Raw Counts)",
      x = "Number of Health Conditions",
      y = "Number of Mice"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.title = element_text(face = "bold", size = 16),
      axis.text = element_text(size = 12),
      legend.title = element_text(face = "bold", size = 16)
    )

  #* Count mice in each bin per class, plot as proportions
    #- Count
      bin_counts <- hc_lifespan_data %>%
        count(new_class_bw, condition_bin, name = "count")
      class_totals <- hc_lifespan_data %>%
        count(new_class_bw, name = "total")
      bin_props <- bin_counts %>%
        left_join(class_totals, by = "new_class_bw") %>%
        mutate(proportion = count / total)
    #- Graph with regular bars stacked
      proportion_LL_HCs <- ggplot(bin_props, aes(x = condition_bin, y = proportion, fill = factor(new_class_bw))) +
        geom_bar(stat = "identity", position = "stack") +
        scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
        scale_fill_manual(values = c("#f2605d", "#36b22b", "#5087fe"), name = "Class") +
        labs(
          title = "Bins of Mice by Number of Conditions (Proportions)",
          x = "Number of Health Conditions",
          y = "Proportion of Mice"
        ) +
        theme_minimal(base_size = 14) +
        theme(
          plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
          axis.title = element_text(face = "bold", size = 14),
          axis.text = element_text(size = 12)
      )
    #- Graph with stacked proportions/parts of whole
      bin_props_stacked <- hc_lifespan_data %>%
        count(new_class_bw, condition_bin, name = "count") %>%
        group_by(new_class_bw) %>%
        mutate(proportion = count / sum(count)) %>%
        ungroup()

      proportion_LL_HCs_stacked <- ggplot(bin_props_stacked, aes(x = factor(new_class_bw), y = proportion, fill = condition_bin)) +
        geom_bar(stat = "identity", position = "stack") +
        scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
        scale_fill_manual(values = bin_colors, name = "# of HCs") +
        labs(
          title = "Distribution of Condition Burden by Class",
          x = "Class",
          y = "Proportion of Mice"
        ) +
        unified_plot_theme
  #* Plot time-normalized incidence
    bin_props_yr <- hc_lifespan_data %>%
      count(new_class_bw, hc_yr_bin, name = "count") %>%
      group_by(new_class_bw) %>%
      mutate(total = sum(count), proportion = count / total) %>%
      ungroup()
    proportion_LL_HCs_per_year <- ggplot(bin_props_yr, aes(x = factor(new_class_bw), y = proportion, fill = hc_yr_bin)) +
      geom_bar(stat = "identity", position = "stack") +
      scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
      scale_fill_manual(values = bin_colors_per_yr, name = "HCs/Yr") +
      labs(
        title = "Distribution of Condition Incidence (per Year) by Class",
        x = "Class",
        y = "Proportion of Mice"
      ) +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
        axis.title = element_text(face = "bold", size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 12)
      )
```

## Examine cumulative HC free
```{r HC free study}
  #* Prepare cumulative incidence tibble
    #- Bring in the DOB from the census
      census <- as_tibble(census) %>%
        mutate(dob = as.Date(dob, format = "%m/%d/%Y")) %>%
        select(idno, dob)
    #- Create a time-to-event tibble and join DOB
      HC_relevant_LL <- classes_i %>%
        left_join(census, by = "idno") %>%
        mutate(
          le_days = le_wk * 7,
          event_day = as.integer(created_date - dob) # days since birth
        ) %>%
        select(idno, tod,dob, created_date, le_days, new_class_bw, event_day, abscess:condition_to_death) %>%
        arrange(desc(event_day))
    #- Compute first event per mouse or censor at le_days if none
      cumulative_df_fixed <- HC_relevant_LL %>%
        group_by(idno, new_class_bw, dob, le_days) %>%
        reframe(
          first_event_day = if (all(is.na(event_day))) NA_integer_ else min(event_day, na.rm = TRUE),
          event_day = if (all(is.na(event_day))) le_days else min(event_day, na.rm = TRUE)
        ) %>%
        mutate(
          Class = factor(new_class_bw, levels = c(1, 2, 3)),
          event = if_else(is.na(first_event_day), 0, 1),
          pct_lifespan = event_day / le_days
        ) %>%
        filter(is.finite(pct_lifespan), pct_lifespan >= 0, pct_lifespan <= 1)
    #- Compute cumulative proportion disease-free by class, fixed denominator, raw days
      cumulative_class_curves_day <- cumulative_df_fixed %>%
        group_by(Class) %>%
        arrange(event_day, .by_group = TRUE) %>%
        mutate(
          class_n = n(),
          prop_disease_free_day = 1 - cumsum(event) / class_n
        ) %>%
        ungroup()
    #- Compute cumulative proportion disease-free by class, fixed denominator, % lifespan
      cumulative_class_curves_pct <- cumulative_df_fixed %>%
        group_by(Class) %>%
        arrange(pct_lifespan, .by_group = TRUE) %>%
        mutate(
          class_n = n(),
          prop_disease_free_pct = 1 - cumsum(event) / class_n
        ) %>%
        ungroup()
  #* Plot
    #- As raw days
      days_HCfree <- ggplot(cumulative_class_curves, aes(x = event_day, y = prop_disease_free, color = Class)) +
        geom_step(size = 1.1) +
        scale_color_manual(values = class_colors) +
        labs(
          title = "Cumulative Proportion of Mice HC-free Over Time",
          x = "Days Since Birth",
          y = "Proportion HC-Free",
          color = "Class"
        ) +
        theme_minimal(base_size = 14) +
        theme(
          plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
          axis.title = element_text(face = "bold"),
          legend.title = element_text(face = "bold")
        ) +
        ylim(0, 1)
    #- As % LE
    percent_LE_HCfree <- ggplot(cumulative_class_curves_pct, aes(x = pct_lifespan, y = prop_disease_free_pct, color = Class)) +
      geom_step(size = 1.1) +
      scale_color_manual(values = class_colors) +
      labs(
        title = "Cumulative Proportion of Mice HC-free Over % of Lifespan",
        x = "% Lifespan",
        y = "Proportion HC-Free",
        color = "Class"
      ) +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")
      ) +
      ylim(0, 1)
```

## Examine cumulative incidence
```{r Cumulative HCs study}
  #* Compute cumulative HC tibbles
    #- Compute class sizes
      class_sizes <- classes_i %>%
        distinct(idno, new_class_bw) %>%
        count(new_class_bw, name = "class_size") %>%
        mutate(Class = factor(new_class_bw, levels = c(1, 2, 3)))
    #- Compute a version for only mice who had HCs
      cumulative_HC_day_events_only <- HC_relevant_LL %>%
        filter(!is.na(event_day)) %>% # excludes mice with 0 events
        mutate(Class = factor(new_class_bw, levels = c(1, 2, 3))) %>%
        group_by(Class) %>%
        mutate(class_size = n_distinct(idno)) %>%
        group_by(Class, event_day) %>%
        summarise(
          total_HC = n(),
          class_size = first(class_size),
          .groups = "drop"
        ) %>%
        arrange(Class, event_day) %>%
        group_by(Class) %>%
        mutate(
          cum_HC = cumsum(total_HC),
          cum_HC_per_mouse_events_only = cum_HC / class_size
        ) %>%
        ungroup()
    #- Compute a version which considers all mice
      cumulative_HC_day_all_mice <- HC_relevant_LL %>%
        filter(!is.na(event_day)) %>%
        mutate(Class = factor(new_class_bw, levels = c(1, 2, 3))) %>%
        group_by(Class, event_day) %>%
        summarise(total_HC = n(), .groups = "drop") %>%
        left_join(class_sizes, by = "Class") %>%
        arrange(Class, event_day) %>%
        group_by(Class) %>%
        mutate(
          cum_HC = cumsum(total_HC),
          cum_HC_per_mouse_all = cum_HC / class_size
        ) %>%
        ungroup()
  #* Plots
  #- Including only mice with HCs
    cum_inc_HCmice <- ggplot(cumulative_HC_day_events_only, aes(x = event_day, y = cum_HC_per_mouse_events_only, color = Class)) +
      geom_step(size = 1.1) +
      scale_color_manual(values = class_colors) +
      labs(
        title = "Cumulative HCs per Mouse (Only Mice With Events)",
        x = "Days Since Birth",
        y = "Cumulative HCs / Mouse",
        color = "Class"
      ) +
      theme_minimal(base_size = 14)
  #- Including mice who never had HC
    cum_inc_allmice <- ggplot(cumulative_HC_day_all_mice, aes(x = event_day, y = cum_HC_per_mouse_all, color = Class)) +
      geom_step(size = 1.1) +
      scale_color_manual(values = class_colors) +
      labs(
        title = "Cumulative HCs per Mouse (Including Mice with No Events)",
        x = "Days Since Birth",
        y = "Cumulative HCs / Mouse",
        color = "Class"
      ) +
      theme_minimal(base_size = 14)
```

## Examine clusters of conditions
```{r plot clusters and heatmaps}
  #* Examine overrepresentation of any HC in classes, only counting once per mouse
    #- Select relevant columns from HC data
      HC_relevant_LL_clusters <- HC_relevant_LL %>%
        select(idno, new_class_bw, abscess:weight_loss)
    #- Pivot long, count conditions
      long_cond <- HC_relevant_LL_clusters %>%
        pivot_longer(cols = -c(idno, new_class_bw), names_to = "condition", values_to = "present") %>%
        filter(present == 1) %>%
        distinct(idno, new_class_bw, condition)
      cond_counts <- long_cond %>%
        count(new_class_bw, condition, name = "num_mice_with_condition")
    #- Get total number of unique mice per class, calculate prevalence
      class_totals <- HC_relevant_LL_clusters %>%
        distinct(idno, new_class_bw) %>%
        count(new_class_bw, name = "n_mice")
      cond_prevalence <- cond_counts %>%
        left_join(class_totals, by = "new_class_bw") %>%
        mutate(prevalence = num_mice_with_condition / n_mice * 100)
    #- Identify conditions to keep (over 1% prevalence in all classes)
      conditions_to_keep <- cond_prevalence %>%
        group_by(condition) %>%
        summarise(max_prevalence = max(prevalence)) %>%
        filter(max_prevalence > 1) %>%
        pull(condition)
    #- Retain only those conditions
      HC_LL_filtered <- HC_relevant_LL_clusters %>%
        select(idno, new_class_bw, all_of(conditions_to_keep))
    #- Count number of unique mice with each condition in each class
      cond_cluster_counts <- HC_LL_filtered %>%
        pivot_longer(cols = -c(idno, new_class_bw), names_to = "condition", values_to = "present") %>%
        filter(present == 1) %>%
        distinct(idno, new_class_bw, condition) %>%
        count(new_class_bw, condition, name = "num_mice_with_condition")
    #- Calculate prevalence per class
      cond_cluster_prevalence <- cond_cluster_counts %>%
        left_join(class_totals, by = "new_class_bw") %>%
        mutate(prevalence = num_mice_with_condition / n_mice * 100)
    #- Prepare wide binary format for chi-squared/Fisher's test
      cond_test_input <- HC_LL_filtered %>%
        pivot_longer(cols = -c(idno, new_class_bw), names_to = "condition", values_to = "present") %>%
        group_by(condition) %>%
        nest()
    #- Run chi-squared or Fisher’s test per condition
      cond_cluster_tests <- cond_test_input %>%
        mutate(
          test = map(data, ~ {
            tab <- table(.x$present, .x$new_class_bw)
            # Check expected counts from chi-sq test
            expected <- suppressWarnings(chisq.test(tab)$expected)
            if (any(expected < 5)) {
              fisher.test(tab)
            } else {
              chisq.test(tab)
            }
          }),
          tidy = map(test, broom::tidy)
        ) %>%
        unnest(tidy) %>%
        select(condition, method, p.value, statistic) %>%
        arrange(p.value)
    #- Join p-values back to prevalence table, rename with clean names
      cond_cluster_results_clean <- cond_cluster_prevalence %>%
        left_join(cond_cluster_tests, by = "condition") %>%
        mutate(
          Significance = if_else(p.value <= 0.05, "Y", "N"),
          condition = recode(condition, !!!condition_renames)
        )
    #- Create condition matrix
    condition_matrix <- cond_cluster_results_clean %>%
      select(new_class_bw, condition, prevalence) %>%
      pivot_wider(names_from = new_class_bw, values_from = prevalence) %>%
      column_to_rownames("condition") %>%
      as.matrix()
    condition_matrix[is.na(condition_matrix)] <- 0
  #* Create heatmap
    #- Format display numbers to one decimal place,rename col labels
      display_vals <- format(round(condition_matrix, 1), nsmall = 1)
      colnames(condition_matrix) <- paste("Class", colnames(condition_matrix))
    #- Create row annotation for significance
      sig_annot <- cond_cluster_results_clean %>%
        distinct(condition, Significance) %>%
        column_to_rownames("condition")
      colnames(sig_annot) <- "Significant Difference"
      annotation_colors <- list(
        "Significant Difference" = c("Y" = "black", "N" = "white")
      )
    #- Plot heatmap
      heatmap_grob <- pheatmap::pheatmap(
        condition_matrix,
        scale = "none", # Fix inconsistent coloring
        cluster_rows = TRUE,
        cluster_cols = TRUE,
        annotation_row = sig_annot,
        annotation_colors = annotation_colors,
        annotation_legend = TRUE,
        annotation_names_row = FALSE,
        display_numbers = display_vals,
        number_color = "gray20",
        fontsize_row = 13,
        fontsize_col = 13,
        fontsize_number = 10,
        angle_col = 0,
        legend = FALSE
      )
```

## Examine clusters of conditions but normalize by time
```{r plot clusters and heatmaps norm by time}
  #* Examine overrepresentation, but norm by time
    #- Reshape and include lifespan
      condition_cols <- names(HC_relevant_LL)[which(names(HC_relevant_LL) == "abscess"):which(names(HC_relevant_LL) == "weight_loss")]
    #- Create full idno × condition table including 0s, then compute rate per year
      long_cond_all <- HC_relevant_LL %>%
        select(idno, new_class_bw, le_days, all_of(condition_cols)) %>%
        pivot_longer(cols = all_of(condition_cols), names_to = "condition", values_to = "present") %>%
        mutate(present = replace_na(present, 0)) %>%
        group_by(idno, new_class_bw, condition, le_days) %>%
        summarise(n = sum(present), .groups = "drop") %>%
        mutate(
          le_years = le_days / 365.25,
          rate_per_year = n / le_years
        )
    #- Compute average rate per condition and class
      cond_class_rates <- long_cond_all %>%
        group_by(new_class_bw, condition) %>%
        summarise(mean_rate = mean(rate_per_year), .groups = "drop")
    #- Now compute p values via Kruskal-Wallis test
      cond_test_input_norm <- long_cond_all %>%
        filter(condition %in% conditions_to_keep)
      cond_cluster_tests_norm <- cond_test_input_norm %>%
        group_by(condition) %>%
        nest() %>%
        mutate(
          test = map(data, ~ kruskal.test(rate_per_year ~ as.factor(new_class_bw), data = .x)),
          tidy = map(test, broom::tidy)
        ) %>%
        unnest(tidy) %>%
        select(condition, p.value, statistic, method = method)
    #- Filter down to only the common conditions defined earlier
      filtered_rates <- cond_class_rates %>%
        filter(condition %in% conditions_to_keep) %>%
        left_join(cond_cluster_tests_norm, by = "condition") %>%
          mutate(
            Significance = if_else(p.value <= 0.05, "Y", "N"),
            condition = recode(condition, !!!condition_renames)
          )
    #- Now structure matrix for heatmap
    condition_matrix_norm <- filtered_rates %>%
      select(new_class_bw, condition, mean_rate) %>% # Keep only numeric data
      pivot_wider(names_from = new_class_bw, values_from = mean_rate) %>%
      replace(is.na(.), 0) %>%
      column_to_rownames("condition") %>%
      as.matrix()
  #* Create heatmap
    #- Format display numbers to one decimal place,rename col labels
      colnames(condition_matrix_norm) <- paste("Class", colnames(condition_matrix_norm))
      display_vals_norm <- format(round(condition_matrix_norm, 2), nsmall = 2)
    #- Create row annotation for significance
      sig_annot_timed <- filtered_rates %>%
        distinct(condition, Significance) %>%
        column_to_rownames("condition")
      colnames(sig_annot_timed) <- "Significant Difference"
      annotation_colors_timed <- list(
        "Significant Difference" = c("Y" = "black", "N" = "white")
      )
    #- Plot heatmap
      heatmap_timed <- pheatmap::pheatmap(
        condition_matrix_norm,
        scale = "none", # Fix inconsistent coloring
        cluster_rows = TRUE,
        cluster_cols = TRUE,
        annotation_row = sig_annot_timed,
        annotation_colors = annotation_colors_timed,
        annotation_legend = TRUE,
        annotation_names_row = FALSE,
        display_numbers = display_vals_norm,
        number_color = "gray20",
        fontsize_row = 13,
        fontsize_col = 13,
        fontsize_number = 10,
        angle_col = 0,
        legend = TRUE # You probably want this to show the color key now
      )
```     

## Compile plots
```{r compiled plots}
  # Unified theme
  unified_plot_theme <- theme_minimal(base_size = 18) +
    theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 16, face = "bold"),
      axis.text = element_text(size = 13),
      legend.title = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 13, face = "bold")
    )
  #* Start a PDF device, compile all the pages
  pdf("Healthcard_Figures.pdf", width = 8.5, height = 11, onefile = TRUE)

    # Helper: top-left caption only (no plot shift)
    draw_top_left_caption <- function(fig_num, caption_text) {
      title_part <- paste0("Healthcard Figure ", fig_num, ".")
      grid.text(
        label = title_part,
        x = unit(0.01, "npc"),
        y = unit(0.99, "npc"),
        hjust = 0, vjust = 1,
        gp = gpar(fontsize = 13, fontface = "bold.italic")
      )
      grid.text(
        label = paste(strwrap(caption_text, width = 100), collapse = "\n"),
        x = unit(0.01, "npc"),
        y = unit(0.965, "npc"),
        hjust = 0, vjust = 1,
        gp = gpar(fontsize = 13, fontface = "italic")
      )
    }

    # --- Figure 1: EOL_sig_plot_unified ---
    grid.newpage()
    draw_top_left_caption(1, "Prevalence of conditions that differed between classes at the end of life.")
    pushViewport(viewport(width = 0.8, height = 0.8))
    g1 <- arrangeGrob(
      EOL_sig_plot_unified + unified_plot_theme +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)),
      nullGrob(),
      heights = c(0.6, 0.3)
    )
    grid.draw(g1)
    popViewport()

    # --- Figure 2 ---
    grid.newpage()
    draw_top_left_caption(2, "Distribution of number of conditions per mouse accross classes expressed as raw counts (upper panel) and proportions of the class size (lower panel).")
    pushViewport(viewport(width = 0.8, height = 0.8))
    g2 <- arrangeGrob(
      raw_counts_LL_HCs + unified_plot_theme,
      proportion_LL_HCs + unified_plot_theme,
      heights = c(0.45, 0.45),
      ncol = 1
    )
    grid.draw(g2)
    popViewport()

    # --- Figure 3 ---
    grid.newpage()
    draw_top_left_caption(3, "Parts-of-whole distribution of condition burden per mouse accross classes expressed as prevalence (upper panel) and time-normalized incidence, i.e., health conditions per year (lower panel).")
    pushViewport(viewport(width = 0.8, height = 0.8))
    g3 <- arrangeGrob(
      proportion_LL_HCs_stacked + unified_plot_theme,
      proportion_LL_HCs_per_year + unified_plot_theme,
      heights = c(0.45, 0.45),
      ncol = 1
    )
    grid.draw(g3)
    popViewport()

    # --- Figure 4 ---
    grid.newpage()
    draw_top_left_caption(4, "Cumulative proportion of mice without any health conditions/health cards over the lifespan, expressed according to time (upper panel) and as a percentage of the entire lifespan (lower panel).")
    pushViewport(viewport(width = 0.8, height = 0.8))
    g4 <- arrangeGrob(
      days_HCfree + unified_plot_theme,
      percent_LE_HCfree + unified_plot_theme,
      heights = c(0.45, 0.45),
      ncol = 1
    )
    grid.draw(g4)
    popViewport()

    # --- Figure 5 ---
    grid.newpage()
    draw_top_left_caption(5, "Cumulative incidence of health conditions/health cards per mouse over time with only mice who ever had HCs (top panel) and including mice who never had HCs (bottom panel).")
    pushViewport(viewport(width = 0.8, height = 0.8))
    g5 <- arrangeGrob(
      cum_inc_HCmice + unified_plot_theme,
      cum_inc_allmice + unified_plot_theme,
      heights = c(0.45, 0.45),
      ncol = 1
    )
    grid.draw(g5)
    popViewport()

    # --- Figure 6 ---
    grid.newpage()
    draw_top_left_caption(6, "Heatmap with clustering of prevalence of specific conditions accross classes. Black rectangular annotations on the left indicate significant differences between groups. Prevalence per class are printed in the individual squares.")
    pushViewport(viewport(width = 0.80, height = 0.75))
    grid.draw(heatmap_grob$gtable)
    popViewport()

    # --- Figure 7 ---
    grid.newpage()
    draw_top_left_caption(7, "Heatmap with clustering of time-normalized incidence, i.e., specific health conditions per year, accross classes. Black rectangular annotations on the left indicate significant differences between groups. Incidence per class (condition events/yr) are printed in the squares.")
    pushViewport(viewport(width = 0.85, height = 0.75))
    grid.draw(heatmap_timed$gtable)
    popViewport()
  #* Close the PDF device
  dev.off()
```  