---
title: "Dataset Creation for Trajectory SLAM"
author: "Eric Shiroma, William Mueller"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
library(psych)
library(kableExtra)
library(mgcv)
library(ggplot2)
library(ggpubr)
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis 
#### census data read in
  - input file census.csv (10/30/2019)
  - edit csv column name 'Animal_ID'
  
```{r slam_census, warning=FALSE, message=FALSE, echo=FALSE}
# edit csv column name 'Animal_ID'
census <- read.csv("data/census.csv") %>%
  # census <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/census.csv") %>%
  select(idno, animal = Animal_ID, tag, sex, strain, dob, cohort, taghistory)
census$dob <- as.Date(census$dob, "%m/%d/%Y")
census$idno <- as.character(census$idno)
census$tag <- as.character(census$tag)
census$taghistory <- as.character(census$taghistory)
``` 

#### body weight data read in and clean
  - input file BW_Temp_FC_All_2019-11-16 (11/16/2019)
  - input file NMR.csv (10/30/2019)
  - clean body weights (funny values) 
    - text to missing 
    - bw < 10 to missing 
    - bw > 100 to missing
  - merge nmr body weights into main body weight dataset (bodywt)
    - bodywt_nmr
  
```{r slam_bodyweight, warning=FALSE, message=FALSE, echo=FALSE}
bodywt <- read.csv("data/BW_Temp_FC_All_2021-12-23.csv") %>%
  # bodywt <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/BW_Temp_FC_All_2019-11-16.csv") %>%
  select(idno, date, bw, foodcons, temp)

bodywt <- bodywt %>%
  mutate(
    bw = ifelse(bw == "", NA,
      ifelse(bw == "NA", NA,
        ifelse(bw == "wholecaedead", NA,
          ifelse(bw == "MIA", NA,
            ifelse(bw == "fd5-23", NA,
              ifelse(bw == "FD", NA,
                ifelse(bw == "f/d6/7/2018", NA,
                  ifelse(bw == "f/d2/7/18", NA,
                    ifelse(bw == "f/d2/27", NA,
                      ifelse(bw == "DVR", NA,
                        ifelse(bw == "deadd", NA,
                          ifelse(bw == "dead5-9-18", NA,
                            ifelse(bw == "dead5-8-18", NA,
                              ifelse(bw == "dead5-21-18", NA,
                                ifelse(bw == "dead5-14", NA,
                                  ifelse(bw == "dead5-10", NA,
                                    ifelse(bw == "dead1-30", NA,
                                      ifelse(bw == "dead", NA,
                                        ifelse(bw == "Dead", NA,
                                          ifelse(bw == "DEAD", NA,
                                            ifelse(bw == "BW", NA,
                                              ifelse(bw == "culled5-18", NA, as.character(bw))
                                            )
                                          )
                                        )
                                      )
                                    )
                                  )
                                )
                              )
                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  )
bodywt$date <- as.Date(bodywt$date, "%Y-%m-%d")
bodywt$bw <- as.numeric(as.character(bodywt$bw))
bodywt$idno <- as.character(bodywt$idno)
bw_na_before <- sum(is.na(bodywt$bw))
bodywt <- bodywt %>%
  mutate(
    bw = ifelse(bw < 10, NA,
      ifelse(bw > 100, NA, bw)
    )
  )
bw_na_after <- sum(is.na(bodywt$bw))

bodywt <- bodywt %>%
  filter(!is.na(bw))

duplicated_bw <- sum(duplicated(bodywt[c(1, 2)]))
bodywt <- bodywt[!duplicated(bodywt[c(1, 2)]), ]

## from NMR
bodywt_nmr <- read.csv("data/nmr_compiled_2021-11-10.csv") %>%
  select(idno, date, bw)

bodywt_nmr$date <- as.Date(bodywt_nmr$date, "%Y-%m-%d")
bodywt_nmr$bw <- as.numeric(as.character(bodywt_nmr$bw))
bodywt_nmr$idno <- as.character(bodywt_nmr$idno)
bodywt_nmr <- bodywt_nmr %>%
  mutate(
    bw = ifelse(bw < 10, NA,
      ifelse(bw > 100, NA, bw)
    )
  )

bodywt <- bodywt %>%
  full_join(bodywt_nmr)

bodywt <- bodywt[!duplicated(bodywt[c(1, 2)]), ]

hist(bodywt$bw)
```

- duplicated bw = `r duplicated_bw`
- NA bw before outlier removal `r bw_na_before`
- NA bw after outlier removal `r bw_na_after`
- bw outlier `r bw_na_after - bw_na_before`

#### glucose data read in and clean 
  - input file Glucose_Lactate.csv (10/30/2019)
  - edit csv column name 'Stress_Notes'
  - clean glucose (funny values) 
    - text to missing 
    - gluc < 20 or gluc >= 300 to missing 
    
```{r slam_glucose, warning=FALSE, message=FALSE, echo=FALSE}
# edit csv column name 'Stress_Notes'
glucose <- read.csv("data/gluc_lact_2021-12-01.csv") %>%
  # glucose <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/Glucose_Lactate.csv") %>%
  select(idno, date, gluc = Glucose) %>%
  mutate(
    dead = ifelse(gluc == "DEAD", 1, NA),
    gluc = ifelse(gluc == "DEAD", NA,
      ifelse(gluc == "dead", NA,
        ifelse(gluc == "missed", NA,
          ifelse(gluc == "missing", NA,
            ifelse(gluc == "NOT BLED", NA,
              ifelse(gluc == "SAC", NA,
                ifelse(gluc == "89/148", NA,
                  ifelse(gluc == "", NA,
                    ifelse(gluc == "no glucose", NA,
                      ifelse(gluc == "Not accounted for", NA,
                        ifelse(gluc == "would not bleed", NA,
                          ifelse(gluc == "", NA, as.character(gluc))
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  )
glucose$date <- as.Date(glucose$date, "%Y-%m-%d")
glucose$gluc <- as.numeric(glucose$gluc)
glucose$idno <- as.character(glucose$idno)
gluc_na_before <- sum(is.na(glucose$gluc))
glucose <- glucose %>%
  mutate(
    gluc = ifelse(gluc < 20 | gluc >= 300, NA, gluc)
  )

gluc_na_after <- sum(is.na(glucose$gluc))
duplicated_gluc <- sum(duplicated(glucose[c(1, 2)]))

glucose <- glucose[!duplicated(glucose[c(1, 2)]), ]

hist(glucose$gluc)
```

- duplicated gluc = `r duplicated_gluc`
- NA gluc before outlier removal `r gluc_na_before`
- NA gluc after outlier removal `r gluc_na_after`
- gluc outlier `r gluc_na_after - gluc_na_before`

#### body comp data read in and clean 
  - input file NMR.csv (10/30/2019)
  - clean nmr (funny values) 
    - text to missing 
    - fat <= 0 to missing
    - lean <= 5 to NA
    - lean >= 40 to NA

```{r slam_nmr, warning=FALSE, message=FALSE, echo=FALSE}
nmr <- read.csv("data/nmr_compiled_2021-11-10.csv") %>%
  # nmr <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/NMR.csv") %>%
  select(idno, date, fat, lean, fluid)

invalid_fat <- nmr$fat[is.na(as.numeric(as.character(nmr$fat)))]

nmr <- nmr %>%
  mutate(fat = ifelse(fat %in% invalid_fat, NA, fat))

nmr$date <- as.Date(nmr$date, "%Y-%m-%d")
nmr$idno <- as.character(nmr$idno)
nmr$fat <- as.numeric(as.character(nmr$fat))
fat_na_before <- sum(is.na(nmr$fat))
lean_na_before <- sum(is.na(nmr$lean))
nmr <- nmr %>%
  mutate(
    fat = ifelse(fat <= 0, NA, fat)
  ) %>%
  mutate(
    lean = ifelse(lean <= 5, NA,
      ifelse(lean >= 40, NA, lean)
    )
  )
fat_na_after <- sum(is.na(nmr$fat))
lean_na_after <- sum(is.na(nmr$lean))

duplicated_nmr <- sum(!duplicated(nmr[c(1, 2)]))
nmr <- nmr[!duplicated(nmr[c(1, 2)]), ]

hist(nmr$fat)
hist(nmr$lean)
```

- duplicated nmr = `r duplicated_nmr`
- NA fat before outlier removal `r fat_na_before`
- NA fat after outlier removal `r fat_na_after`
- fat outlier `r fat_na_after - fat_na_before`
- NA lean before outlier removal `r lean_na_before`
- NA lean after outlier removal `r lean_na_after`
- lean outlier `r lean_na_after - lean_na_before`

```{r all_to_new_idno, warning=FALSE, message=FALSE, echo=FALSE}
# changing old idnos to new idnos in body comp file
fixID_history <- function(idno, census) {
  census <- census %>%
    separate(taghistory, c("th1", "th2", "th3", "th4", "th5", "th6"), sep = ",", extra = "warn")

  idnos <- trimws(idno)
  idnos[idnos %in% census$Animal.ID] <- census$idno[match(idnos[idnos %in% census$Animal.ID], census$Animal.ID)]
  idnos[idnos %in% census$tag] <- census$idno[match(idnos[idnos %in% census$tag], census$tag)]
  idnos[idnos %in% census$th1] <- census$idno[match(idnos[idnos %in% census$th1], census$th1)]
  idnos[idnos %in% census$th2] <- census$idno[match(idnos[idnos %in% census$th2], census$th2)]
  idnos[idnos %in% census$th3] <- census$idno[match(idnos[idnos %in% census$th3], census$th3)]
  idnos[idnos %in% census$th4] <- census$idno[match(idnos[idnos %in% census$th4], census$th4)]
  idnos[idnos %in% census$th5] <- census$idno[match(idnos[idnos %in% census$th5], census$th5)]
  idnos[idnos %in% census$th6] <- census$idno[match(idnos[idnos %in% census$th6], census$th6)]

  return(idnos)
}

# pull tag history if needed
nmr$idno <- fixID_history(nmr$idno, census)
```

#### survival data read in and clean 
  - input file Survival.csv (10/30/2019)
  - clean survival (funny values) 
    - text to missing 
    - recode old to new idnos 
    
```{r slam_survival, warning=FALSE, message=FALSE, echo=FALSE}
surv <- read.csv("data/survival_2021-12-17.csv") 
# surv <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/Survival.csv")
surv$tod <- as.Date(surv$tod, "%m/%d/%y")
surv$tag <- as.character(surv$tag)
surv$cod <- as.character((surv$cod))
surv <- surv %>%
  mutate(tag = ifelse(tag == "R", "179060",
    ifelse(tag == "L", "179217", tag)
  ))
```

#### merging all datasets into a main file and clean
  - glucose
  - bodywt (body weight)
  - nmr (body fat)
  - census
  - removal of timepoints before dob or mouse facility arrival date
  
```{r slam_main_all, warning=FALSE, message=FALSE, echo=FALSE}
main_all <- glucose %>%
  full_join(bodywt, by = c("idno", "date")) %>%
  full_join(nmr, by = c("idno", "date")) %>%
  left_join(census, by = "idno") %>%
  mutate(
    date = if_else(cohort == 1 & date >= "2015-11-12", date,
      if_else(cohort == 2 & date >= "2016-02-03", date,
        if_else(cohort == 3 & date >= "2016-06-06", date,
          if_else(cohort == 4 & date >= "2016-08-08", date,
            if_else(cohort == 5 & date >= "2016-11-07", date,
              if_else(cohort == 6 & date >= "2017-01-11", date,
                if_else(cohort == 7 & date >= "2017-03-14", date,
                  if_else(cohort == 8 & date >= "2017-05-30", date,
                    if_else(cohort == 9 & date >= "2017-08-22", date,
                      if_else(cohort == 10 & date >= "2017-11-14", date, as.Date(NA))
                    )
                  )
                )
              )
            )
          )
        )
      )
    ),
    age_wk = round(((date - dob) / 7), 2) + 1
  ) %>%
  filter(!is.na(date) & !is.na(idno))

before_age_filter <- nrow(main_all)

main_all <- main_all %>%
  filter(age_wk > 0)

after_age_filter <- nrow(main_all)

main_all$age_wk <- as.numeric(main_all$age_wk)
```

- Before age filter = `r before_age_filter`
- After age filter = `r after_age_filter`
  
#### restriction to cohorts 1 to 10

```{r slam_ratio, warning=FALSE, message=FALSE, echo=FALSE}
main_all2 <- main_all %>% filter(cohort >= 1 & cohort <= 10)

# getting unique idnos
main_all2_n <- n_distinct(main_all2$idno)

sexstrain <- main_all2 %>%
  select(idno, sex, strain) %>%
  distinct()

sextable <- table(sexstrain$sex)
straintable <- table(sexstrain$strain)

# calculate % sex and strain
slam_sex_ratio <- format(round(((sextable[1] / (sextable[1] + sextable[2])) * 100), 2), nsmall = 2)
slam_strain_ratio <- format(round(((straintable[1] / (straintable[1] + straintable[2])) * 100), 2), nsmall = 2)

# renamed for isg template
slam_main_all <- main_all2
```

#### population characteristics
- `r main_all2_n` mice with valid indices 
- `r slam_sex_ratio` % female 
- `r slam_strain_ratio` % B6 

#### age characteristics

```{r slam_age_desc, warning=FALSE, message=FALSE, echo=FALSE}
main_all2_desc <- main_all2 %>%
  group_by(idno, sex, strain) %>%
  summarise(
    age_wk = min(age_wk, na.rm = T)
  )

main_all2_desc2_all <- format(round(psych::describe(main_all2_desc$age_wk), 2), nsmall = 2) %>%
  mutate_each(funs(as.numeric(.))) %>%
  mutate(group1 = "all") %>%
  select(-c("vars"))
main_all2_desc_sex <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$sex, mat = TRUE, digits = 2)) %>%
  select(-c("item", "vars")) %>%
  mutate(n = as.numeric(n))
main_all2_desc_strain <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$strain, mat = TRUE, digits = 2)) %>%
  select(-c("item", "vars")) %>%
  mutate(n = as.numeric(n))

main_all2_desc2 <- bind_rows(main_all2_desc2_all, main_all2_desc_sex, main_all2_desc_strain) %>%
  select(group = group1, everything())
```

```{r slam_age_desc_kable, warning=FALSE, message=FALSE, echo=FALSE}
kable(main_all2_desc2, caption = "Age Characteristics", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

*** 

#### census_c16_death creation
  - merge 
    - census
    - surv
    - main_all2
  - restrict to cohorts 1 to 10
  - remove mice with missing time of death
  - create all cause death censor category
    - all-cause mortality included: "Found dead", "Per PI", "Culled", "DVR or Pathology", "Culled Per Vet"
    
*** 

```{r slam_censor_df, warning=FALSE, message=FALSE, echo=FALSE}
census_c16_death <- main_all2 %>%
  select(idno, tag, dob) %>%
  distinct() %>%
  left_join(surv, by = "tag") %>%
    select(-age_wk) %>%
  left_join(main_all2, by = c("idno", "dob")) %>%

  group_by(
    idno, cod, dob, sex, strain,
    tod
  ) %>%
  summarise(
    maxdate = max(tod),
    age_first = min(age_wk, na.rm = T),
    firstdate = min(date, na.rm = T),
    lastdate = max(date, na.rm = T)
  ) %>%
  mutate(cod = tolower(cod)) %>%
  mutate(
    fu_age_wk = floor(((maxdate - firstdate) + 1) / 7),
    le_wk = floor(((maxdate - dob) + 1) / 7),
    dead_censor = ifelse(cod == "found dead", 1,
      ifelse(cod == "per pi", 1,
        ifelse(cod == "per vet", 1,
          ifelse(cod == "culled", 1,
            ifelse(cod == "culled per vet", 1,
              ifelse(cod == "cull per vet", 1,
                ifelse(cod == "cull per pi", 1,
                  ifelse(cod == "dvr or pathology", 1, 0)
                )
              )
            )
          )
        )
      )
    )
  ) %>%
  ungroup() %>%
  filter(!is.na(maxdate)) %>%
  select(
    idno, dead_censor, fu_age_wk, le_wk, cod, sex, strain,
    lastdate, maxdate, tod
  )

# ghost mice: timepoints after date of death
ghost <- census_c16_death %>%
  filter(lastdate > tod) %>%
  mutate(afterlife = lastdate - tod)

codtable <- as.data.frame(table(census_c16_death$cod))
dctable1 <- as.data.frame(table(census_c16_death$dead_censor))
fu_time_desc <- format(round(as.data.frame(psych::describe(as.numeric(census_c16_death$fu_age_wk))), 2), nsmall = 2)
```

#### cause of death 

```{r cod_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(codtable, caption = "Cause of Death", col.names = c("Cause of Death", "Frequency"), row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### death frequency 

```{r death_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(dctable1, caption = "Death Frequency", col.names = c("Dead = 1", "Frequency"), row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### follow up time 

```{r followup_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(fu_time_desc, caption = "Follow Up Time", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

*** 

#### life expectancy 

```{r slam_life_expectancy, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE}
# function to create life expectancy for different groups
create_LE <- function(data, criteria, filter) {
  life_exp <- data %>%
    ungroup() %>%
    filter(!!filter) %>%
    summarise(
      total = n_distinct(idno),
      dead = n_distinct(idno[!!criteria]),
      life_exp_wk = round(median(le_wk[!!criteria]), 2),
      q1 = round(quantile(le_wk[!!criteria])["25%"], 2),
      q3 = round(quantile(le_wk[!!criteria])["75%"], 2),
      sd = round(sd(as.numeric(le_wk[!!criteria])), 2),
      se = round(sd(as.numeric(le_wk[!!criteria])) / sqrt(dead), 2),
      lcl = round(life_exp_wk - 1.96 * se, 2),
      ucl = round(life_exp_wk + 1.96 * se, 2)
    )
  life_exp$life_exp_wk <- as.numeric(life_exp$life_exp_wk)

  return(life_exp)
}

life_exp <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("M", "F"))) %>%
  mutate(group = "all")
life_exp_f <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("F"))) %>%
  mutate(group = "female")
life_exp_m <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("M"))) %>%
  mutate(group = "male")
life_exp_het3 <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(strain %in% c("HET3"))) %>%
  mutate(group = "HET3")
life_exp_b6 <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(strain %in% c("B6"))) %>%
  mutate(group = "B6")

life_exp2 <- bind_rows(life_exp, life_exp_f, life_exp_m, life_exp_het3, life_exp_b6) %>%
  select(group, everything())

## for isg template
slam_life_exp <- life_exp
```

```{r slam_life_exp_kable, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE}
kable(life_exp2[, c(1:6)], caption = "Life Expectancy", col.names = c("Group", "Total", "Dead", "Life Expectancy <br> (weeks)", "iqr", "sd"), row.names = TRUE, align = "c", escape = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left")
```

*** 

#### create timepoints as percent life expectancy
- "per_age_wk" in main_all2
- "per_fu_wk" and "per_le_wk" in census_c16_death

```{r percent_le, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE}
census_c16_death <- census_c16_death %>%
  mutate(
    per_fu_age_wk =
      ifelse(sex == "M", (fu_age_wk / life_exp_m$life_exp_wk) * 100,
        ifelse(sex == "F", (fu_age_wk / life_exp_f$life_exp_wk) * 100, fu_age_wk / life_exp$life_exp_wk)
      ),
    per_le_wk =
      ifelse(sex == "M", (le_wk / life_exp_m$life_exp_wk) * 100,
        ifelse(sex == "F", (le_wk / life_exp_f$life_exp_wk) * 100, le_wk / life_exp$life_exp_wk)
      ),
  )

census_c16_death2 <- census_c16_death %>%
  select(idno, le_wk) %>%
  mutate(le_wk = as.numeric(le_wk))

# filter out timepoints after date of death
main_all2 <- main_all2 %>%
  left_join(census_c16_death2, by = "idno") %>%
  filter(age_wk <= le_wk) %>%
  mutate(per_age_wk = (age_wk / le_wk) * 100)

### filter for only dead
main_all2_dead_and_alive <- main_all2

main_all2 <- main_all2 %>%
  left_join(select(census_c16_death, idno, dead_censor), by = "idno") %>%
  filter(dead_censor == 1)

### rename for igs template
slam_main_all <- main_all2
```

***

#### frequencies of measures 

```{r slam_follow_up, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
n_asmt <- main_all2 %>%
  group_by(idno, sex) %>%
  summarise_all(funs(
    n = sum(!is.na(.)),
    sd = format(round(sd(!is.na(.)), 2), nsmall = 2)
  ))

gluc_asmt <- format(round(as.data.frame(psych::describe(n_asmt$gluc_n)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, everything(), -vars) %>%
  mutate(n = as.numeric(n))
bw_asmt <- format(round(as.data.frame(psych::describe(n_asmt$bw_n)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, everything(), -vars) %>%
  mutate(n = as.numeric(n))
fat_asmt <- format(round(as.data.frame(psych::describe(n_asmt$fat_n)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, everything(), -vars) %>%
  mutate(n = as.numeric(n))

## by sex
gluc_asmt_sex <- as.data.frame(describeBy(n_asmt$gluc_n, n_asmt$sex, mat = TRUE, digits = 2)) %>%
  select(-c("item", "vars")) %>%
  mutate(n = as.numeric(n))
bw_asmt_sex <- as.data.frame(describeBy(n_asmt$bw_n, n_asmt$sex, mat = TRUE, digits = 2)) %>%
  select(-c("item", "vars")) %>%
  mutate(n = as.numeric(n))
fat_asmt_sex <- as.data.frame(describeBy(n_asmt$fat_n, n_asmt$sex, mat = TRUE, digits = 2)) %>%
  select(-c("item", "vars")) %>%
  mutate(n = as.numeric(n))


## age desc by life expectancy
main_all2_desc_gluc <- main_all2 %>%
  filter(!is.na(gluc)) %>%
  group_by(idno, sex)
main_all2_desc_bw <- main_all2 %>%
  filter(!is.na(bw)) %>%
  group_by(idno, sex)
main_all2_desc_fat <- main_all2 %>%
  filter(!is.na(fat)) %>%
  group_by(idno, sex)

# by age weeks
allgluc_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_gluc$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, age_min = min, age_max = max)
allbw_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_bw$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, age_min = min, age_max = max)
allfat_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_fat$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, age_min = min, age_max = max)
## by sex
gluc_age_range <- as.data.frame(describeBy(main_all2_desc_gluc$age_wk, main_all2_desc_gluc$sex, mat = TRUE, digits = 2)) %>%
  select(group1, age_min = min, age_max = max)
bw_age_range <- as.data.frame(describeBy(main_all2_desc_bw$age_wk, main_all2_desc_bw$sex, mat = TRUE, digits = 2)) %>%
  select(group1, age_min = min, age_max = max)
fat_age_range <- as.data.frame(describeBy(main_all2_desc_fat$age_wk, main_all2_desc_fat$sex, mat = TRUE, digits = 2)) %>%
  select(group1, age_min = min, age_max = max)

## by percent age
allgluc_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_gluc$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, le_min = min, le_max = max)
allbw_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_bw$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, le_min = min, le_max = max)
allfat_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_fat$age_wk)), 2), nsmall = 2) %>%
  mutate(group1 = "all") %>%
  select(group1, le_min = min, le_max = max)
# by sex
gluc_perage_range <- as.data.frame(describeBy(main_all2_desc_gluc$age_wk, main_all2_desc_gluc$sex, mat = TRUE, digits = 2)) %>%
  select(group1, le_min = min, le_max = max)
bw_perage_range <- as.data.frame(describeBy(main_all2_desc_bw$age_wk, main_all2_desc_bw$sex, mat = TRUE, digits = 2)) %>%
  select(group1, le_min = min, le_max = max)
fat_perage_range <- as.data.frame(describeBy(main_all2_desc_fat$age_wk, main_all2_desc_fat$sex, mat = TRUE, digits = 2)) %>%
  select(group1, le_min = min, le_max = max)

##

gluc_asmt2 <- gluc_asmt %>%
  left_join(allgluc_age_range) %>%
  left_join(allgluc_perage_range)
gluc_asmt_sex2 <- gluc_asmt_sex %>%
  left_join(gluc_age_range) %>%
  left_join(gluc_perage_range)
bw_asmt2 <- bw_asmt %>%
  left_join(allbw_age_range) %>%
  left_join(allbw_perage_range)
bw_asmt_sex2 <- bw_asmt_sex %>%
  left_join(bw_age_range) %>%
  left_join(bw_perage_range)
fat_asmt2 <- fat_asmt %>%
  left_join(allfat_age_range) %>%
  left_join(allfat_perage_range)
fat_asmt_sex2 <- fat_asmt_sex %>%
  left_join(fat_age_range) %>%
  left_join(fat_perage_range)

gluc_amst3 <- rbind(gluc_asmt2, gluc_asmt_sex2)
bw_amst3 <- rbind(bw_asmt2, bw_asmt_sex2)
fat_amst3 <- rbind(fat_asmt2, fat_asmt_sex2)

all_asmt <- rbind(gluc_amst3, bw_amst3, fat_amst3)
all_asmt$indices <- c("glucose", "glucose", "glucose", "body weight", "body weight", "body weight", "NMR", "NMR", "NMR")
all_asmt <- all_asmt %>%
  select(indices, group = group1, everything())

## for template
slam_gluc_asmt <- gluc_asmt
slam_bw_asmt <- bw_asmt
slam_fat_asmt <- fat_asmt
```

```{r asmt_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}
kable(all_asmt, caption = "Assessment Frequencies", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  column_spec(c(1), border_right = TRUE)
```

#### age at peak indices value creation 

```{r slam_life_peaks, warning=FALSE, message=FALSE, echo=FALSE, cache=FALSE}
scaling <- round(0.05 * life_exp$life_exp_wk)
scaling_f <- round(0.05 * life_exp_f$life_exp_wk)
scaling_m <- round(0.05 * life_exp_m$life_exp_wk)

# for isg output
slam_scaling <- scaling
slam_scaling_f <- scaling_f
slam_scaling_m <- scaling_m

# functions to create peaks for different groups
lifespan_peaks <- function(metric, data, filter, mle, scale) {
  # set boundary conditions for peak to lowest and highest life expectancy timepoints
  # low_bound <- 0.25*mle
  # high_bound <- 1.25*mle
  low_bound <- 25
  high_bound <- 100

  # subset arguments
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)")
  metric2 <- paste0("!is.na(", metric, ")")

  # subset for non-missing data
  df <- base::subset(data, eval(parse(text = metric2)))

  # subset data by xlimit and subgroup restraints
  peaks_df <- subset(df, eval(parse(text = sub))) %>%
    arrange(per_age_wk) %>%
    filter(!!filter)

  # gam model fit using geom_smooth parameters
  gam_fit <- mgcv::gam(eval(parse(text = metric)) ~ s(per_age_wk, bs = "cs"), method = "REML", data = peaks_df)

  # create fitted values using gam model from dataframe age values
  predict1 <- as.data.frame(predict.gam(gam_fit, peaks_df, se.fit = TRUE))

  # create n and sd by age group
  fit2 <- peaks_df %>%
    bind_cols(predict1) %>%
    select(per_age_wk, fit, se.fit) %>%
    filter(per_age_wk >= low_bound, per_age_wk <= high_bound) %>%
    group_by(per_age_wk) %>%
    mutate(
      n = n(),
      sd = se.fit * sqrt(n)
    ) %>%
    ungroup()

  # create low bound peak object
  low_bound_peak <- fit2 %>%
    filter(per_age_wk == min(per_age_wk)) %>%
    select(per_age_wk, fit, se.fit, n, sd) %>%
    distinct()

  # create high bound peak object
  high_bound_peak <- fit2 %>%
    filter(per_age_wk == max(per_age_wk)) %>%
    select(per_age_wk, fit, se.fit, n, sd) %>%
    distinct()

  # pulling peak age and value
  infl <- NULL
  infl$logic <- c(NA, diff(diff(fit2$fit)) < 0, NA)
  infl <- as.data.frame(infl)

  # finding age of the first point significantly different from peak for the lower range
  # setting to lower bound if no values are present
  peak <- fit2 %>%
    bind_cols(infl) %>%
    filter(infl$logic == "TRUE") %>%
    group_by(per_age_wk, n, sd, se.fit) %>%
    summarise(
      fit = max(fit)
    ) %>%
    ungroup() %>%
    filter(fit == max(fit, na.rm = T))

  peak <- peak[1, ]

  # finding age of the first point significantly different from peak for the lower range
  # setting to lower bound if no values are present
  fit_diff_min <- tryCatch(
    fit2 %>%
      filter(per_age_wk < peak$per_age_wk) %>%
      mutate(fit_diff = fit - peak$fit) %>%
      mutate(age_diff = peak$per_age_wk - per_age_wk) %>%
      mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>%
      filter(pval <= 0.05) %>%
      filter(
        age_diff == min(age_diff)
      ) %>%
      distinct() %>%
      select(per_age_wk, fit),
    error = function(e) low_bound_peak
  )

  # finding age of the first point significantly different from peak for the higher range
  # setting to higher bound if no values are present
  fit_diff_max <- tryCatch(
    fit2 %>%
      filter(per_age_wk > peak$per_age_wk) %>%
      mutate(fit_diff = fit - peak$fit) %>%
      mutate(age_diff = per_age_wk - peak$per_age_wk) %>%
      mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>%
      filter(pval <= 0.05) %>%
      filter(
        age_diff == min(age_diff)
      ) %>%
      distinct() %>%
      select(per_age_wk, fit),
    error = function(e) high_bound_peak
  )

  # setting to lower bound if no values are present
  if (dim(fit_diff_min)[1] == 0) {
    fit_diff_min <- low_bound_peak
  } else {
  }

  # setting to higher bound if no values are present
  if (dim(fit_diff_max)[1] == 0) {
    fit_diff_max <- high_bound_peak
  } else {
  }


  if (!is.na(scale)) {
    # find rate of change prepeak using linear model
    prepeak_data <- fit2 %>%
      filter(per_age_wk <= peak$per_age_wk)
    # filter(per_age_wk <= peak$per_age_wk)


    summary_prepeak <- summary(lm(prepeak_data$fit ~ prepeak_data$per_age_wk))

    pre_est <- summary_prepeak$coefficients[, "Estimate"][2]
    pre_ste <- summary_prepeak$coefficients[, "Std. Error"][2]
    pre_pval <- format(round(summary_prepeak$coefficients[, "Pr(>|t|)"][2]), 3, nsmall = 3)

    pre_lcl <- pre_est - 1.96 * pre_ste
    pre_ucl <- pre_est + 1.96 * pre_ste

    pre_est_scale <- pre_est * scale
    pre_lcl_scale <- (pre_lcl * scale) - (1.96 * pre_ste * scale)
    pre_ucl_scale <- (pre_ucl * scale) + (1.96 * pre_ste * scale)

    prepeak_rate <- paste0(
      format(round(pre_est_scale, 2), nsmall = 2), " (",
      format(round(pre_lcl_scale, 2), nsmall = 2), ", ",
      format(round(pre_ucl_scale, 2), nsmall = 2), ")"
    )

    # find rate of change postpeak using linear model
    postpeak_data <- fit2 %>%
      filter(per_age_wk > peak$per_age_wk)

    summary_postpeak <- summary(lm(postpeak_data$fit ~ postpeak_data$per_age_wk))

    post_est <- summary_postpeak$coefficients[, "Estimate"][2]
    post_ste <- summary_postpeak$coefficients[, "Std. Error"][2]
    post_pval <- format(round(summary_postpeak$coefficients[, "Pr(>|t|)"][2]), 3, nsmall = 3)

    post_lcl <- post_est - 1.96 * post_ste
    post_ucl <- post_est + 1.96 * post_ste

    post_est_scale <- post_est * scale
    post_lcl_scale <- (post_lcl * scale) - (1.96 * post_ste * scale)
    post_ucl_scale <- (post_ucl * scale) + (1.96 * post_ste * scale)

    postpeak_rate <- paste0(
      format(round(post_est_scale, 2), nsmall = 2), " (",
      format(round(post_lcl_scale, 2), nsmall = 2), ", ",
      format(round(post_ucl_scale, 2), nsmall = 2), ")"
    )

    # putting all outputs into list
    all_peaks <- list(peak, fit_diff_min, fit_diff_max, high_bound_peak, low_bound_peak, prepeak_rate, postpeak_rate, pre_pval, post_pval)
    names(all_peaks) <- c("all", "min", "max", "highbound", "lowbound", "prerate", "postrate", "pre_pval", "post_pval")
  } else {
    # putting all outputs into list
    all_peaks <- list(peak, fit_diff_min, fit_diff_max, high_bound_peak, low_bound_peak)
    names(all_peaks) <- c("all", "min", "max", "highbound", "lowbound")
  }

  return(all_peaks)
}

gluc_peak <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("M", "F")), mle = life_exp$life_exp_wk, scale = scaling)
bw_peak <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("M", "F")), mle = life_exp$life_exp_wk, scale = scaling)
fat_peak <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("M", "F")), mle = life_exp$life_exp_wk, scale = scaling)

gluc_peak_male <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("M")), mle = life_exp_m$life_exp_wk, scale = scaling_m)
bw_peak_male <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("M")), mle = life_exp_m$life_exp_wk, scale = scaling_m)
fat_peak_male <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("M")), mle = life_exp_m$life_exp_wk, scale = scaling_m)

gluc_peak_female <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("F")), mle = life_exp_f$life_exp_wk, scale = scaling_f)
bw_peak_female <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("F")), mle = life_exp_f$life_exp_wk, scale = scaling_f)
fat_peak_female <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("F")), mle = life_exp_f$life_exp_wk, scale = scaling_f)

gluc_peak_HET3 <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(strain %in% c("HET3")), mle = life_exp_het3$life_exp_wk, scale = NA)
bw_peak_HET3 <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(strain %in% c("HET3")), mle = life_exp_het3$life_exp_wk, scale = NA)
fat_peak_HET3 <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(strain %in% c("HET3")), mle = life_exp_het3$life_exp_wk, scale = NA)

gluc_peak_B6 <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(strain %in% c("B6")), mle = life_exp_b6$life_exp_wk, scale = NA)
bw_peak_B6 <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(strain %in% c("B6")), mle = life_exp_b6$life_exp_wk, scale = NA)
fat_peak_B6 <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(strain %in% c("B6")), mle = life_exp_b6$life_exp_wk, scale = NA)
```

#### main_cat2 creation 
- merge 
  - main_all2 
  - census_c16_death 

#### imputation 
- carry forward previous observed non-missing value 
  - filter out all missing values 
  - filter for rows less than than 25 + 1, 50 + 1, 75 + 1, and 90 + 1 (previous observed values)
  - identify last non-missing value for imputation if needed 
  - leave missing if no last non-missing value 
  - merge into main_cat2 dataset
  
```{r slam_main_cat, warning=TRUE, message=FALSE, echo=FALSE}
census_c16_death$le_wk <- as.numeric(census_c16_death$le_wk)
main_cat2 <- main_all2 %>%
  left_join(census_c16_death) %>%
  select(idno) %>%
  distinct()

carry_forward <- function(index, cat) {
  index_cat <- paste0(index, "_", cat)

  output <- main_all2 %>%
    left_join(census_c16_death) %>%
    group_by(idno, sex) %>%
    filter(!is.na(!!index)) %>%
    # get all previous values to carry forward if needed
    filter(
      dplyr::case_when(
        sex == "F" ~ age_wk < ((!!cat / 100) * life_exp_f$life_exp_wk + 1),
        sex == "M" ~ age_wk < ((!!cat / 100) * life_exp_m$life_exp_wk + 1)
      )
    ) %>%
    mutate(!!index_cat := !!index) %>%
    # choose closest value
    filter(per_age_wk == max(per_age_wk)) %>%
    # filter for those alive at timepoint
    filter(
      dplyr::case_when(
        sex == "F" ~ le_wk >= ((!!cat / 100) * life_exp_f$life_exp_wk),
        sex == "M" ~ le_wk >= ((!!cat / 100) * life_exp_m$life_exp_wk)
      )
    ) %>%
    # mean of data at same timepoint
    summarise(
      !!index_cat := mean(!!index, na.rm = T)
    ) %>%
    select(idno, !!index_cat) %>%
    distinct() %>%
    right_join(main_cat2, by = "idno")

  return(output)
}
# main_cat2 <- carry_forward(quote(gluc), quote(25))
main_cat2 <- carry_forward(quote(gluc), quote(50))
main_cat2 <- carry_forward(quote(gluc), quote(75))
main_cat2 <- carry_forward(quote(gluc), quote(90))

# main_cat2 <- carry_forward(quote(bw), quote(25))
main_cat2 <- carry_forward(quote(bw), quote(50))
main_cat2 <- carry_forward(quote(bw), quote(75))
main_cat2 <- carry_forward(quote(bw), quote(90))

# main_cat2 <- carry_forward(quote(fat), quote(25))
main_cat2 <- carry_forward(quote(fat), quote(50))
main_cat2 <- carry_forward(quote(fat), quote(75))
main_cat2 <- carry_forward(quote(fat), quote(90))

# age adjustment
# main_cat2 <- carry_forward(quote(per_age_wk), quote(25))
main_cat2 <- carry_forward(quote(per_age_wk), quote(50))
main_cat2 <- carry_forward(quote(per_age_wk), quote(75))
main_cat2 <- carry_forward(quote(per_age_wk), quote(90))

# age_wk
main_cat2 <- carry_forward(quote(age_wk), quote(50))
main_cat2 <- carry_forward(quote(age_wk), quote(75))
main_cat2 <- carry_forward(quote(age_wk), quote(90))
```

#### main_all3 creation
- merge
  - main_all2
  - census_c16_death
  - main_cat2
- create variables for imputation and survival bias analysis

```{r slam_main_all3, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
main_all3 <- main_all2 %>%
  left_join(census_c16_death) %>%
  left_join(main_cat2) %>%
  mutate(
    whendead =

    # if(sex == "F")
    # as.factor(
    #           ifelse(per_le_wk >= 0*life_exp_f$life_exp_wk & per_le_wk < 0.375*life_exp_f$life_exp_wk, 25,
    #           ifelse(per_le_wk >= 0.375*life_exp_f$life_exp_wk & per_le_wk < 0.625*life_exp_f$life_exp_wk, 50,
    #           ifelse(per_le_wk >= 0.625*life_exp_f$life_exp_wk & per_le_wk < 0.875*life_exp_f$life_exp_wk, 75,
    #           ifelse(per_le_wk >= 0.875*life_exp_f$life_exp_wk, 90, NA)))))
    # else
    # as.factor(
    #           ifelse(per_le_wk >= 0*life_exp_m$life_exp_wk & per_le_wk < 0.375*life_exp_m$life_exp_wk, 25,
    #           ifelse(per_le_wk >= 0.375*life_exp_m$life_exp_wk & per_le_wk < 0.625*life_exp_m$life_exp_wk, 50,
    #           ifelse(per_le_wk >= 0.625*life_exp_m$life_exp_wk & per_le_wk < 0.875*life_exp_m$life_exp_wk, 75,
    #           ifelse(per_le_wk >= 0.875*life_exp_m$life_exp_wk, 90, NA))))),
      as.factor(
        ifelse(per_le_wk >= 0 & per_le_wk < 37.5, 25,
          ifelse(per_le_wk >= 37.5 & per_le_wk < 62.5, 50,
            ifelse(per_le_wk >= 62.5 & per_le_wk < 87.5, 75,
              ifelse(per_le_wk >= 87.5, 90, NA)
            )
          )
        )
      ),
    whenvisit =

    # if(sex == "F")
    # as.factor(
    #           ifelse(per_age_wk >= 0*life_exp_f$life_exp_wk & per_age_wk < 0.375*life_exp_f$life_exp_wk, 25,
    #           ifelse(per_age_wk >= 0.375*life_exp_f$life_exp_wk & per_age_wk < 0.625*life_exp_f$life_exp_wk, 50,
    #           ifelse(per_age_wk >= 0.625*life_exp_f$life_exp_wk & per_age_wk < 0.875*life_exp_f$life_exp_wk, 75,
    #           ifelse(per_age_wk >= 0.875*life_exp_f$life_exp_wk, 90, NA)))))
    # else
    # as.factor(
    #           ifelse(per_age_wk >= 0*life_exp_m$life_exp_wk & per_age_wk < 0.375*life_exp_m$life_exp_wk, 25,
    #           ifelse(per_age_wk >= 0.375*life_exp_m$life_exp_wk & per_age_wk < 0.625*life_exp_m$life_exp_wk, 50,
    #           ifelse(per_age_wk >= 0.625*life_exp_m$life_exp_wk & per_age_wk < 0.875*life_exp_m$life_exp_wk, 75,
    #           ifelse(per_age_wk >= 0.875*life_exp_m$life_exp_wk, 90, NA))))),
      as.factor(
        ifelse(per_age_wk >= 0 & per_age_wk < 37.5, 25,
          ifelse(per_age_wk >= 37.5 & per_age_wk < 62.5, 50,
            ifelse(per_age_wk >= 62.5 & per_age_wk < 87.5, 75,
              ifelse(per_age_wk >= 87.5, 90, NA)
            )
          )
        )
      ),
  )
```

## Results 
#### distribution figure with trend line
- generalized additive model for main indices
- linear model with quadratic fit for lifespan groups (survival bias analyses)
- remove timepoints before max life expectancy period 

```{r slam_ss_dist_funct, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
surv_color <- c("green", "yellow", "red", "purple", "gray")

fcn_plotdist_smooth <- function(metric, x, y, by, xlabel, ylabel, title, peak1, peak2) {
  # subsetting
  metric2 <- paste0("!is.na(", metric, ")")
  metric3 <- paste0("df$", metric)
  main_all3 <- main_all3 %>%
    filter(per_age_wk < as.numeric(all_asmt$le_max[1]))

  df <- base::subset(main_all3, eval(parse(text = metric2)))
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)")

  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) +

    # dots
    geom_point(alpha = 0.5, position = "jitter", size = 1, aes_string(colour = by), shape = 1) +

    # trendlines
    geom_smooth(aes_string(group = by, linetype = "solid"), linetype = "solid", color = "black", size = 1.8) +
    geom_smooth(aes(linetype = "solid")) +

    # peaks
    geom_point(aes(x = peak1$all$per_age_wk, y = peak1$all$fit), size = 2, color = "black", fill = "white", alpha = 0.7, shape = 23, stroke = 2) +
    geom_point(aes(x = peak2$all$per_age_wk, y = peak2$all$fit), size = 2, color = "black", fill = "white", alpha = 0.7, shape = 23, stroke = 2) +

    # labels
    xlab(xlabel) +
    ylab(ylabel) +
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE)
}

fcn_plotdist_lm <- function(metric, x, y, by, xlabel, ylabel, title) {
  # subsetting
  metric2 <- paste0("!is.na(", metric, ")")
  metric3 <- paste0("df$", metric)
  main_all3 <- main_all3 %>%
    filter(as.numeric(whenvisit) <= as.numeric(whendead)) %>%
    filter(as.numeric(whendead) > 0)
  df <- base::subset(main_all3, eval(parse(text = metric2)))
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)")
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) +
    # trendlines
    geom_smooth(aes_string(group = by, linetype = "solid"), method = "loess", linetype = "solid", color = "black", size = 1.8, se = FALSE) +
    geom_smooth(aes(linetype = "solid"), method = "loess", size = 1, se = FALSE) +

    # labels
    xlab(xlabel) +
    ylab(ylabel) +
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE) +
    scale_color_manual(values = surv_color, name = "Survival\n(% Life Expectancy)")
}
``` 

```{r slam_ss_dist_plots, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
cbPalette_dot <- c("#f8766d", "#00b0f6", "black")

# sex
slam_fig_i1 <- fcn_plotdist_smooth(
  metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "gluc", ylabel = "Glucose (mg/dL)", by = "sex", peak1 = gluc_peak_male, peak2 = gluc_peak_female
) +
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
slam_fig_i2 <- fcn_plotdist_smooth(
  metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "bw", ylabel = "Body Weight (g)", by = "sex", peak1 = bw_peak_male, peak2 = bw_peak_female
) +
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
slam_fig_i3 <- fcn_plotdist_smooth(
  metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "fat", ylabel = "Body Fat (g)", by = "sex", peak1 = fat_peak_male, peak2 = fat_peak_female
) +
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)

# strain
slam_fig_j1 <- fcn_plotdist_smooth(
  metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "gluc", ylabel = "Glucose (mg/dL)", by = "strain", peak1 = gluc_peak_HET3, peak2 = gluc_peak_B6
) +
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)
slam_fig_j2 <- fcn_plotdist_smooth(
  metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "bw", ylabel = "Body Weight (g)", by = "strain", peak1 = bw_peak_HET3, peak2 = bw_peak_B6
) +
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)
slam_fig_j3 <- fcn_plotdist_smooth(
  metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "fat", ylabel = "Body Fat (g)", by = "strain", peak1 = fat_peak_HET3, peak2 = fat_peak_B6
) +
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)

# death
slam_fig_d1 <- fcn_plotdist_lm(
  metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "gluc", ylabel = "Glucose (mg/dL)", by = "whendead"
)
slam_fig_d2 <- fcn_plotdist_lm(
  metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "bw", ylabel = "Body Weight (g)", by = "whendead"
)
slam_fig_d3 <- fcn_plotdist_lm(
  metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy",
  y = "fat", ylabel = "Body Fat (g)", by = "whendead"
)

#
slam_fig_i13 <- ggarrange(slam_fig_i1, slam_fig_i2, slam_fig_i3,
  ncol = 2, nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

slam_fig_j13 <- ggarrange(slam_fig_j1, slam_fig_j2, slam_fig_j3,
  ncol = 2, nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

slam_fig_d13 <- ggarrange(slam_fig_d1, slam_fig_d2, slam_fig_d3,
  ncol = 2, nrow = 2,
  common.legend = TRUE,
  legend = "right"
)
``` 

```{r slam_sex_dist_fig, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
annotate_figure(slam_fig_i13,
  top = text_grob("Trends in Metabolic Indices over the Lifespan by Sex", size = 12, hjust = 0, x = unit(0, "lines")),
  bottom = text_grob("Note: Diamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines"))
)
``` 
 
*** 

```{r slam_strain_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}
annotate_figure(slam_fig_j13,
  top = text_grob("Trends in Metabolic Indices over the Lifespan by Strain", size = 12, hjust = 0, x = unit(0, "lines")),
  bottom = text_grob("Note: Diamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines"))
)
``` 

*** 

#### pre and post-peak change creation 

```{r slam_phases, warning=FALSE, message=FALSE, echo=FALSE}
# all
gluc_peak_ci <- paste0(format(round(gluc_peak$all$per_age_wk, 2), nsmall = 2), " (", format(round(gluc_peak$min$per_age_wk, 2), nsmall = 2), ", ", format(round(gluc_peak$max$per_age_wk, 2), nsmall = 2), ")")

bw_peak_ci <- paste0(format(round(bw_peak$all$per_age_wk, 2), nsmall = 2), " (", format(round(bw_peak$min$per_age_wk, 2), nsmall = 2), ", ", format(round(bw_peak$max$per_age_wk, 2), nsmall = 2), ")")

fat_peak_ci <- paste0(format(round(fat_peak$all$per_age_wk, 2), nsmall = 2), " (", format(round(fat_peak$min$per_age_wk, 2), nsmall = 2), ", ", format(round(fat_peak$max$per_age_wk, 2), nsmall = 2), ")")

# male
male_gluc_peak_ci <- paste0(format(round(gluc_peak_male$all$per_age_wk, 2), nsmall = 2), " (", format(round(gluc_peak_male$min$per_age_wk, 2), nsmall = 2), ", ", format(round(gluc_peak_male$max$per_age_wk, 2), nsmall = 2), ")")

male_bw_peak_ci <- paste0(format(round(bw_peak_male$all$per_age_wk, 2), nsmall = 2), " (", format(round(bw_peak_male$min$per_age_wk, 2), nsmall = 2), ", ", format(round(bw_peak_male$max$per_age_wk, 2), nsmall = 2), ")")

male_fat_peak_ci <- paste0(format(round(fat_peak_male$all$per_age_wk, 2), nsmall = 2), " (", format(round(fat_peak_male$min$per_age_wk, 2), nsmall = 2), ", ", format(round(fat_peak_male$max$per_age_wk, 2), nsmall = 2), ")")

# female
female_gluc_peak_ci <- paste0(format(round(gluc_peak_female$all$per_age_wk, 2), nsmall = 2), " (", format(round(gluc_peak_female$min$per_age_wk, 2), nsmall = 2), ", ", format(round(gluc_peak_female$max$per_age_wk, 2), nsmall = 2), ")")

female_bw_peak_ci <- paste0(format(round(bw_peak_female$all$per_age_wk, 2), nsmall = 2), " (", format(round(bw_peak_female$min$per_age_wk, 2), nsmall = 2), ", ", format(round(bw_peak_female$max$per_age_wk, 2), nsmall = 2), ")")

female_fat_peak_ci <- paste0(format(round(fat_peak_female$all$per_age_wk, 2), nsmall = 2), " (", format(round(fat_peak_female$min$per_age_wk, 2), nsmall = 2), ", ", format(round(fat_peak_female$max$per_age_wk, 2), nsmall = 2), ")")

# combine all indices
roc_gluc <- c(
  "gluc", gluc_peak_ci, female_gluc_peak_ci, male_gluc_peak_ci,
  gluc_peak$prerate, gluc_peak_female$prerate, gluc_peak_male$prerate,
  gluc_peak$postrate, gluc_peak_female$postrate, gluc_peak_male$postrate
)

roc_bw <- c(
  "bw", bw_peak_ci, female_bw_peak_ci, male_bw_peak_ci,
  bw_peak$prerate, bw_peak_female$prerate, bw_peak_male$prerate,
  bw_peak$postrate, bw_peak_female$postrate, bw_peak_male$postrate
)

roc_fat <- c(
  "fat", fat_peak_ci, female_fat_peak_ci, male_fat_peak_ci,
  fat_peak$prerate, fat_peak_female$prerate, fat_peak_male$prerate,
  fat_peak$postrate, fat_peak_female$postrate, fat_peak_male$postrate
)

phases <- rbind(roc_gluc, roc_bw, roc_fat) %>%
  as.data.frame()

colnames(phases) <- c("index", "peak_all", "peak_female", "peak_male", "all", "female", "male", "all2", "female2", "male2")
``` 

```{r slam_phase_formatting, warning=FALSE, message=FALSE, echo=FALSE}
# html output formatting
phases2 <- phases %>%
  mutate(
    peak_all = noquote((gsub("\\(", "<br>(", peak_all))),
    peak_female = noquote((gsub("\\(", "<br>(", peak_female))),
    peak_male = noquote((gsub("\\(", "<br>(", peak_male))),
    all = noquote((gsub("\\(", "%<br>(", all))),
    female = noquote((gsub("\\(", "%<br>(", female))),
    male = noquote((gsub("\\(", "%<br>(", male))),
    all2 = noquote((gsub("\\(", "%<br>(", all2))),
    female2 = noquote((gsub("\\(", "%<br>(", female2))),
    male2 = noquote((gsub("\\(", "%<br>(", male2)))
  )

slam_phases <- phases2
```

#### change in metabolic indices pre and post-peak 
- rate of percent change 
 
```{r slam_phase_table_prate, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}
slam_phases_out <- kable(slam_phases,
  row.names = FALSE, col.names = c("Metabolic <br> Index", "All", "Female", "Male", "All", "Female", "Male", "All", "Female", "Male"), align = c("c", "c", "c", "c", "c", "c", "c"), escape = FALSE,
  caption = "Rate of Percent Change for Metabolic Indices Pre- and Post-peak"
) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE, position = "left", font_size = 12) %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%
  column_spec(4, border_right = TRUE) %>%
  column_spec(7, border_right = TRUE) %>%
  add_header_above(c(" " = 1, "Age at Peak Value$^†$ \n" = 3, "Pre-Peak$^‡$$^§$" = 3, "Post-Peak$^‡$$^§$" = 3)) %>%
  footnote(symbol = c(
    "p value < 0.05",
    "Mean (range of ages at similar values) age at peak index value",
    "Mean (95% confidence interval) rate of percent change",
    "Pre-peak is defined as 25% life expectancy to age at peak index value; post-peak phase is defined as peak to 100% life expectancy",
    "Male and female estimates are based on sex specific life expectancy"
  ))

slam_phases_out
```

*** 

#### quartile creation

```{r slam_scaling, warning=FALSE, message=FALSE, echo=FALSE}
main_cont_surv <- main_all3 %>%
  select(idno, sex, strain, per_age_wk_50, per_age_wk_75, per_age_wk_90) %>%
  distinct() %>%
  inner_join(census_c16_death)

main_cat_surv_f <- main_cat2 %>%
  inner_join(main_cont_surv) %>%
  filter(sex == "F")
main_cat_surv_m <- main_cat2 %>%
  inner_join(main_cont_surv) %>%
  filter(sex == "M")
```
 
```{r slam_quartiles, warning=FALSE, message=FALSE, echo=FALSE}
quart_breaks <- c(0, 0.25, 0.50, 0.75, 1.0)
quart_labs <- c("Low", "Medlow", "Medhigh", "High")

## female
main_cat_surv_f$gluc_50_quart <- cut(main_cat_surv_f$gluc_50, quantile(main_cat_surv_f$gluc_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$gluc_75_quart <- cut(main_cat_surv_f$gluc_75, quantile(main_cat_surv_f$gluc_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$gluc_90_quart <- cut(main_cat_surv_f$gluc_90, quantile(main_cat_surv_f$gluc_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

main_cat_surv_f$bw_50_quart <- cut(main_cat_surv_f$bw_50, quantile(main_cat_surv_f$bw_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$bw_75_quart <- cut(main_cat_surv_f$bw_75, quantile(main_cat_surv_f$bw_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$bw_90_quart <- cut(main_cat_surv_f$bw_90, quantile(main_cat_surv_f$bw_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

main_cat_surv_f$fat_50_quart <- cut(main_cat_surv_f$fat_50, quantile(main_cat_surv_f$fat_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$fat_75_quart <- cut(main_cat_surv_f$fat_75, quantile(main_cat_surv_f$fat_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_f$fat_90_quart <- cut(main_cat_surv_f$fat_90, quantile(main_cat_surv_f$fat_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

## male
main_cat_surv_m$gluc_50_quart <- cut(main_cat_surv_m$gluc_50, quantile(main_cat_surv_m$gluc_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$gluc_75_quart <- cut(main_cat_surv_m$gluc_75, quantile(main_cat_surv_m$gluc_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$gluc_90_quart <- cut(main_cat_surv_m$gluc_90, quantile(main_cat_surv_m$gluc_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

main_cat_surv_m$bw_50_quart <- cut(main_cat_surv_m$bw_50, quantile(main_cat_surv_m$bw_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$bw_75_quart <- cut(main_cat_surv_m$bw_75, quantile(main_cat_surv_m$bw_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$bw_90_quart <- cut(main_cat_surv_m$bw_90, quantile(main_cat_surv_m$bw_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

main_cat_surv_m$fat_50_quart <- cut(main_cat_surv_m$fat_50, quantile(main_cat_surv_m$fat_50, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$fat_75_quart <- cut(main_cat_surv_m$fat_75, quantile(main_cat_surv_m$fat_75, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)
main_cat_surv_m$fat_90_quart <- cut(main_cat_surv_m$fat_90, quantile(main_cat_surv_m$fat_90, breaks = quart_breaks, na.rm = T), labels = quart_labs, include.lowest = T)

# merge into all dataset
main_cat_surv <- main_cat_surv_f %>%
  rbind(main_cat_surv_m)

slam_main_cat_surv <- main_cat_surv
```

```{r save main_all2}
save(main_all2, file = "output/main_all2.RDATA")
save(main_cat_surv, file = "output/main_cat_surv.RDATA")
```

