---
title: "SLAM Metabolic Class Primary Figures"
author: "..."
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
geometry:
  - top=1in
  - bottom=1in
  - left=0.5in
  - right=0.5in
---

```{r figs setup, warning=FALSE, message=FALSE, echo=FALSE}
# library(helphlme)
library(tidyverse)
library(kableExtra)
# library(survival)
# library(survminer)
library(ggplotify)
# library(sjPlot)
# library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
library(flextable)
# library(future)
library(magick)
library(gridExtra)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```


```{r figs read input yaml}
# this loads in all the tables
all_path <- "../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/workspace.RDATA"
fb6_path <- "../pipeline/07_display_figures/output/slam_c1-c10_age_fb6_bwfatgluc/workspace.RDATA"
fhet3_path <- "../pipeline/07_display_figures/output/slam_c1-c10_age_fhet3_bwfatgluc/workspace.RDATA"
mb6_path <- "../pipeline/07_display_figures/output/slam_c1-c10_age_mb6_bwfatgluc/workspace.RDATA"
mhet3_path <- "../pipeline/07_display_figures/output/slam_c1-c10_age_mhet3_bwfatgluc/workspace.RDATA"
itp_path <- "../pipeline/07_display_figures/output/itp_c10c11c13c16_age_controls_bw/workspace.RDATA"
held_out_path <- "../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/workspace.RDATA"

config <- NULL
```

```{r figs load models, include = FALSE}
# Define a named list of input paths and environment names
render_tasks <- list(
  all_env = all_path,
  fb6_env = fb6_path,
  fhet3_env = fhet3_path,
  mb6_env = mb6_path,
  mhet3_env = mhet3_path,
  itp_env = itp_path,
  held_out_env = held_out_path
)

lapply(names(render_tasks), function(env_name) {
  env <- new.env()
  input_path <- render_tasks[[env_name]]
  load(input_path, envir = env)
  assign(env_name, env, envir = .GlobalEnv)
})
```

```{r figs fig no }
fig_no <- 1
tab_no <- 1
max_width <- 7
```

```{r save figure function}
save_figure <- function(image_obj, fig_num, label) {
  dir_path <- file.path("output", "primary", paste0("fig", fig_num))
  file_path <- file.path(dir_path, paste0(fig_num, label, ".png"))

  if (!dir.exists(dir_path)) {
    dir.create(dir_path, recursive = TRUE)
  }

  magick::image_write(image_obj, path = file_path, format = "png")
}
```

```{r annotate figure function}
annotate_figure_label <- function(image_obj, label) {
  magick::image_annotate(
    image_obj,
    text = toupper(label),
    size = 50,
    gravity = "northwest",
    location = "+20+20",
    color = "black"
  )
}
```

```{r grob to magick}
grob_to_magick <- function(grob, width = 400, height = 400, res = 150) {
  tf <- tempfile(fileext = ".png")
  png(tf, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(grob)
  dev.off()
  image_read(tf)
}
```

\newpage

\tableofcontents

\newpage

### Graphical Abstract

```{r figs overview, out.width = "100%"}
fig <- list()
fig$A <- image_read("images/Overview.png")
save_figure(fig$A, as.character(fig_no), "A")

fig$A
```

\newpage

### Figure `r as.character(fig_no)`: Defining Metabolic Classes

```{r define classes, out.width = "100%"}
fig <- list()
fig$A <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_1.png")
fig$B <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_4.png")
fig$C <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_7.png")
fig$D <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_10.png")
fig$E <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_2.png")
fig$F <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_5.png")
fig$G <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_8.png")
fig$H <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_11.png")
fig$I <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_3.png")
fig$J <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_6.png")
fig$K <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_9.png")
fig$L <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/outcome/plot_12.png")
fig$M <- image_read("images/Basic LCM.png")


for (label in names(fig)) {
  save_figure(fig[[label]], as.character(fig_no), label) # Save with correct label
  fig[[label]] <- annotate_figure_label(fig[[label]], label) # Annotate
}

# create table
hr_table <- do.call(rbind, all_env$save_figtabs$mortality_panel_hr)
hr_table <- cbind(Class = rownames(hr_table), hr_table)
hr_table <- flextable(hr_table) %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

invisible(save_as_image(hr_table, "raw/mortality_panel_hr.png", zoom = 10))
fig$N <- image_read("raw/mortality_panel_hr.png")
fig$N <- image_scale(fig$N, "400%")
fig$N <- image_border(fig$N, color = "white", geometry = "100x100")
fig$N <- annotate_figure_label(fig$N, "N")


legend1 <- grob_to_magick(all_env$save_figtabs$legends_only[[1]]) %>%
  image_scale("200%")
legend2 <- grob_to_magick(all_env$save_figtabs$legends_only[[2]]) %>%
  image_scale("200%")
legend3 <- grob_to_magick(all_env$save_figtabs$legends_only[[3]]) %>%
  image_scale("200%")


# Build rows of 4 images
row1 <- image_append(c(fig$A, fig$B, fig$C, fig$D, legend1), stack = FALSE)
row2 <- image_append(c(fig$E, fig$F, fig$G, fig$H, legend2), stack = FALSE)
row3 <- image_append(c(fig$I, fig$J, fig$K, fig$L, legend3), stack = FALSE)
row4 <- image_append(c(fig$M, fig$N))
# Stack the rows vertically
final_panel <- image_append(c(row1, row2, row3, row4), stack = TRUE)

# View or save
final_panel
# image_write(final_panel, "raw/fig/panel.png")

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Sex/Strain Classes BW

```{r sex strain bw classes, out.width = "100%"}
fig <- list()
fig$A <- image_read("raw/fb6_env/define_class/bwbw.png")
fig$B <- image_read("raw/fb6_env/define_class/bwkm.png")
fig$C <- image_read("raw/mb6_env/define_class/bwbw.png")
fig$D <- image_read("raw/mb6_env/define_class/bwkm.png")
fig$E <- image_read("raw/fhet3_env/define_class/bwbw.png")
fig$F <- image_read("raw/fhet3_env/define_class/bwkm.png")
fig$G <- image_read("raw/mhet3_env/define_class/bwbw.png")
fig$H <- image_read("raw/mhet3_env/define_class/bwkm.png")

for (label in names(fig)) {
  fig[[label]] <- annotate_figure_label(fig[[label]], label) # Annotate
}

cbind_hr_table <- function(strata, hr_table) {
  cbind(sex_strain = strata, Class = rownames(hr_table), hr_table)
}

hr_table <- rbind(
  cbind_hr_table("fb6", fb6_env$save_figtabs$mortality_panel_hr[[1]]),
  cbind_hr_table("mb6", mb6_env$save_figtabs$mortality_panel_hr[[1]]),
  cbind_hr_table("fhet3", fhet3_env$save_figtabs$mortality_panel_hr[[1]]),
  cbind_hr_table("mhet3", mhet3_env$save_figtabs$mortality_panel_hr[[1]])
)
rownames(hr_table) <- NULL

hr_table <- flextable(hr_table) %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

invisible(save_as_image(hr_table, "raw/mortality_panel_hr.png", zoom = 10))
fig$I <- image_read("raw/mortality_panel_hr.png")
fig$I <- image_scale(fig$I, "400%")
fig$I <- image_border(fig$I, color = "white", geometry = "100x100")
fig$I <- annotate_figure_label(fig$I, "I")

for (label in names(fig)) {
  save_figure(fig[[label]], as.character(fig_no), label) # Save with correct label
}

legend1 <- image_read("raw/fb6_env/define_class/bwbw_legend.png")
legend2 <- image_read("raw/mb6_env/define_class/bwbw_legend.png")
legend3 <- image_read("raw/fhet3_env/define_class/bwbw_legend.png")
legend4 <- image_read("raw/mhet3_env/define_class/bwbw_legend.png")

row1 <- image_append(c(fig$A, fig$B, legend1, fig$C, fig$D, legend2), stack = FALSE)
row2 <- image_append(c(fig$E, fig$F, legend3, fig$G, fig$H, legend4), stack = FALSE)

final_panel <- image_append(c(row1, row2, fig$I), stack = TRUE)

final_panel

fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Graded Mortality Response

```{r dose response, out.width = "100%"}
fig <- list()
fig$A <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/dose_response/plot_1.png")
fig$B <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/dose_response/plot_4.png")
fig$C <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/dose_response/plot_7.png")
fig$D <- image_read("images/Basic LCM.png")
fig$E <- image_read("images/Graded Mortality.png")


for (label in names(fig)) {
  save_figure(fig[[label]], as.character(fig_no), label) # Save with correct label
  fig[[label]] <- annotate_figure_label(fig[[label]], label) # Annotate
}

# HR table removed -- hazard ratios are now shown in the forest-plot figure
# below (Graded Mortality Response - Forest Plot)

# Stack the KM dose-response curves
col1 <- image_append(c(fig$A, fig$B, fig$C), stack = TRUE)

final_panel <- col1

# View or save
final_panel
# image_write(final_panel, "raw/fig2/panel.png")
fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Graded Mortality Response - Forest Plot

```{r dose response forest, out.width = "100%", fig.height=11, fig.width=12}
all_env$save_figtabs$dose_plot_panel_forest

fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Held Out Cohort Validation

```{r external validation, out.width = "100%"}
fig <- list()
fig$A <- image_read("images/External Validation.png")
fig$B <- image_read("../pipeline/07_display_figures/output/slam_c1-c10_age_all_bwfatgluc/forest/plot_5.png")
# fig$C <- image_read("../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/km_combined_hr_validation/plot_[6.5, 32.5]_2.png")
fig$C <- image_read("../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/km_combined_hr_validation/plot_[6.5, 58.5]_4.png")
fig$D <- image_read("../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/km_combined_hr_validation/plot_[6.5, 84.5]_6.png")
fig$E <- image_read("../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/km_combined_hr_validation/plot_[6.5, 110.5]_8.png")
fig$F <- image_read("../pipeline/07_display_figures/output/slam_c1-10_x_slam_c16-18_age_bwfatgluc/km_combined_hr_validation/plot_[6.5, 136.5]_10.png")

fig$A <- image_scale(fig$A, "75%")
fig$B <- image_scale(fig$B, "125%")

for (label in names(fig)) {
  save_figure(fig[[label]], as.character(fig_no), label) # Save with correct label
  fig[[label]] <- annotate_figure_label(fig[[label]], label) # Annotate
}

# HR table removed -- hazard ratios are now shown in the forest-plot figure
# below (Held Out Cohort Validation - Forest Plot)

# Stack the rows vertically
col1 <- image_append(c(fig$A, fig$B), stack = TRUE)
col2 <- image_append(c(fig$C, fig$E), stack = TRUE)
col3 <- image_append(c(fig$D, fig$F), stack = TRUE)

final_panel <- image_append(c(col1, col2, col3), stack = FALSE)

# View or save
final_panel
# image_write(final_panel, "raw/fig2/panel.png")
fig_no <- fig_no + 1
```

\newpage

### Figure `r as.character(fig_no)`: Held Out Cohort Validation - Forest Plot

```{r external validation forest, out.width = "100%", fig.height=13, fig.width=12}
held_out_env$save_figtabs$km_hr_combine_panel_forest

fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: ITP Validation

```{r ITP validation, out.width = "100%"}
fig <- list()
fig$A <- image_read("../pipeline/07_display_figures/output/itp_c10c11c13c16_age_controls_bw/outcome/plot_1.png")
fig$B <- image_read("../pipeline/07_display_figures/output/itp_c10c11c13c16_age_controls_bw/outcome/plot_2.png")
fig$D <- image_read("images/ITP Treatment.png")
fig$E <- image_read("../pipeline/99_treatment_response/output/treatment_response/plot_1.png")
fig$F <- image_read("../pipeline/99_treatment_response/output/treatment_response/plot_2.png")
fig$G <- image_read("../pipeline/99_treatment_response/output/treatment_response/plot_3.png")

for (label in names(fig)) {
  save_figure(fig[[label]], as.character(fig_no), label) # Save with correct label
  fig[[label]] <- annotate_figure_label(fig[[label]], label) # Annotate
}

load("../pipeline/99_treatment_response/output/hr_table/hr_table.RDATA")

hr_table <- do.call(rbind, hrs_table) %>%
  flextable() %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

invisible(save_as_image(hr_table, "raw/treatment_mortlity_panel_hr.png", zoom = 10))
fig$H <- image_read("raw/treatment_mortlity_panel_hr.png")
fig$H <- image_scale(fig$H, "400%")
fig$H <- image_border(fig$H, color = "white", geometry = "100x100")
fig$H <- annotate_figure_label(fig$H, "H")

# create table

hr_table <- itp_env$save_figtabs$mortality_panel_hr[[1]]
hr_table <- cbind(Class = rownames(hr_table), hr_table)
hr_table <- flextable(hr_table) %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

invisible(save_as_image(hr_table, "raw/mortality_panel_hr.png", zoom = 10))
fig$C <- image_read("raw/mortality_panel_hr.png")
fig$C <- image_scale(fig$C, "400%")
fig$C <- image_border(fig$C, color = "white", geometry = "100x100")
fig$C <- annotate_figure_label(fig$C, "C")


# Stack the rows vertically
row1col1 <- image_append(c(fig$A, fig$B), stack = FALSE)
row1col2 <- image_append(c(fig$C, fig$D), stack = TRUE)
row1 <- image_append(c(row1col1, row1col2), stack = FALSE)
row2 <- image_append(c(fig$E, fig$F, fig$G), stack = FALSE)
row3 <- fig$H

final_panel <- image_append(c(row1, row2, row3), stack = TRUE)

# View or save
final_panel
# image_write(final_panel, "raw/fig2/panel.png")
fig_no <- fig_no + 1
```

\newpage

### Table `r as.character(tab_no)`: SLAM C1-C10 Class Demographics

```{r slam class demographics, results = 'asis'}
all_env$save_figtabs$t1_df %>%
  select(-oc_name) %>%
  mutate(Class = row_names) %>%
  flextable() %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

tab_no <- 1 + tab_no
```

\newpage

### Table `r as.character(tab_no)`: ITP Class Demographics

```{r itp class demographics, results = 'asis'}
itp_env$save_figtabs$t1_df %>%
  select(-oc_name) %>%
  mutate(Class = row_names) %>%
  flextable() %>%
  theme_vanilla() %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(max_width = max_width, inc = .25, max_iter = 100)

tab_no <- 1 + tab_no
```

