---
title: "SLAM Metabolic Class Figures Display"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../05_figures/output/test_local.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```


## Analysis

### Load Input File

```{r read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  params$input_path <- "../05_figures/output/test_local.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
```

### Load Models

```{r load models}
load(input$datasets_path)
load(input$final_models_path)
config <- yaml::read_yaml(input$config_path)


# load in main_cat_surv
main_cat_surv <- read.csv(file = config$survival_dataset$path)
main_cat_surv <- as.data.frame(main_cat_surv)

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

## Create Figures

### Dispaly How latent class models work

```{r how they work, fig.height=5, fig.width=10}
p1 <- final_models[, "spag_plot"][[1]]
p3 <- final_models[, "traj_obs"][[1]]

fig1 <- cowplot::plot_grid(p1, p3, nrow = 1)
```

### Dispaly predicted and observed

- Supplement will include sex strain strata 

```{r pred_obs_panel, fig.height= 16, fig.width=12}
og_index <- final_models$data_mod == "og"
pred_plots <- final_models[og_index, "traj_pred"]
obs_plots <- final_models[og_index, "traj_obs"]

pred_plots <- lapply(pred_plots, ggdraw, clip = "on")

title_theme <- ggdraw() +
  draw_label("Predicted and Observed Metabolic Life Course Trajectories")

plot_list <- c(pred_plots, obs_plots)
pred_obs_panel <- cowplot::plot_grid(
  plotlist = plot_list, nrow = 3, byrow = FALSE,
  greedy = FALSE, labels = "AUTO"
)

pred_obs_panel <- cowplot::plot_grid(title_theme, pred_obs_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Table Who's in what class

```{r table class demographics, results='asis'}
t1_list <- final_models[og_index, "clean_t1"]
outcome_names <- final_models[og_index, "oc_name"]

t1_list <- lapply(seq_along(t1_list), function(i) {
  t1 <- t1_list[[i]]
  oc_name <- outcome_names[[i]]

  if (isFALSE(all(is.na(t1[[1]])))) {
    t1 <- t1
    t1 <- as.data.frame(t(t1))
    t1$row_names <- row.names(t1)
    row.names(t1) <- NULL
    t1$oc_name <- oc_name
  } else {
    t1 <- NULL
  }
  t1
})
t1_df <- do.call(rbind, t1_list)
first_names <- match(c("oc_name", "row_names"), names(t1_df))
last_names <- seq_along(names(t1_df))[-first_names]

t1_df <- t1_df[, c(first_names, last_names)]
```


### Mortality outcomes by class

- Supplement will include same figures but for each sex strain strata

```{r mortality outcomes by class, fig.height= 16, fig.width=12}
kap_plots <- final_models[og_index, "kap_plot"]
kap_plots <- lapply(kap_plots, `[[`, "plot")
plot_list <- c(obs_plots, kap_plots)

mortality_panel <- cowplot::plot_grid(
  plotlist = plot_list, nrow = 3,
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)
title_theme <- ggdraw() +
  draw_label(
    "Observed Metabolic Life Course Trajectories and Kaplan Meiers by Class"
  )

mortality_panel <- cowplot::plot_grid(title_theme, mortality_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Table of Hazard Ratios, Raw and Sex/Strain Adjusted

```{r hazard ratio table, results = 'asis'}
surv_gethr <- function(fit, vars, names = NULL, ndec = 4) {
  sum_fit <- summary(fit) ## get summary of fit
  # save coeffecients as dataframe
  coef <- data.frame(sum_fit$coefficients, check.names = F)
  # save confidence interval as dataframe
  conf <- data.frame(sum_fit$conf.int, check.names = F)
  # initialize hr dataframe
  hr <- data.frame()
  # loop through vars
  for (var in vars) {
    # save hr
    value <- format(round(conf[var, "exp(coef)"], ndec), nsmall = 2)
    # save lower HR bound
    lower <- format(round(conf[var, "lower .95"], ndec), nsmall = 2)
    # save upper HR bound
    upper <- format(round(conf[var, "upper .95"], ndec), nsmall = 2)
    # save pval
    pval <- coef[var, "Pr(>|z|)"]

    ## add stars depending on significance
    if (pval < .001) {
      final <- paste0(value, " (", lower, ", ", upper, ")***")
    } else if (pval < .005) {
      final <- paste0(value, " (", lower, ", ", upper, ")**")
    } else if (pval < 0.050) {
      final <- paste0(value, " (", lower, ", ", upper, ")*")
    } else {
      final <- paste0(value, " (", lower, ", ", upper, ")")
    }

    ## save stats above to hr dataframe
    hr[var, c("value", "lower", "upper", "pval", "final")] <- c(
      value,
      lower,
      upper,
      pval,
      final
    )
  }

  ## if names are provided, rename variables
  if (!is.null(names)) {
    rownames(hr) <- names
  }

  ## create a column that has the rownames printed before HR
  hr[, "with_names"] <- unlist(lapply(rownames(hr), function(name) {
    el <- paste(name, hr[name, "final"], sep = ": ")
  }))

  ## create text that can be used to easily add to KM plot in place of pval
  one_text <- paste(as.vector(hr[, "with_names"]), collapse = "\n")

  return(list(hr_table = hr, hr_text = one_text))
}


make_cox_table <- function(cox_models, mo_select = c(1, 3), outcome) {
  xlevels_list <- lapply(cox_models, `[[`, "xlevels")
  xlevels_names <- lapply(xlevels_list, names)

  xlevels_list <- lapply(seq_along(xlevels_names), function(i) {
    vars <- xlevels_list[[i]]
    names <- xlevels_names[[i]]
    new_names <- mapply(paste0, names, vars, SIMPLIFY = FALSE)
    new_names <- unlist(lapply(new_names, `[`, -1))
  })


  hr_table_list <- list()
  for (i in mo_select) {
    hr_table <- surv_gethr(cox_models[[i]],
      vars = xlevels_list[[i]]
    )$hr_table
    hr_table <- cbind(row.names(hr_table), hr_table)
    names(hr_table)[[1]] <- "Class"
    row.names(hr_table) <- NULL
    model_name <- paste("HR Model", i)
    hr_table <- hr_table %>%
      select(Class, hr = final)
    hr_table <- cbind(outcome, hr_table)
    names(hr_table) <- c("Outcome", "Class", model_name)
    hr_table_list[[i]] <- hr_table
  }
  result <- purrr::reduce(hr_table_list, dplyr::left_join, by = c("Class", "Outcome"))
}


combine_cox_table <- function(final_models, mo_names, cox_mo_select) {
  x <- final_models
  cox_table_list <- list()
  for (name in mo_names) {
    cox_models <- x[x$model_name == name, ]$cox_models[[1]]
    outcome <- x[x$model_name == name, ]$oc_name
    ng <- x[x$model_name == name, ]$ng
    if (ng != 1) {
      cox_table_list[[name]] <- make_cox_table(
        cox_models,
        cox_mo_select,
        outcome
      )
    } else {
      cox_table_list[[name]] <- NA
    }
  }
  names(cox_table_list) <- NULL
  do.call(rbind, cox_table_list)
}

model_names <- final_models[og_index, "model_name"]
hr_table <- combine_cox_table(final_models, model_names, c(1, 2, 3))
```

### Metabolic Outcomes by Class

- Supplement will show the same figures but fore each sex strain combo

```{r metabolic outcomes by class, fig.height=16, fig.width=24}
og_plots_index <- grepl("^plot.*_og_.*across$", names(final_models))

og_plots <- unlist(final_models[og_index, og_plots_index], recursive = FALSE)

other_met_panel <- cowplot::plot_grid(
  plotlist = og_plots,
  nrow = length(which(og_index)),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)


title_theme <- ggdraw() +
  draw_label(paste(
    "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
    sep = " "
  ))

other_met_panel <- cowplot::plot_grid(title_theme, other_met_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Table of lme models
- Model1 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model2 : $outcome = Age + Age^2 + Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model3 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model4 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model5 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Female * HET3 + (Age + Age^2 | Subject)$
```{r table lme, results = 'asis'}
rows <- final_models[og_index, ]

age_var <- rows$age_var[[1]]
age_var_name <- rows$age_var_name[[1]]
age_var_units <- rows$age_var_units[[1]]

age_var2_name <- paste(age_var_name, "2</sup>", sep = "<sup>")
age_var2_units <- paste(age_var_units, "2</sup>", sep = "<sup>")

age_var2 <- paste(age_var, 2, sep = "")
age_var_ns <- paste(age_var, "ns", sep = "_")
age_var2_ns <- paste(age_var2, "ns", sep = "_")
class_names <- paste("Class", seq_len(100))
names(class_names) <- paste0("new_class", seq_len(100))
age_names <- paste(age_var_name, age_var_units)
names(age_names) <- age_var_ns
age2_names <- paste(age_var2_name, age_var2_units)
names(age2_names) <- age_var2_ns
age_interaction_names <- paste(age_names, class_names, sep = ":")
names(age_interaction_names) <- paste(names(age_names),
  names(class_names),
  sep = ":"
)
age2_interaction_names <- paste(age2_names, class_names, sep = ":")
names(age2_interaction_names) <- paste(names(age2_names), names(class_names), sep = ":")
fixcov_names <- c("female" = "Female", "het3" = "HET3")
fixcov_names_interacted <- paste(fixcov_names, collapse = ":")
names(fixcov_names_interacted) <- paste(names(fixcov_names), collapse = ":")

all_names <- c(
  class_names,
  age_names,
  age2_names,
  age_interaction_names,
  age2_interaction_names,
  fixcov_names,
  fixcov_names_interacted
)

lmes_index <- grepl("^model.*_og_.*across$", names(final_models))
lmes_list <- final_models[og_index, lmes_index]

model_list <- lapply(lmes_list, function(row) {
  lapply(row, `[[`, 4)
})

invert_list <- function(list) {
  list_matrix <- do.call(rbind, list)
  output <- apply(list_matrix, 2, as.list)
  output
}
model_list <- invert_list(model_list)



oc_names_list <- rep(list(rows$oc_name), times = length(lmes_list))

lme_table_list <- list()
title_list <- paste(rows$oc_name, "Class")

lme_table <- lapply(seq_along(model_list), function(i) {
  title <- paste(title_list[[i]], "Mixed Effects Models")
  lme_table <- tab_model(model_list[[i]],
    show.ci = FALSE,
    show.re.var = FALSE,
    pred.labels = all_names,
    show.reflvl = TRUE,
    wrap.labels = 50,
    show.icc = FALSE,
    show.r2 = FALSE,
    show.ngroups = FALSE,
    show.obs = FALSE,
    dv.labels = oc_names_list[[i]],
    title = title
  )
  lme_table
})

lme_table
```

### Class Prediction Panel

- Supplmenet will include same table for glucose and body fat

```{r class prediction panel, fig.height=16, fig.width=16}
mo_names <- c("bw_main|bw_age_wk", "gluc_main|gluc_age_wk", "fat_main|fat_age_wk")

interval_plots <- final_models[og_index, "interval_boot_accuracy_plot"]
window_plots <- final_models[og_index, "window_boot_accuracy_plot"]
cumulative_plots <- final_models[og_index, "cumulative_boot_accuracy_plot"]
resampled_plots <- final_models[og_index, "resampled_boot_accuracy_plot"]

plot_list <- c(interval_plots, window_plots, cumulative_plots, resampled_plots)

title <- paste("Classification Accuracy")

title_theme <- ggdraw() +
  draw_label(title)

class_prediction_panel <- cowplot::plot_grid(
  plotlist = plot_list,
  ncol = 4,
  labels = "AUTO",
  byrow = FALSE
)

class_prediction_panel <- cowplot::plot_grid(title_theme,
  class_prediction_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```


## Display Figures

### Figure 1:How it works

```{r figure 1}
fig1
```

### Figure 2: Predicted Observed
```{r figure 2, fig.height=18, fig.width=12}
pred_obs_panel
```

### Table 1: Whos in each Class
```{r table 1, results = 'asis'}
print_table <- function(env, table_name, title) {
  cap <- paste(env$subgroup,
    title,
    sep = " "
  )
  table <- kable(env[[table_name]],
    format = "html",
    caption = cap
  ) %>%
    column_spec(1:2, bold = TRUE, border_right = TRUE) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    kable_styling(bootstrap_options = "striped")

  return(table)
}

print_table(.GlobalEnv,
  table_name = "t1_df",
  title = "Demographics by Metabolic Index and Class"
)
```

### Figure 3: Metabolic Classes with Other Metabolic Metrics

```{r figure 3, fig.height=18, fig.width=24}
other_met_panel
```

### Table 2: Metabolic Classes Predict other Metabolic Metrics

```{r table 2, results = 'asis'}
for(table in lme_table) {
  cat(table$knitr)
}
```

### Figure 4: Mortality and Class

```{r figure 4, fig.height=18, fig.width=12}
mortality_panel
```

### Table 3: Hazard Ratios

- Model 1: $ Surv ~ Class $
- Model 2: $ Surv ~ Class + Sex + Strain$
- Model 3: $ Surv ~ Class + Sex * Strain$

```{r table 3, results = 'asis'}
print_table(.GlobalEnv,
  table_name = "hr_table",
  title = "Table of Hazard Ratios by Metabolic Indices"
)
```

### Figure 5: Class Prediction Panel

```{r figure 5, fig.width=18, fig.height=18}
class_prediction_panel
```

### Figure 6: Internal Validation (Preliminary)

```{r figure 6}
final_models[!og_index, ]$filter_window_test_lp_plot[[1]]
final_models[!og_index, ]$filter_window_test_concord[[1]]
final_models[!og_index, ]$filter_window_test_wave_lp_plot[[1]]
final_models[!og_index, ]$filter_window_test_wave_concord[[1]]
final_models[!og_index, ]$filter_cumulative_test_lp_plot[[1]]
final_models[!og_index, ]$filter_cumulative_test_concord[[1]]
final_models[!og_index, ]$filter_cumulative_test_wave_lp_plot[[1]]
final_models[!og_index, ]$filter_cumulative_test_wave_concord[[1]]
```
