---
title: "SLAM Metabolic Class Figures Display - `r config$out_tag` Mice"
subtitle: "External Validation: `r config$external_validate`"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../05_figures/output/test_local.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)

window_legend <- TRUE
window_legend_palette <- magma(3, begin = .2, end = .7, direction = -1)
window_legend_labels <- c("2 measures", "3 measures", "4 measures")

cumulative_legend <- TRUE
cumulative_legend_palette <- viridis(3, begin = .2, end = .7, direction = -1)
```

## Analysis

### Load Input File

```{r read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  # params$input_path <- "../05_figures/output/test_local.yaml"
  params$input_path <- "../05_figures/output/all.yaml"
  # params$input_path <- "../05_figures/output/all_lean.yaml"
  # params$input_path <- "../05_figures/output/fb6.yaml"
  # params$input_path <- "../05_figures/output/external_validation.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
```

### Load Models

```{r load models}
load(input$datasets_path)
load(input$final_models_path)
config <- yaml::read_yaml(input$config_path)


# load in main_cat_surv
main_cat_surv <- read.csv(file = config$survival_dataset$path)
main_cat_surv <- as.data.frame(main_cat_surv)

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

## Create Figures

### Dispaly How latent class models work

```{r how they work, fig.height=5, fig.width=10}
p1 <- final_models[, "spag_plot"][[1]]
p3 <- final_models[, "traj_obs"][[1]]

how_it_works <- cowplot::plot_grid(p1, p3, nrow = 1)
```

### Dispaly predicted and observed

- Supplement will include sex strain strata 

#### Complete Data

```{r pred_obs_panel, fig.height= 16, fig.width=12}
og_index <- final_models$data_mod == "og"
pred_plots <- final_models[og_index, "traj_pred"]
obs_plots <- final_models[og_index, "traj_obs"]

if (config$custom_palette) {
  j <- 0
  for (i in seq_along(obs_plots)) {
    group <- rlang::eval_tidy(obs_plots[[i]]$mapping$colour,
      data = obs_plots[[i]]$data
    )
    k <- length(unique(group)) + j
    obs_plots[[i]] <- obs_plots[[i]] +
      scale_color_manual(
        values = config$palette_colors[j + 1:k],
        name = "Class",
        labels = config$legend_labels[j + 1:k]
      )
    j <- k
  }
}

pred_plots <- lapply(pred_plots, ggdraw, clip = "on")

title_theme <- ggdraw() +
  draw_label("Predicted and Observed Metabolic Life Course Trajectories")

plot_list <- c(pred_plots, obs_plots)
pred_obs_panel <- cowplot::plot_grid(
  plotlist = plot_list, nrow = length(pred_plots), byrow = FALSE,
  greedy = FALSE, labels = "AUTO"
)

pred_obs_panel <- cowplot::plot_grid(title_theme, pred_obs_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

#### Training Data

```{r pred_obs_panel training, fig.height= 16, fig.width=12}
train_index <- final_models$data_mod == "train_test"
pred_plots_train <- final_models[train_index, "traj_pred"]
obs_plots_train <- final_models[train_index, "traj_obs"]

if (config$custom_palette_train) {
  j <- 0
  for (i in seq_along(obs_plots_train)) {
    group <- rlang::eval_tidy(obs_plots_train[[i]]$mapping$colour,
      data = obs_plots_train[[i]]$data
    )
    k <- length(unique(group)) + j
    obs_plots_train[[i]] <- obs_plots_train[[i]] +
      scale_color_manual(
        values = config$palette_colors_train[j + 1:k],
        name = "Class",
        labels = config$legend_labels_train[j + 1:k]
      )
    j <- k
  }
}

pred_plots_train <- lapply(pred_plots_train, ggdraw, clip = "on")

title_theme <- ggdraw() +
  draw_label("Predicted and Observed Metabolic Life Course Trajectories")

plot_list <- c(pred_plots_train, obs_plots_train)
pred_obs_panel_train <- cowplot::plot_grid(
  plotlist = plot_list, nrow = length(pred_plots_train), byrow = FALSE,
  greedy = FALSE, labels = "AUTO"
)

pred_obs_panel_train <- cowplot::plot_grid(title_theme, pred_obs_panel_train,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Table Who's in what class

#### Complete Data

```{r table class demographics, results='asis'}
t1_list <- final_models[og_index, "clean_t1"]
outcome_names <- final_models[og_index, "oc_name"]

t1_list <- lapply(seq_along(t1_list), function(i) {
  t1 <- t1_list[[i]]
  oc_name <- outcome_names[[i]]

  if (isFALSE(all(is.na(t1[[1]])))) {
    t1 <- t1
    t1 <- as.data.frame(t(t1))
    t1$row_names <- row.names(t1)
    row.names(t1) <- NULL
    t1$oc_name <- oc_name
  } else {
    t1 <- NULL
  }
  t1
})
t1_df <- do.call(rbind, t1_list)
first_names <- match(c("oc_name", "row_names"), names(t1_df))
last_names <- seq_along(names(t1_df))[-first_names]

t1_df <- t1_df[, c(first_names, last_names)]

if (config$custom_palette) {
  index <- grepl("^class_\\d+_final$", t1_df$row_names)
  t1_df$row_names[index] <- config$legend_labels
}
```

#### Training Data

```{r table class demographics train, results='asis'}
t1_list_train <- final_models[train_index, "clean_t1"]
outcome_names <- final_models[train_index, "oc_name"]

t1_list_train <- lapply(seq_along(t1_list_train), function(i) {
  t1 <- t1_list_train[[i]]
  oc_name <- outcome_names[[i]]

  if (isFALSE(all(is.na(t1[[1]])))) {
    t1 <- t1
    t1 <- as.data.frame(t(t1))
    t1$row_names <- row.names(t1)
    row.names(t1) <- NULL
    t1$oc_name <- oc_name
  } else {
    t1 <- NULL
  }
  t1
})
t1_df_train <- do.call(rbind, t1_list_train)
first_names <- match(c("oc_name", "row_names"), names(t1_df_train))
last_names <- seq_along(names(t1_df_train))[-first_names]

t1_df_train <- t1_df_train[, c(first_names, last_names)]

if (config$custom_palette_train) {
  index <- grepl("^class_\\d+_final$", t1_df_train$row_names)
  t1_df_train$row_names[index] <- config$legend_labels_train
}
```

### Heatmap census
```{r heatmap, fig.height = 8, fig.width = 8, include = FALSE}
census <- final_models$combined_census[[1]]
prob_cols <- names(census)[grepl("^prob\\d+", names(census))]
cov <- datasets[[1]]$covariates_dummy
census1 <- census[, c(prob_cols, cov)]
if (config$custom_palette) {
  colnames(census1) <- c(config$legend_labels, cov)
}
heatmap(as.matrix(census1), margins = c(10, 4))
heatmap1 <- recordPlot()
census2 <- census[, prob_cols]
if (config$custom_palette) {
  colnames(census2) <- config$legend_labels
}
heatmap(as.matrix(census2), margins = c(10, 4))
heatmap2 <- recordPlot()
```

```{r save heatamps}
getwd()
output_path <- file.path("output", config$out_tag, "heatmap")
dir.create(output_path, recursive = TRUE)
output_path <- normalizePath(output_path)

file_path <- file.path(output_path, "heatmap1.png")
jpeg(file = file_path)
heatmap(as.matrix(census1), margins = c(10, 4))
dev.off()

file_path <- file.path(output_path, "heatmap2.png")
jpeg(file = file_path)
heatmap(as.matrix(census2), margins = c(10, 4))
dev.off()
```

### Show validation data

```{r validation data}
plot_validation <- function(data1,
                            data2,
                            age_var,
                            outcome,
                            age_name,
                            outcome_name) {
  ggplot2::ggplot(
    data = data1,
    mapping = ggplot2::aes(
      x = eval(as.symbol((age_var))),
      y = eval(as.symbol(outcome))
    )
  ) +
    ggplot2::geom_point(alpha = .03, color = "navy") +
    ggplot2::geom_point(
      data = data2,
      mapping = ggplot2::aes(
        x = eval(as.symbol((age_var))),
        y = eval(as.symbol(outcome))
      ),
      alpha = .5,
      color = "navy"
    ) +
    ggplot2::labs(
      x = age_name,
      y = outcome_name
    )
}


datasets <- lapply(datasets, function(dataset) {
  data1 <- dataset$data
  outcome <- dataset$outcome
  age_var <- dataset$age_var[[1]]
  age_var <- paste0(age_var, "_ns")
  age_var_name <- paste(
    dataset$labels$age_var_name,
    dataset$labels$age_var_units
  )
  outcome_name <- paste(
    dataset$labels$oc_name,
    dataset$labels$oc_units
  )

  prediction_data <- dataset$prediction_data$data

  dataset$validation_plots <- lapply(prediction_data, function(data) {
    lapply(data, plot_validation,
      data1 = data1,
      age = age_var, outcome = outcome,
      age_name = age_var_name,
      outcome_name = outcome_name
    )
  })

  dataset
})
```

### Mortality outcomes by class

- Supplement will include same figures but for each sex strain strata

#### Complete data

```{r mortality outcomes by class, fig.height= 16, fig.width=12}
kap_plots <- final_models[og_index, "kap_plot"]

if (config$custom_palette) {
  j <- 0
  for (i in seq_along(kap_plots)) {
    group <- kap_plots[[i]]$plot$data$strata
    k <- length(unique(group)) + j
    kap_plots[[i]]$plot <- kap_plots[[i]]$plot +
      scale_color_manual(
        values = config$palette_colors[j + 1:k],
        name = "Class",
        labels = config$legend_labels[j + 1:k]
      )
    j <- k
  }
}
kap_plots <- lapply(kap_plots, `[[`, "plot")
plot_list <- c(obs_plots, kap_plots)

mortality_panel <- cowplot::plot_grid(
  plotlist = plot_list, nrow = length(kap_plots),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)
title_theme <- ggdraw() +
  draw_label(
    "Observed Metabolic Life Course Trajectories and Kaplan Meiers by Class"
  )

mortality_panel <- cowplot::plot_grid(title_theme, mortality_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

#### Training data

```{r mortality outcomes by class training, fig.height= 16, fig.width=12}
kap_plots_train <- final_models[train_index, "kap_plot"]

if (config$custom_palette_train) {
  j <- 0
  for (i in seq_along(kap_plots_train)) {
    group <- kap_plots_train[[i]]$plot$data$strata
    k <- length(unique(group)) + j
    kap_plots_train[[i]]$plot <- kap_plots_train[[i]]$plot +
      scale_color_manual(
        values = config$palette_colors_train[j + 1:k],
        name = "Class",
        labels = config$legend_labels_train[j + 1:k]
      )
    j <- k
  }
}
kap_plots_train <- lapply(kap_plots_train, `[[`, "plot")
plot_list <- c(obs_plots, kap_plots_train)

mortality_panel_train <- cowplot::plot_grid(
  plotlist = plot_list, nrow = length(kap_plots_train),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)
title_theme <- ggdraw() +
  draw_label(
    "Observed Metabolic Life Course Trajectories and Kaplan Meiers by Class"
  )

mortality_panel_train <- cowplot::plot_grid(title_theme, mortality_panel_train,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Table of Hazard Ratios, Raw and Sex/Strain Adjusted

```{r hazard ratio table, results = 'asis'}
surv_gethr <- function(fit, vars, names = NULL, ndec = 4) {
  sum_fit <- summary(fit) ## get summary of fit
  # save coeffecients as dataframe
  coef <- data.frame(sum_fit$coefficients, check.names = F)
  # save confidence interval as dataframe
  conf <- data.frame(sum_fit$conf.int, check.names = F)
  # initialize hr dataframe
  hr <- data.frame()
  # loop through vars
  for (var in vars) {
    # save hr
    value <- format(round(conf[var, "exp(coef)"], ndec), nsmall = 2)
    # save lower HR bound
    lower <- format(round(conf[var, "lower .95"], ndec), nsmall = 2)
    # save upper HR bound
    upper <- format(round(conf[var, "upper .95"], ndec), nsmall = 2)
    # save pval
    pval <- coef[var, "Pr(>|z|)"]

    ## add stars depending on significance
    if (pval < .001) {
      final <- paste0(value, " (", lower, ", ", upper, ")***")
    } else if (pval < .005) {
      final <- paste0(value, " (", lower, ", ", upper, ")**")
    } else if (pval < 0.050) {
      final <- paste0(value, " (", lower, ", ", upper, ")*")
    } else {
      final <- paste0(value, " (", lower, ", ", upper, ")")
    }

    ## save stats above to hr dataframe
    hr[var, c("value", "lower", "upper", "pval", "final")] <- c(
      value,
      lower,
      upper,
      pval,
      final
    )
  }

  ## if names are provided, rename variables
  if (!is.null(names)) {
    rownames(hr) <- names
  }

  ## create a column that has the rownames printed before HR
  hr[, "with_names"] <- unlist(lapply(rownames(hr), function(name) {
    el <- paste(name, hr[name, "final"], sep = ": ")
  }))

  ## create text that can be used to easily add to KM plot in place of pval
  one_text <- paste(as.vector(hr[, "with_names"]), collapse = "\n")

  return(list(hr_table = hr, hr_text = one_text))
}


make_cox_table <- function(cox_models, mo_select = c(1, 3), outcome) {
  xlevels_list <- lapply(cox_models, `[[`, "xlevels")
  xlevels_names <- lapply(xlevels_list, names)

  xlevels_list <- lapply(seq_along(xlevels_names), function(i) {
    vars <- xlevels_list[[i]]
    names <- xlevels_names[[i]]
    new_names <- mapply(paste0, names, vars, SIMPLIFY = FALSE)
    new_names <- unlist(lapply(new_names, `[`, -1))
  })


  hr_table_list <- list()
  for (i in mo_select) {
    hr_table <- surv_gethr(cox_models[[i]],
      vars = xlevels_list[[i]]
    )$hr_table
    hr_table <- cbind(row.names(hr_table), hr_table)
    names(hr_table)[[1]] <- "Class"
    row.names(hr_table) <- NULL
    model_name <- paste("HR Model", i)
    hr_table <- hr_table %>%
      select(Class, hr = final)
    hr_table <- cbind(outcome, hr_table)
    names(hr_table) <- c("Outcome", "Class", model_name)
    hr_table_list[[i]] <- hr_table
  }
  result <- purrr::reduce(hr_table_list, dplyr::left_join, by = c("Class", "Outcome"))
}


combine_cox_table <- function(final_models, mo_names, cox_mo_select) {
  x <- final_models
  cox_table_list <- list()
  for (name in mo_names) {
    cox_models <- x[x$model_name == name, ]$cox_models[[1]]
    outcome <- x[x$model_name == name, ]$oc_name
    ng <- x[x$model_name == name, ]$ng
    if (ng != 1) {
      cox_table_list[[name]] <- make_cox_table(
        cox_models,
        cox_mo_select,
        outcome
      )
    } else {
      cox_table_list[[name]] <- NA
    }
  }
  names(cox_table_list) <- NULL
  do.call(rbind, cox_table_list)
}

model_names <- final_models[og_index, "model_name"]
hr_table <- combine_cox_table(final_models, model_names, c(1, 2, 3))

if (config$custom_palette) {
  splits <- base::strsplit(hr_table$Class, split = "Class")
  splits <- as.numeric(unlist(lapply(splits, function(split) {
    if (!anyNA(split)) {
      split <- split[[2]]
    } else {
      split <- NA
    }
    split
  })))
  hr_table$Class <- config$legend_labels[splits]
}
```

### Metabolic Outcomes by Class

- Supplement will show the same figures but fore each sex strain combo

#### Complete Data

```{r metabolic outcomes by class, fig.height=16, fig.width=24}
og_plots_index <- grepl("^plot.*_og_.*across$", names(final_models))
length_plots <- length(which(og_plots_index))

og_plots <- final_models[og_index, og_plots_index]


if (config$custom_palette) {
  for (n in seq_along(og_plots)) {
    j <- 0
    plots <- og_plots[[n]]
    for (i in seq_along(plots)) {
      group <- rlang::eval_tidy(plots[[i]]$mapping$colour,
        data = plots[[i]]$data
      )
      k <- length(unique(group)) + j
      plots[[i]] <- plots[[i]] +
        scale_color_manual(
          values = config$palette_colors[j + 1:k],
          name = "Class",
          labels = config$legend_labels[j + 1:k]
        )
      j <- k
    }
    og_plots[[n]] <- plots
  }
}



og_plots <- unlist(og_plots, recursive = FALSE)

other_met_panel <- cowplot::plot_grid(
  plotlist = og_plots,
  nrow = length(which(og_index)),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
    sep = " "
  ))

other_met_panel <- cowplot::plot_grid(title_theme, other_met_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

#### Training Data

```{r metabolic outcomes by class train, fig.height=16, fig.width=24}
train_plots_index <- grepl("^plot.*_og_.*across$", names(final_models))
length_plots <- length(which(train_plots_index))

train_plots <- final_models[train_index, train_plots_index]


if (config$custom_palette) {
  for (n in seq_along(train_plots)) {
    j <- 0
    plots <- train_plots[[n]]
    for (i in seq_along(plots)) {
      group <- rlang::eval_tidy(plots[[i]]$mapping$colour,
        data = plots[[i]]$data
      )
      k <- length(unique(group)) + j
      plots[[i]] <- plots[[i]] +
        scale_color_manual(
          values = config$palette_colors[j + 1:k],
          name = "Class",
          labels = config$legend_labels[j + 1:k]
        )
      j <- k
    }
    train_plots[[n]] <- plots
  }
}



train_plots <- unlist(train_plots, recursive = FALSE)

other_met_panel_train <- cowplot::plot_grid(
  plotlist = train_plots,
  nrow = length(which(og_index)),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
    sep = " "
  ))

other_met_panel_train <- cowplot::plot_grid(title_theme, other_met_panel_train,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### New Figure 1

```{r new figure 1, fig.height = 24, fig.width = 24}
selector <- matrix(seq_len(length_plots * length_plots), nrow = length_plots)
diag(selector) <- NA
selector <- t(matrix(t(selector)[which(!is.na(selector))], ncol = length_plots))
selector <- as.vector(selector)
unique_plots <- og_plots[selector]

new_figure1_plot_list <- c(obs_plots, kap_plots, unique_plots)


new_figure1_plot_list <- lapply(new_figure1_plot_list, function(p) {
  p +
    coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
})


new_figure1 <- cowplot::plot_grid(
  plotlist = new_figure1_plot_list,
  nrow = length(which(og_index)),
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
    sep = " "
  ))

new_figure1 <- cowplot::plot_grid(title_theme, new_figure1,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

```{r save plots from above chunk}
for (i in seq_along(new_figure1_plot_list)) {
  plot <- new_figure1_plot_list[[i]]
  name <- paste0("plot_", i, ".png")
  output_path <- file.path("output", config$out_tag, "outcome")
  dir.create(output_path, recursive = TRUE)
  output_path <- normalizePath(output_path)


  ggsave(filename = name, plot = plot, device = "png", path = output_path)
}
```

```{r model_display}
model_display <- 4
```

### Table of lme models
- Model1 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model2 : $outcome = Age + Age^2 + Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model3 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model4 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model5 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Female * HET3 + (Age + Age^2 | Subject)$
- Model displayed: `r model_display`
```{r table lme, results = 'asis'}
rows <- final_models[og_index, ]

age_var <- rows$age_var[[1]]
age_var_name <- rows$age_var_name[[1]]
age_var_units <- rows$age_var_units[[1]]

age_var2_name <- paste(age_var_name, "2</sup>", sep = "<sup>")
age_var2_units <- paste(age_var_units, "2</sup>", sep = "<sup>")

age_var2 <- paste(age_var, 2, sep = "")
age_var_ns <- paste(age_var, "ns", sep = "_")
age_var2_ns <- paste(age_var2, "ns", sep = "_")
class_names <- paste("Class", seq_len(100))
names(class_names) <- paste0("new_class", seq_len(100))
if (config$custom_palette) {
  class_names <- config$legend_labels
  names(class_names) <- paste0("new_class", seq_along(config$legend_labels))
}
age_names <- paste(age_var_name, age_var_units)
names(age_names) <- age_var_ns
age2_names <- paste(age_var2_name, age_var2_units)
names(age2_names) <- age_var2_ns
age_interaction_names <- paste(age_names, class_names, sep = ":")
names(age_interaction_names) <- paste(names(age_names),
  names(class_names),
  sep = ":"
)
age2_interaction_names <- paste(age2_names, class_names, sep = ":")
names(age2_interaction_names) <- paste(names(age2_names), names(class_names), sep = ":")
fixcov_names <- c("female" = "Female", "het3" = "HET3")
fixcov_names_interacted <- paste(fixcov_names, collapse = ":")
names(fixcov_names_interacted) <- paste(names(fixcov_names), collapse = ":")

all_names <- c(
  class_names,
  age_names,
  age2_names,
  age_interaction_names,
  age2_interaction_names,
  fixcov_names,
  fixcov_names_interacted
)

lmes_index <- grepl("^model.*_og_.*across$", names(final_models))
lmes_list <- final_models[og_index, lmes_index]

model_list <- lapply(lmes_list, function(row) {
  lapply(row, `[[`, model_display)
})

model_calls <- lapply(model_list, function(models) {
  lapply(models, function(model) {
    model@call
  })
})

invert_list <- function(list) {
  list_matrix <- do.call(rbind, list)
  output <- apply(list_matrix, 2, as.list)
  output
}
model_list <- invert_list(model_list)



oc_names_list <- rep(list(rows$oc_name), times = length(lmes_list))

lme_table_list <- list()
title_list <- paste(rows$oc_name, "Class")

lme_table <- lapply(seq_along(model_list), function(i) {
  title <- paste(title_list[[i]], "Mixed Effects Models")
  lme_table <- tab_model(model_list[[i]],
    show.ci = FALSE,
    show.re.var = FALSE,
    pred.labels = all_names,
    show.reflvl = TRUE,
    wrap.labels = 50,
    show.icc = FALSE,
    show.r2 = FALSE,
    show.ngroups = FALSE,
    show.obs = FALSE,
    dv.labels = oc_names_list[[i]],
    title = title
  )
  lme_table
})

lme_table
```

### Class Prediction Panel

- Supplmenet will include same table for glucose and body fat

```{r class prediction panel, fig.height=16, fig.width=16}
interval_plots <- final_models[og_index, "interval_boot_accuracy_plot"]
window_plots <- final_models[og_index, "window_boot_accuracy_plot"]
cumulative_plots <- final_models[og_index, "cumulative_boot_accuracy_plot"]
resampled_plots <- final_models[og_index, "resampled_boot_accuracy_plot"]

if (window_legend) {
  for (i in seq_along(window_plots)) {
    window_plots[[i]] <- window_plots[[i]] +
      scale_color_manual(
        values = window_legend_palette,
        name = "Window Size",
        labels = window_legend_labels
      )
  }
}

if (cumulative_legend) {
  for (i in seq_along(cumulative_plots)) {
    cumulative_plots[[i]] <- cumulative_plots[[i]] +
      scale_color_viridis(begin = .2, end = .8, direction = -1, discrete = TRUE)
  }
}

interval_plots <- lapply(interval_plots, function(plot) {
  plot +
    coord_cartesian(ylim = c(0, 1))
})

window_plots <- lapply(window_plots, function(plot) {
  plot +
    coord_cartesian(xlim = c(0, NA), ylim = c(0, 1))
})

cumulative_plots <- lapply(cumulative_plots, function(plot) {
  plot +
    coord_cartesian(xlim = c(0, NA), ylim = c(0, 1))
})

resampled_plots <- lapply(resampled_plots, function(plot) {
  plot +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
})

plot_list <- c(interval_plots, window_plots, cumulative_plots, resampled_plots)


title <- paste("Classification Accuracy")

title_theme <- ggdraw() +
  draw_label(title)

class_prediction_panel <- cowplot::plot_grid(
  plotlist = plot_list,
  ncol = 4,
  labels = "AUTO",
  byrow = FALSE
)

class_prediction_panel <- cowplot::plot_grid(title_theme,
  class_prediction_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

```{r save plots from chunk above}
for (i in seq_along(plot_list)) {
  plot <- plot_list[[i]]
  name <- paste0("plot_", i, ".png")
  output_path <- file.path("output", config$out_tag, "classification")
  dir.create(output_path, recursive = TRUE)
  output_path <- normalizePath(output_path)


  ggsave(filename = name, plot = plot, device = "png", path = output_path)
}
```

### Combinded Cox Model 

```{r combined cox model}
test_index <- final_models$data_mod == "train_test"
rest_index <- !test_index
test_index <- which(test_index)[[1]]
rest_index <- which(rest_index)[[1]]

test_combined_cox <- list(
  survminer::ggforest(
    final_models$cox_combined_models[[test_index]][[1]],
    final_models$combined_census[[test_index]],
    main = "HR Testing Combined Cox"
  ),
  survminer::ggforest(
    final_models$cox_combined_models[[test_index]][[2]],
    final_models$combined_census[[test_index]],
    main = "HR Testing Combined Cox"
  ),
  survminer::ggforest(
    final_models$cox_combined_models[[test_index]][[3]],
    final_models$combined_census[[test_index]],
    main = "HR Testing Combined Cox"
  )
)

rest_combined_cox <- list(
  survminer::ggforest(
    final_models$cox_combined_models[[rest_index]][[1]],
    final_models$combined_census[[rest_index]],
    main = "HR Complete Data Combined Cox"
  ),
  survminer::ggforest(
    final_models$cox_combined_models[[rest_index]][[2]],
    final_models$combined_census[[rest_index]],
    main = "HR Complete Data Combined Cox"
  ),
  survminer::ggforest(
    final_models$cox_combined_models[[rest_index]][[3]],
    final_models$combined_census[[rest_index]],
    main = "HR Complete Data Combined Cox"
  )
)
```

### Dose response

```{r source surv analysis, file = 'R/surv_dose.R'}
source("R/surv_dose.R")
```

```{r dose response, fig.height = 8, fig.width = 8, include = FALSE}
test_index <- final_models$data_mod == "train_test"
rest_index <- !test_index
test_index <- which(test_index)[[1]]
rest_index <- which(rest_index)[[1]]

census <- final_models[rest_index, "combined_census"][[1]]
prob_cols <- names(census)[grepl("^prob\\d+", names(census))]

for (prob_col in prob_cols) {
  name <- paste0(prob_col, "_cut")
  name_tile <- paste0(prob_col, "_tile")
  census[[name]] <- cut(census[[prob_col]], breaks = 5, dig.lab = 2)
  census[[name_tile]] <- Hmisc::cut2(census[[prob_col]], g = 5)
}

prob_cols_tile <- paste0(prob_cols, "_tile")

dose_plots <- lapply(seq_along(prob_cols_tile), function(i) {
  if (length(unique(census[[prob_cols_tile[[i]]]])) == 1) {
    print("returning NA")
    return(NA)
  }
  if (config$custom_palette) {
    kap_plot_all_dose(
      df = census, var = prob_cols_tile[[i]],
      age_death = config$survival_dataset$age_death,
      event = config$survival_dataset$event,
      color = config$palette_colors[[i]],
      ptitle = paste(
        "Probability of ", config$legend_labels[[i]],
        "Class Quintile"
      ),
      class_number = i
    )
  } else {
    kap_plot_all_dose(
      df = census, var = prob_cols_tile[[i]],
      age_death = config$survival_dataset$age_death,
      event = config$survival_dataset$event,
      ptitle = paste("Probability of Class", i, "Quintile"),
      class_number = i
    )
  }
})

dose_plots <- lapply(dose_plots, function(plot) {
  if (anyNA(plot)) {
    plot <- NA
  } else {
    plot <- plot[["plot"]]
  }
  plot
})

dose_plot_panel <- cowplot::plot_grid(
  plotlist = dose_plots,
  nrow = round(sqrt(length(prob_cols))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Dose Response of Probability of Class Membership",
    sep = " "
  ))

dose_plot_panel <- cowplot::plot_grid(title_theme, dose_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Dose Response for test data with cumulative validation data

```{r cumulative validation dose response, include = FALSE}
test_index <- final_models$data_mod == "train_test"
rest_index <- !test_index
test_index <- which(test_index)[[1]]
rest_index <- which(rest_index)[[1]]

# 6 to 58 weeks
censuses <- final_models[test_index, "filter_cumulative_test_tmerged"][[1]]
census_names <- names(censuses)

dose_plot_cum_validation <- lapply(censuses, function(census) {
  prob_cols <- names(census)[grepl("^prob\\d+", names(census))]

  for (prob_col in prob_cols) {
    name <- paste0(prob_col, "_cut")
    name_tile <- paste0(prob_col, "_tile")
    census[[name]] <- cut(census[[prob_col]], breaks = 5, dig.lab = 2)
    census[[name_tile]] <- Hmisc::cut2(census[[prob_col]], digits = 7, g = 5)
  }

  prob_cols_tile <- paste0(prob_cols, "_tile")

  dose_plots_cum_validation <- lapply(seq_along(prob_cols_tile), function(i) {
    if (length(unique(census[[prob_cols_tile[[i]]]])) == 1) {
      print("returning NA")
      return(NA)
    }
    if (config$custom_palette_train) {
      kap_plot_all_dose(
        df = census, var = prob_cols_tile[[i]],
        age_death = config$survival_dataset$age_death,
        event = config$survival_dataset$event,
        color = config$palette_colors[[i]],
        ptitle = paste(
          "Probability of ", config$legend_labels_train[[i]],
          "Class Quintile"
        ),
        class_number = i
      )
    } else {
      kap_plot_all_dose(
        df = census, var = prob_cols_tile[[i]],
        age_death = config$survival_dataset$age_death,
        event = config$survival_dataset$event,
        ptitle = paste("Probability of Class", i, "Quntile"),
        class_number = i
      )
    }
  })

  dose_plots_cum_validation <- lapply(dose_plots_cum_validation, function(plot) {
    if (anyNA(plot)) {
      plot <- NA
    } else {
      plot <- plot[["plot"]]
    }
    plot
  })

  dose_plot_cum_validation_panel <- cowplot::plot_grid(
    plotlist = dose_plots_cum_validation,
    nrow = round(sqrt(length(prob_cols))),
    byrow = TRUE,
    greedy = FALSE,
    labels = "AUTO"
  )

  title_theme <- ggdraw() +
    draw_label(paste(
      "Dose Response of Probability of Class Membership",
      sep = " "
    ))

  dose_plot_cum_validation_panel <- cowplot::plot_grid(title_theme,
    dose_plot_cum_validation_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )
})
```

### KM by class test data with cumulative validation data

```{r source surv analysis for km validation, file = 'R/surv_all.R'}
source("R/surv_all.R")
```

```{r cumulative validation by kaplan meier class}
test_index <- final_models$data_mod == "train_test"
rest_index <- !test_index
test_index <- which(test_index)[[1]]
rest_index <- which(rest_index)[[1]]

# 6 to 58 weeks
censuses <- final_models[test_index, "filter_cumulative_test_tmerged"][[1]]
census_names <- names(censuses)

kap_plot_cum_validation <- lapply(censuses, function(census) {
  class_cols <- names(census)[grepl("^new_class", names(census))]

  kap_plots_cum_validation <- lapply(seq_along(class_cols), function(i) {
    if (config$custom_palette) {
      kap_plot_all(
        df = census,
        var = class_cols[[i]],
        age_death = config$survival_dataset$age_death,
        event = config$survival_dataset$event
      )
    } else {
      kap_plot_all(
        df = census,
        var = class_cols[[i]],
        age_death = config$survival_dataset$age_death,
        event = config$survival_dataset$event
      )
    }
  })

  kap_plots_cum_validation <- lapply(kap_plots_cum_validation, `[[`, "plot")

  kap_plot_cum_validation_panel <- cowplot::plot_grid(
    plotlist = kap_plots_cum_validation,
    ncol = length(class_cols),
    byrow = TRUE,
    greedy = FALSE,
    labels = "AUTO"
  )

  title_theme <- ggdraw() +
    draw_label(paste(
      "KM Validation of Test Data",
      sep = " "
    ))

  kap_plot_cum_validation_panel <- cowplot::plot_grid(title_theme,
    kap_plot_cum_validation_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )
})
```

### Histogram of outcomes for complete data

```{r hist outcome, echo = FALSE, include = FALSE}
hist_outcome_plots <- lapply(datasets, function(dataset) {
  data <- dataset$data
  outcome <- dataset$outcome
  data_mod <- dataset$data_mod
  ggplot2::ggplot(data = data, aes_string(x = outcome)) +
    ggplot2::geom_histogram() +
    ggplot2::labs(title = data_mod)
})

hist_outcome_plot_panel <- cowplot::plot_grid(
  plotlist = hist_outcome_plots,
  nrow = round(sqrt(length(datasets))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of outcomes for complete and training data",
    sep = " "
  ))

hist_outcome_plot_panel <- cowplot::plot_grid(title_theme,
  hist_outcome_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Histogram of log(outcomes) for complete data

```{r hist log outcome, echo = FALSE, include = FALSE}
hist_log_outcome_plots <- lapply(datasets, function(dataset) {
  data <- dataset$data
  outcome <- dataset$outcome
  outcome <- paste0("log(", outcome, ")")
  data_mod <- dataset$data_mod
  ggplot2::ggplot(data = data, aes_string(x = outcome)) +
    ggplot2::geom_histogram() +
    ggplot2::labs(title = data_mod)
})

hist_log_outcome_plot_panel <- cowplot::plot_grid(
  plotlist = hist_log_outcome_plots,
  nrow = round(sqrt(length(datasets))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of log(outcomes) for complete and training data",
    sep = " "
  ))

hist_log_outcome_plot_panel <- cowplot::plot_grid(title_theme,
  hist_log_outcome_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Histogram of probabilites of class assignment for all data 

```{r hist prob, echo = FALSE, include = FALSE}
census <- final_models[rest_index, "combined_census"][[1]]
hist_plots <- lapply(prob_cols, function(prob_col) {
  ggplot2::ggplot(data = census, aes_string(x = prob_col)) +
    ggplot2::geom_histogram()
})

hist_plot_panel <- cowplot::plot_grid(
  plotlist = hist_plots,
  nrow = round(sqrt(length(prob_cols))),
  byrow = TRUE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Histograms of probabilites of class membership for all subjects",
    sep = " "
  ))

hist_plot_panel <- cowplot::plot_grid(title_theme, hist_plot_panel,
  rel_heights = c(.05, 1),
  ncol = 1
)
```

### Descriptive stats of probabilities of class assignments - Complete Data

```{r describe prob class, results = 'asis'}
census <- final_models[rest_index, "combined_census"][[1]]
kable(psych::describe(census))
```

### Kaplan Meier by Cohort

This may explain our trouble externally validating with cohorts 9 and 10, since
both have lower levels of survival

```{r source surv analysis for cohort, file = 'R/surv_all.R'}
source("R/surv_all.R")
```

```{r kap plot by cohort}
kap_plot_all(census, "cohort",
  age_death = config$survival_dataset$age_death,
  event = config$survival_dataset$event
)
```

## Display Figures

```{r initiate figure and table number}
fig_no <- 1
tab_no <- 1
```

### Figure `r as.character(fig_no)`: How it works

```{r how it works, fig.width = 12, fig.height = 6}
how_it_works
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Predicted Observed

#### Complete Data

```{r predicted and observed, fig.height=18, fig.width=12}
pred_obs_panel
```

#### Training Data

```{r predicted and observed train, fig.height=18, fig.width=12}
pred_obs_panel_train
fig_no <- fig_no + 1
```

### Table `r as.character(tab_no)`: Whos in each Class

#### Complete Data

```{r table whos in each class, results = 'asis'}
print_table <- function(env, table_name, title) {
  cap <- paste(env$subgroup,
    title,
    sep = " "
  )
  table <- kable(env[[table_name]],
    format = "html",
    caption = cap
  ) %>%
    column_spec(1:2, bold = TRUE, border_right = TRUE) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    kable_styling(bootstrap_options = "striped")

  return(table)
}

print_table(.GlobalEnv,
  table_name = "t1_df",
  title = "Demographics by Metabolic Index and Class"
)
tab_no <- tab_no + 1
```

#### Training Data

```{r table whos in each class train, results = 'asis'}
print_table(.GlobalEnv,
  table_name = "t1_df_train",
  title = "Demographics by Metabolic Index and Class"
)
```

### Figure `r as.character(fig_no)`: Metabolic Classes with Other Metabolic Metrics

#### Complete Data

```{r plot classes by other outcomes, fig.height=18, fig.width=24}
other_met_panel
```

#### Training Data

```{r plot classes by other outcomes train, fig.height=18, fig.width=24}
other_met_panel_train
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Heatmap of census by class probability 

#### Without Covariates

```{r heatmap of census with out covariate, fig.height=12, fig.width=12}
heatmap2
```

#### With Covariates

```{r heatmap of census with covariate, fig.height=12, fig.width=12}
heatmap1
fig_no <- fig_no + 1
```

### Table `r as.character(tab_no)`: Metabolic Classes Predict other Metabolic Metrics

- Model1 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model2 : $outcome = Age + Age^2 + Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model3 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model4 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model5 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Female * HET3 + (Age + Age^2 | Subject)$

#### Model displayed: `r model_display`

#### Complete Data

```{r display model_calls}
model_calls
```

```{r table 2, results = 'asis'}
for (table in lme_table) {
  cat(table$knitr)
}
tab_no <- tab_no + 1
```

### Figure `r as.character(fig_no)`: Mortality and Class

#### Complete Data

```{r mortality and class, fig.height=18, fig.width=12}
mortality_panel
fig_no <- fig_no + 1
```

#### Training Data

```{r mortality and class training, fig.height=18, fig.width=12}
mortality_panel_train
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Class, Mortality, and other outcome - Complete Data

```{r class mortality and other outcome, fig.height=24, fig.width=30}
new_figure1
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Dose Response - Complete Data

```{r dose reponse, fig.height=18, fig.width=18}
dose_plot_panel
fig_no <- fig_no + 1
```

### Table `r as.character(tab_no)`: Hazard Ratios - Complete Data

- Model 1: $ Surv ~ Class $
- Model 2: $ Surv ~ Class + Sex + Strain$
- Model 3: $ Surv ~ Class + Sex * Strain$

```{r table hazard ratios, results = 'asis'}
print_table(.GlobalEnv,
  table_name = "hr_table",
  title = "Table of Hazard Ratios by Metabolic Indices"
)
tab_no <- tab_no + 1
```

### Figure `r as.character(fig_no)`: Class Prediction Panel - Complete Data

```{r Class prediction panel, fig.width=30, fig.height=24}
class_prediction_panel
fig_no <- fig_no + 1
```

### KM Class Validation - Testing Data

The 20% testing data was applied to the 80% training latent class mixed model
to derive their predicted class assignments. These class assignments are then 
used to make the KM Curves below

#### Figure `r as.character(fig_no)`: KM Class Validation @ `r names(kap_plot_cum_validation)[[3]]` weeks

```{r km cum validation 1, fig.height=6, fig.width=18}
kap_plot_cum_validation[[3]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: KM Class Validation @ `r names(kap_plot_cum_validation)[[4]]` weeks

```{r km cum validation 2, fig.height=6, fig.width=18}
kap_plot_cum_validation[[4]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: KM Class Validation @ `r names(kap_plot_cum_validation)[[5]]` weeks

```{r km cum validation 3, fig.height=6, fig.width=18}
kap_plot_cum_validation[[5]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: KM Class Validation @ `r names(kap_plot_cum_validation)[[6]]` weeks

```{r km cum validation 4, fig.height=6, fig.width=18}
kap_plot_cum_validation[[6]]
fig_no <- fig_no + 1
```

### Dose Response Validation - Testing Data

The 20% testing data was applied to the 80% training latent class mixed model
to derive their predicted class assignments probabilities. These class 
assignment probabilities were then quintiled and plotted in the KM curves 
below. 

#### Figure `r as.character(fig_no)`: Dose Response Cumulative Validation @ `r names(dose_plot_cum_validation)[[3]]` weeks

```{r dose reponse cum validation 1, fig.height=18, fig.width=18}
dose_plot_cum_validation[[3]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: Dose Response Cumulative Validation @ `r names(dose_plot_cum_validation)[[4]]` weeks

```{r dose reponse cum validation 2, fig.height=18, fig.width=18}
dose_plot_cum_validation[[4]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: Dose Response Cumulative Validation @ `r names(dose_plot_cum_validation)[[5]]` weeks

```{r dose reponse cum validation 3, fig.height=18, fig.width=18}
dose_plot_cum_validation[[5]]
fig_no <- fig_no + 1
```

#### Figure `r as.character(fig_no)`: Dose Response Cumulative Validation @ `r names(dose_plot_cum_validation)[[6]]` weeks

```{r dose reponse cum validation 4, fig.height=18, fig.width=18}
dose_plot_cum_validation[[6]]
fig_no <- fig_no + 1
```

### Table `r as.character(tab_no)`: Combined Classes Cox Models

```{r get model calls}
model_formulas <- lapply(
  final_models$cox_combined_models[[rest_index]],
  `[[`, "formula"
)
model_formulas <- lapply(model_formulas, function(formula) {
  Reduce(paste, deparse(formula))
})
```

These are the combined models we will use to predict mortality there are
three model calls

- Model 1: `r model_formulas[[1]]`
- Model 2: `r model_formulas[[2]]`
- Model 3: `r model_formulas[[3]]`

#### Complete Data

```{r print combined cox models all data}
lapply(final_models$cox_combined_models[[rest_index]], function(model) {
  list(cox_model = model, proportiona_hazards_test = cox.zph(model))
})
```

#### Training Data
```{r print combined cox models training}
lapply(final_models$cox_combined_models[[test_index]], function(model) {
  list(cox_model = model, proportiona_hazards_test = cox.zph(model))
})
tab_no <- tab_no + 1
```

### Figure `r as.character(fig_no)`: Combined Cox Forrest Plots

These take the models shown in the previous table and display them in forrest
plot format. It also drops any interaction variables

#### Complete Data

Key: 

```{r combined cox key, results = 'asis'}
if (config$custom_palette) {
  df <- data.frame(
    Class = paste("Class", seq_along(config$legend_labels)),
    Label = config$legend_labels
  )
  kable(df)
}
```

```{r combined cox forrest plots all data}
rest_combined_cox
```

#### Training data

Key: 

```{r combined cox key train, results = 'asis'}
if (config$custom_palette_train) {
  df <- data.frame(
    Class = paste("Class", seq_along(config$legend_labels_train)),
    Label = config$legend_labels_train
  )
  kable(df)
}
```

```{r combined cox forrest plots training}
test_combined_cox
fig_no <- fig_no + 1
```

```{r save cominbed cox forrest plots}
p <- c(test_combined_cox, rest_combined_cox)
for (i in seq_along(p)) {
  plot <- p[[i]]
  name <- paste0("plot_", i, ".png")
  output_path <- file.path("output", config$out_tag, "forest")
  dir.create(output_path, recursive = TRUE)
  output_path <- normalizePath(output_path)
  file_name <- file.path(output_path, name)

  ggsave(filename = name, plot = plot, device = "png", path = output_path)
  png(file = file_name)
  plot
  dev.off()
}
```

### Table `r as.character(tab_no)`: Combined Cox Feature Importance

This takes the combined cox models from above and assesses the importance 
of each added variable in predicting mortality

#### Analysis of Deviance: Complete Data

```{r feature importance all data}
models <- final_models$cox_combined_models[[rest_index]]
for (model in models) {
  print(model$formula)
  print(anova(model))
}
```

#### Analysis of Deviance: Training Data

```{r feature importance training data}
models <- final_models$cox_combined_models[[test_index]]
for (model in models) {
  print(model$formula)
  print(anova(model))
}

tab_no <- tab_no + 1
```

```{r validation is this internal or external}
if (!isFALSE(config$external_validate)) {
  validation_type <- "External"
} else {
  validation_type <- "Internal"
}
```

### Figure `r as.character(fig_no)`: `r validation_type` Validation Compare models - Testing Data

The 20% testing data was applied to the 80% training latent class model to 
determine their predicted class probability. These predicted class probabilities
for the 20% testing data were then applied to the 80% training combined cox
models to determine if the predicted class probabilities could be used as
predictors of mortality. 

Using the three combined cox models from above, we assess their performance 
using the linear predictor and concordance derived from validation data

Again we are still using these three models

- Model 1: `r model_formulas[[1]]`
- Model 2: `r model_formulas[[2]]`
- Model 3: `r model_formulas[[3]]`

```{r validation compare models preliminary}
final_models[!og_index, ]$filter_window_test_lp_plot[[1]]
final_models[!og_index, ]$filter_window_test_concord_plot[[1]]
final_models[!og_index, ]$filter_cumulative_test_lp_plot[[1]]
final_models[!og_index, ]$filter_cumulative_test_concord_plot[[1]]

fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: `r validation_type` Validation Model 2

After selecting the Model 2 from the figure above, we can dispaly it in a 
similar format to that of the reclassifcation figures

```{r validation with model 2, include = FALSE}
p <- list()
p[[1]] <- final_models[!og_index, ]$filter_window_test_concord_plot_model2[[1]]
p[[2]] <- final_models[!og_index, ]$filter_window_test_lp_plot_model2[[1]]
p[[3]] <- final_models[!og_index, ]$filter_cumulative_test_concord_plot_model2[[1]]
p[[4]] <- final_models[!og_index, ]$filter_cumulative_test_lp_plot_model2[[1]]
if (window_legend && cumulative_legend) {
  p[[1]] <- final_models[!og_index, ]$filter_window_test_lp_plot_model2[[1]] +
    scale_color_manual(
      values = window_legend_palette,
      name = "Window Size",
      labels = window_legend_labels
    ) +
    coord_cartesian(ylim = c(-1, 2))

  p[[2]] <- final_models[!og_index, ]$filter_window_test_concord_plot_model2[[1]] +
    scale_color_manual(
      values = window_legend_palette,
      name = "Window Size",
      labels = window_legend_labels
    ) +
    coord_cartesian(ylim = c(.5, 1))
  p[[3]] <- final_models[!og_index, ]$filter_cumulative_test_lp_plot_model2[[1]] +
    viridis::scale_color_viridis(
      begin = .2,
      end = .8,
      direction = -1,
      discrete = TRUE
    ) +
    ggplot2::coord_cartesian(ylim = c(-1, 2)) +
    ggplot2::labs(color = "Lower Bound (weeks)") +
    theme(legend.title = element_text(size = rel(.8)))


  p[[4]] <- final_models[!og_index, ]$filter_cumulative_test_concord_plot_model2[[1]] +
    viridis::scale_color_viridis(
      begin = .2,
      end = .8,
      direction = -1,
      discrete = TRUE
    ) +
    ggplot2::coord_cartesian(ylim = c(.5, 1)) +
    ggplot2::labs(color = "Lower Bound (weeks)") +
    theme(legend.title = element_text(size = rel(.8)))
}
p

fig_no <- fig_no + 1
```

```{r save vaidation plots}
for (i in seq_along(p)) {
  plot <- p[[i]]
  name <- paste0("plot_", i, ".png")
  output_path <- file.path("output", config$out_tag, "validation")
  dir.create(output_path, recursive = TRUE)
  output_path <- normalizePath(output_path)


  ggsave(filename = name, plot = plot, device = "png", path = output_path)
}
```

```{r display model 2 validation, fig.height = 12, fig.width = 12}
model2_grid <- cowplot::plot_grid(
  plotlist = p,
  nrow = 2,
  byrow = FALSE,
  greedy = FALSE,
  labels = "AUTO"
)

title_theme <- ggdraw() +
  draw_label(paste(
    "Validation using window and cumulative data",
    sep = " "
  ))

model2_grid <- cowplot::plot_grid(title_theme, model2_grid,
  rel_heights = c(.05, 1),
  ncol = 1
)

model2_grid
```

## Display Dataset/Model informaton 

Do this for complete dataset, training dataset, testing dataset.

- Number of observations per outcome
- Number of subjects per outcome
- Number of sujbects excluded for completeness

```{r dataset information, results = 'asis'}
make_wave <- function(data, id, age_var) {
  data <- data %>%
    dplyr::group_by(!!sym(id)) %>%
    dplyr::arrange(!!sym(age_var)) %>%
    dplyr::mutate(wave = dplyr::row_number()) %>%
    dplyr::mutate(dif_time = !!sym(age_var) - lag(!!sym(age_var))) %>%
    dplyr::ungroup() %>%
    as.data.frame()

  data
}

summarise_dataset <- function(dataset, data_name) {
  data <- dataset[[data_name]]
  if (is.null(data)) {
    return(NULL)
  }
  data_id <- dataset$data_id
  data_mod <- dataset$data_mod
  if (data_mod == "train_test" && data_name == "data") {
    data_name <- "training_data"
  }

  id <- dataset$id
  age <- paste0(dataset$age[[1]], "_ns")
  if (data_name == "missing_data") {
    age <- dataset$age[[1]]
  }
  outcome <- dataset$outcome
  unique_id <- length(unique(data[[id]]))
  data <- make_wave(data, id, age)
  sum <- round(psych::describe(data[, c(outcome, age, "wave", "dif_time")]), 2)
  rownames(sum) <- c(outcome, age, "wave", "dif_time")
  sum <- cbind(rownames(sum), data.frame(sum, row.names = NULL))
  names(sum)[1] <- "variables"
  sum <- data.frame(unique_id = rep(unique_id, nrow(sum)), sum)
  sum <- data.frame(outcome = rep(outcome, nrow(sum)), sum)
  sum <- data.frame(data_id = rep(data_id, nrow(sum)), sum)
  sum <- data.frame(data_name = rep(data_name, nrow(sum)), sum)
  sum
}


sum_data <- do.call(rbind, lapply(datasets, summarise_dataset, "data"))
rownames(sum_data) <- NULL
sum_missing <- do.call(
  rbind,
  lapply(datasets, summarise_dataset, "missing_data")
)
rownames(sum_missing) <- NULL
sum_test <- do.call(rbind, lapply(datasets, summarise_dataset, "test_data"))
rownames(sum_test) <- NULL
```

#### Complete and Training Data

```{r dataset info complete and training, results = 'asis'}
kable(sum_data)
```

#### Missing Data from Complete and Training Data

```{r dataset info missing complete and training, results = 'asis'}
kable(sum_missing)
```

#### Testing Data

```{r dataset info testing data, results = 'asis'}
kable(sum_test)
```

### Census information

- Study Demographics
- Median Life Expectancy

```{r census information, results = 'asis'}
describe_le <- function(data, age_death, event) {
  surv_object <- survival::Surv(time = data[[age_death]], event = data[[event]])
  surv_fit <- survival::survfit(surv_object ~ 1)
  quant_surv <- quantile(surv_fit)
  names(quant_surv$quantile) <- c("Q1", "Median", "Q3")
  names(quant_surv$lower) <- c("Lower Q1", "Lower Median", "Lower Q3")
  names(quant_surv$upper) <- c("Upper Q1", "Upper Median", "Upper Q3")
  quant_surv <- lapply(quant_surv, t)
  quant_surv <- lapply(quant_surv, as.data.frame)
  names(quant_surv) <- NULL
  quant_surv <- do.call(cbind, quant_surv)
  quant_surv <- data.frame(n = surv_fit$n, quant_surv)
  quant_surv$conf_int <- surv_fit$conf.int
  quant_surv$conf_type <- surv_fit$conf.type
  quant_surv$type <- surv_fit$type
  quant_surv
}
census_complete <- final_models$combined_census[[1]]
age_death <- config$survival_dataset$age_death
event <- config$survival_dataset$event
filter_vars <- config$datasets[[1]]$covariates
filtered_dfs <- lapply(filter_vars, function(filter) {
  print(filter)
  unique_values <- unique(census_complete[[filter]])

  filtered_dfs <- lapply(unique_values, function(value) {
    print(value)
    census_complete[census_complete[[filter]] == value, ]
  })
  names(filtered_dfs) <- unique_values
  filtered_dfs
})
names(filtered_dfs) <- filter_vars
filtered_dfs <- unlist(filtered_dfs, recursive = FALSE)

data <- c(list(all = census_complete), filtered_dfs)
# peek(data)

table_mle <- lapply(data, describe_le, age_death, event)
table_mle <- do.call(rbind, table_mle)

kable(table_mle)
```

### LCMM information 

Include latent class model fit statistics

#### LCMM - Concise Table

```{r lcmm concise, results = 'asis'}
table <- helphlme::compare_models(final_models$final_model)$table
table <- data.frame(
  outcome = final_models$oc,
  data_type = final_models$data_mod,
  table
)
kable(table)
```

#### LCMM - Verbose Table

```{r lcmm model information}
for (i in seq_len(nrow(final_models))) {
  label <- paste(final_models$oc[[i]], final_models$data_mod[[i]])
  cat("#############################################################\n")
  cat(label)
  cat("\n#############################################################\n")
  model <- final_models$final_model[[i]]
  summary(model)
  lcmm::postprob(model)
}
```

#### LCMM - Average Posterior Classifcation Probabilites

```{r lcmm model average posterior probs, results = 'asis'}
post_prob_table <- lapply(seq_len(nrow(final_models)), function(i) {
  label <- paste(final_models$oc[[i]], final_models$data_mod[[i]])
  outcome <- final_models$oc[[i]]
  data_type <- final_models$data_mod[[i]]
  model <- final_models$final_model[[i]]
  post_prob <- lcmm::postprob(model)
  post_prob_table <- data.frame(
    outcome = outcome,
    data_type = data_type,
    label = label,
    class = row.names(post_prob[[2]]),
    t(post_prob[[1]]),
    post_prob[[2]],
    t(post_prob[[3]]),
    check.names = FALSE,
    row.names = NULL
  )
  post_prob_table
})
post_prob_table <- data.table::rbindlist(post_prob_table, fill = TRUE)
post_prob_table <- as.data.frame(post_prob_table)
kable(post_prob_table)
```

#### LCMM - Model residual plots

```{r plot model residuals, fig.height = 12, fig.width = 12}
for (i in seq_len(nrow(final_models))) {
  title <- paste(final_models$data_mod[[i]], final_models$oc[[i]])
  model <- final_models$final_model[[i]]
  cat("##### ")
  cat(title)
  cat("\n")

  plot(model, which = "residuals", var.time = "age_wk_ns", title = title)
}
```

#### LCMM - Compare all possible models

```{r compare all models, results = 'asis'}
load(input$models_path)
load(input$cf_path)
all_models <- helphlme::compare_models(models)$table
all_models <- data.frame(
  model_number = all_models[, 1],
  data = cf$data,
  fixed_effects = cf$fixed,
  mixture = cf$mixture,
  all_models[, 2:length(all_models)],
  check.names = FALSE
)
all_models$`Smallest Class (%)` <- round(all_models$`Smallest Class (%)`, 2)
kable(all_models)
```

### Chi-squared information

```{r chi squared information }
for (i in seq_len(nrow(final_models))) {
  title <- paste(final_models$data_mod[[i]], final_models$oc[[i]])
  chi_test <- final_models$chi_test[[i]]
  cat("#############################################################\n")
  cat(title)
  cat("\n#############################################################\n")
  print(chi_test)
}
```

### LME information

- Selected model - `r model_display`

```{r lme information}
for (models in model_list) {
  for (model in models) {
    cat("#############################################################\n")
    print(model@call)
    cat("\n#############################################################\n")
    print(model)
    print(summary(model))
    print(plot(model))
  }
}
```

### LME table

- Selected model - `r model_display`

```{r lme table}
table_list <- lapply(model_list, function(models) {
  lapply(models, function(model) {
    model_sum <- summary(model)
    data.frame(
      call = as.character(model@call)[2],
      title = model_sum$methTitle,
      loglik = model_sum$logLik,
      n_groups = model_sum$n,
      sigma = model_sum$sigma,
      AICtab = model_sum$AICtab,
      optimizer = model@optinfo$optimizer,
      conv = model@optinfo$conv$opt
    )
  })
})

table_list <- unlist(table_list, recursive = FALSE)
lme_table <- do.call(rbind, table_list)
kable(lme_table)
```

### Individual Cox Model

#### print model output and print cox.zph output

```{r individual cox model}
rest_index <- which(final_models$data_mod == "og")
for (i in rest_index) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]
  title <- paste(outcome, data_mod)
  cat("#############################################################\n")
  print(title)
  cat("\n#############################################################\n")

  for (model in models) {
    print(model)
    print(cox.zph(model))
  }
}
```

#### Table coxzph Individual Cox Model Fits

```{r table coxzph individual cox model, results = 'asis'}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  zph_tables <- lapply(seq_along(models), function(i) {
    formula <- Reduce(paste, deparse(models[[i]]$formula))

    table <- cox.zph(models[[i]])$table
    table <- data.frame(
      outcome = outcome, data_type = data_mod, Model = i, formula = formula,
      variables = rownames(table), table
    )
    rownames(table) <- NULL
    table$chisq <- round(table$chisq, 4)
    table$p <- round(table$p, 4)

    table
  })
  zph_tables <- do.call(rbind, zph_tables)

  zph_tables
})

coxzph_tables <- do.call(rbind, coxzph_tables)
kable(coxzph_tables)
```

#### Plot Shoenfeld Residuals

```{r plot coxzph individual cox model residuals}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  lapply(seq_along(models), function(i) {
    model <- models[[i]]
    formula <- Reduce(paste, deparse(model$formula))

    coxzph <- cox.zph(model)
    plot(coxzph, main = formula)
  })
})
```

### Combined Cox Model

#### print model output and print cox.zph output - Combined Cox

```{r combined cox model print output}
rest_index <- which(final_models$data_mod == "og")
for (i in rest_index) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]
  title <- paste(outcome, data_mod)
  cat("#############################################################\n")
  print(title)
  cat("\n#############################################################\n")

  for (model in models) {
    print(model)
    print(cox.zph(model))
  }
}
```

#### Table coxzph Individual Cox Model Fits - Combined Cox

```{r table coxzph combined cox model, results = 'asis'}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  zph_tables <- lapply(seq_along(models), function(i) {
    formula <- Reduce(paste, deparse(models[[i]]$formula))

    table <- cox.zph(models[[i]])$table
    table <- data.frame(
      outcome = outcome, data_type = data_mod, Model = i, formula = formula,
      variables = rownames(table), table
    )
    rownames(table) <- NULL
    table$chisq <- round(table$chisq, 4)
    table$p <- round(table$p, 4)

    table
  })
  zph_tables <- do.call(rbind, zph_tables)

  zph_tables
})

coxzph_tables <- do.call(rbind, coxzph_tables)
kable(coxzph_tables)
```

#### Plot Shoenfeld Residuals - Combined Cox

```{r plot coxzph combined cox model residuals}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  lapply(seq_along(models), function(i) {
    model <- models[[i]]
    formula <- Reduce(paste, deparse(model$formula))

    coxzph <- cox.zph(model)
    plot(coxzph, main = formula)
  })
})
```


#### KM plots of covariates - complete data

```{r source surv analysis for covs, file = 'R/surv_all.R'}
source("R/surv_all.R")
```

```{r plot km covs}
census <- final_models[rest_index, "combined_census"][[1]]

cov_dummy <- config$datasets[[1]]$covariates_dummy
lapply(cov_dummy, function(cov) {
  census[[cov]] <- factor(census[[cov]])
  kap_plot_all(census, cov,
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )
})
```

### Figure `r as.character(fig_no)`: Histogram of Outcome Variables - Complete and Training Data

```{r disply hist outcome panel, fig.height=12, fig.width=12}
hist_outcome_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Histogram of log transform of outcome Variables - Complete and Training Data

```{r disply hist log outcome panel, fig.height=12, fig.width=12}
hist_log_outcome_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Histogram of probability of class membership - Complete Data

```{r disply hist class prob panel, fig.height=12, fig.width=12}
hist_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Display Testing Data

This is a check that we provided the correct validation data

#### Cumulative

```{r cumulative validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[1]]
plot_grid(plotlist = plots)
```

#### Window

```{r window validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[2]]
plot_grid(plotlist = plots)
```

#### Interval

```{r interval validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[3]]
plot_grid(plotlist = plots)
```

#### Proportion Resampled

```{r resampled validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[4]]
plot_grid(plotlist = plots)

fig_no <- fig_no + 1
```

