---
title: "SLAM Metabolic Class Figures Display"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  all_dir: "20220618_104254_all"
  mb6_dir: "20220617_045809_mb6"
  fb6_dir: "20220617_072941_fb6"
  mhet3_dir: "20220617_052502_mhet3"
  fhet3_dir: "20220617_033852_fhet3"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```

## Analysis

### Load Models

```{r load models}
output_path <- file.path("..", "05_figures", "output")

load_models <- function(subgroup, envir = .GlobalEnv) {
  subgroup_param <- paste(subgroup, "dir", sep = "_")
  subgroup_path <- params[[subgroup_param]]
  path <- file.path(output_path,
                    subgroup_path,
                    "data",
                    "final_models.RDATA")
  env_name <- paste(subgroup, "env", sep = "_")
  assign(env_name, rlang::env(), envir = envir)
  load(path, eval(as.symbol(env_name)))
}

subgroups <- c("all", "mb6", "fb6", "mhet3", "fhet3")

for(subgroup in subgroups) {
  load_models(subgroup)
}

# Add subgroup name to env
all_env$subgroup <- "All Mice"
mb6_env$subgroup <- "Male B6"
fb6_env$subgroup <- "Female B6"
mhet3_env$subgroup <- "Male HET3"
fhet3_env$subgroup <- "Female HET3"

```

## Figures

### Figure 1: Dispaly How latent class models work

```{r figure 1, fig.height=5, fig.width=10}
x <- all_env$final_models
p1 <- x[x$model_name == "bw_main|bw_age_wk", "spag_plot"][[1]]
p2 <- ggplot()
p3 <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]


cowplot::plot_grid(p1, p3, nrow = 1)
```

### Figure 2: Dispaly predicted and observed

- Supplement will include sex strain strata 

```{r figure 2, fig.height= 16, fig.width=12}

create_fig2 <- function(env) {
  x <- env$final_models
  bw_pred <- x[x$model_name == "bw_main|bw_age_wk", "traj_pred"][[1]]
  bw_obs <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]
  gluc_pred <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_pred"][[1]]  
  gluc_obs <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_obs"][[1]]
  fat_pred <- x[x$model_name == "fat_main|fat_age_wk", "traj_pred"][[1]]
  fat_obs <- x[x$model_name == "fat_main|fat_age_wk", "traj_obs"][[1]]
  
  bw_pred <- ggdraw(bw_pred, clip = "on")
  gluc_pred <- ggdraw(gluc_pred, clip = "on")
  fat_pred <- ggdraw(fat_pred, clip = "on")
  
  title_theme <- ggdraw() + 
    draw_label(paste(env$subgroup,
                     "Predicted and Observed Metabolic Life Course Trajectories", 
                     sep = " "))
  
  fig2 <- cowplot::plot_grid(bw_pred, bw_obs,
                   gluc_pred, gluc_obs,
                   fat_pred, fat_obs, 
                   nrow = 3,
                   greedy = FALSE,
                   labels = "AUTO"
                   )
  
  env$fig2 <- cowplot::plot_grid(title_theme, fig2, rel_heights = c(.05, 1),
                                 ncol = 1)
  
  env
  
}

env_list <- list(all_env, mb6_env, fb6_env, mhet3_env, fhet3_env)

env_list <- lapply(env_list, create_fig2)

lapply(env_list, `[[`, "fig2")

```

### Table 1: Who's in what class

```{r table 1, results='asis'}

mo_names <- c("bw_main|bw_age_wk", "gluc_main|gluc_age_wk", "fat_main|fat_age_wk")
create_t1 <- function(env, mo_names) {
  x <- env$final_models
  t1_list <- lapply(mo_names, function(name) {
    x[x$model_name == name, ]$clean_t1[[1]]
  })
  
  outcome_names <- lapply(mo_names, function(name) {
    x[x$model_name == name, ]$oc_name
  })
  
  t1_list <- lapply(seq_along(t1_list), function(i) {
    t1 <- t1_list[[i]]
    oc_name <- outcome_names[[i]]
    
    
   if(!is.na(t1[[1]])) {
     t1 <- t1
    t1 <- as.data.frame(t(t1))
    t1$row_names <- row.names(t1)
    row.names(t1) <- NULL
    t1$oc_name <- oc_name
   } else {
     t1 <- NULL
   }
    t1
  })
  
  t1_df <- do.call(rbind, t1_list) %>%
    select(oc_name, row_names, everything())
  
  env$t1_df <- t1_df
  
  env
}

env_list <- lapply(env_list, function(env) {
  create_t1(env, mo_names)
})

lapply(env_list, function(env) {
  uni_models <- unique(env$t1_df$oc_name)
  j <- 0
  no_rows <- list()
  for (i in seq_along(uni_models)) {
    number_rows <- sum(env$t1_df$oc_name == uni_models[[i]])
    no_rows[[i]] <- number_rows + j
    j <- number_rows
  }
  table <- kable(env$t1_df, format = "html") %>%
    column_spec(1, bold = TRUE, border_right = TRUE) %>%
    row_spec(no_rows[[1]], hline_after = T) %>%
    row_spec(no_rows[[2]], hline_after = T) %>%
    row_spec(no_rows[[3]], hline_after = T) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    kable_styling(bootstrap_options = "striped")
  
  print(table)
    
})
```

### Figure 3: Mortality outcomes by class

- Supplement will include same figures but for each sex strain strata

```{r figure 3, fig.height= 16, fig.width=12}

create_fig3 <- function(env){
  x <- env$final_models
  bw_obs <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]
  gluc_obs <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_obs"][[1]]
  fat_obs <- x[x$model_name == "fat_main|fat_age_wk", "traj_obs"][[1]]
  bw_kap <- x[x$model_name == "bw_main|bw_age_wk", "kap_plot"][[1]]
  gluc_kap <- x[x$model_name == "gluc_main|gluc_age_wk", "kap_plot"][[1]]  
  fat_kap <- x[x$model_name == "fat_main|fat_age_wk", "kap_plot"][[1]]   
  
  if (!is.na(bw_kap)) bw_kap <- bw_kap$plot
  if (!is.na(gluc_kap)) gluc_kap <- gluc_kap$plot
  if (!is.na(fat_kap)) fat_kap <- fat_kap$plot

  fig3 <- cowplot::plot_grid(bw_obs, bw_kap,
                   gluc_obs, gluc_kap,
                   fat_obs, fat_kap, 
                   nrow = 3,
                   greedy = FALSE,
                   labels = "AUTO")
  
  title_theme <- ggdraw() + 
    draw_label(paste(env$subgroup,
                     "Observed Metabolic Life Course Trajectories and Kaplan Meiers by Class", 
                     sep = " "))
  
  env$fig3 <- cowplot::plot_grid(title_theme, fig3, rel_heights = c(.05, 1),
                                 ncol = 1)
  
  env
}
env_list <- lapply(env_list, create_fig3)

lapply(env_list, `[[`, "fig3")
```

### Table 2: Hazard Ratios Adjusted for Sex Strain and proportional hazards

```{r table 2}

```

### Figure 4: Metabolic Outcomes by Class

- Supplement will show the same figures but fore each sex strain combo

```{r figure 4, fig.height=16, fig.width=24}

create_fig4 <- function(env) {
  x <- env$final_models
  bw_pred <- x[x$model_name == "bw_main|bw_age_wk", "traj_pred"][[1]]
  gluc_pred <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_pred"][[1]]  
  fat_pred <- x[x$model_name == "fat_main|fat_age_wk", "traj_pred"][[1]]
  
  bw_pred <- ggdraw(bw_pred, clip = "on")
  gluc_pred <- ggdraw(gluc_pred, clip = "on")
  fat_pred <- ggdraw(fat_pred, clip = "on")
  
  bw_bw <- x[x$model_name == "bw_main|bw_age_wk", "plot_bw"][[1]]
  gluc_bw <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_bw"][[1]]  
  fat_bw <- x[x$model_name == "fat_main|fat_age_wk", "plot_bw"][[1]]
  
  bw_gluc <- x[x$model_name == "bw_main|bw_age_wk", "plot_gluc"][[1]]
  gluc_gluc <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_gluc"][[1]]  
  fat_gluc <- x[x$model_name == "fat_main|fat_age_wk", "plot_gluc"][[1]]
  
  bw_fat <- x[x$model_name == "bw_main|bw_age_wk", "plot_fat"][[1]]
  gluc_fat <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_fat"][[1]]  
  fat_fat <- x[x$model_name == "fat_main|fat_age_wk", "plot_fat"][[1]]
  
  fig4 <- cowplot::plot_grid(bw_pred, bw_bw, bw_gluc, bw_fat,
                   gluc_pred, gluc_bw, gluc_gluc, gluc_fat,
                   fat_pred, fat_bw, fat_gluc, fat_fat,
                   nrow = 3,
                   greedy = FALSE,
                   labels = "AUTO")
  
  title_theme <- ggdraw() + 
    draw_label(paste(env$subgroup,
                     "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
                     sep = " "))
    
  env$fig4 <- cowplot::plot_grid(title_theme, fig4, rel_heights = c(.05, 1),
                                 ncol = 1)
  
  env
  
}

  
env_list <- lapply(env_list, create_fig4)

lapply(env_list, `[[`, "fig4")
```

### Figure 5: Class Prediction 

- Supplmenet will include same table for glucose and body fat

```{r figure 5, fig.height=16, fig.width=16}
bw_interval <- x[x$model_name == "bw_main|bw_age_wk", "interval_plots"][[1]]
bw_threshold <- x[x$model_name == "bw_main|bw_age_wk", "threshold_plots"][[1]]
bw_sample <- x[x$model_name == "bw_main|bw_age_wk", "sample_plots"][[1]]

fig5 <- cowplot::plot_grid(bw_interval, bw_threshold, bw_sample,
                           ncol = 3,
                           labels = "AUTO"
                           )

fig5
```