---
title: "SLAM Metabolic Class Figures Display"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  all_dir: "20220618_104254_all"
  mb6_dir: "20220617_045809_mb6"
  fb6_dir: "20220617_072941_fb6"
  mhet3_dir: "20220617_052502_mhet3"
  fhet3_dir: "20220617_033852_fhet3"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```

## Analysis

### Load Models

```{r load models}
output_path <- file.path("..", "05_figures", "output")

load_models <- function(subgroup, envir = .GlobalEnv) {
  subgroup_param <- paste(subgroup, "dir", sep = "_")
  subgroup_path <- params[[subgroup_param]]
  path <- file.path(
    output_path,
    subgroup_path,
    "data",
    "final_models.RDATA"
  )
  env_name <- paste(subgroup, "env", sep = "_")
  assign(env_name, rlang::env(), envir = envir)
  load(path, eval(as.symbol(env_name)))
}

subgroups <- c("all", "mb6", "fb6", "mhet3", "fhet3")

for (subgroup in subgroups) {
  load_models(subgroup)
}

# Add subgroup name to env
all_env$subgroup <- "All Mice"
mb6_env$subgroup <- "Male B6"
fb6_env$subgroup <- "Female B6"
mhet3_env$subgroup <- "Male HET3"
fhet3_env$subgroup <- "Female HET3"
```

## Figures

### Figure 1: Dispaly How latent class models work

```{r figure 1, fig.height=5, fig.width=10}
x <- all_env$final_models
p1 <- x[x$model_name == "bw_main|bw_age_wk", "spag_plot"][[1]]
p2 <- ggplot()
p3 <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]


cowplot::plot_grid(p1, p3, nrow = 1)
```

### Figure 2: Dispaly predicted and observed

- Supplement will include sex strain strata 

```{r figure 2, fig.height= 16, fig.width=12}

create_pred_obs_panel <- function(env) {
  x <- env$final_models
  bw_pred <- x[x$model_name == "bw_main|bw_age_wk", "traj_pred"][[1]]
  bw_obs <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]
  gluc_pred <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_pred"][[1]]
  gluc_obs <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_obs"][[1]]
  fat_pred <- x[x$model_name == "fat_main|fat_age_wk", "traj_pred"][[1]]
  fat_obs <- x[x$model_name == "fat_main|fat_age_wk", "traj_obs"][[1]]

  bw_pred <- ggdraw(bw_pred, clip = "on")
  gluc_pred <- ggdraw(gluc_pred, clip = "on")
  fat_pred <- ggdraw(fat_pred, clip = "on")

  title_theme <- ggdraw() +
    draw_label(paste(env$subgroup,
      "Predicted and Observed Metabolic Life Course Trajectories",
      sep = " "
    ))

  pred_obs_panel <- cowplot::plot_grid(bw_pred, bw_obs,
    gluc_pred, gluc_obs,
    fat_pred, fat_obs,
    nrow = 3,
    greedy = FALSE,
    labels = "AUTO"
  )

  env$pred_obs_panel <- cowplot::plot_grid(title_theme, pred_obs_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )

  env
}

env_list <- list(all_env, mb6_env, fb6_env, mhet3_env, fhet3_env)

env_list <- lapply(env_list, create_pred_obs_panel)

lapply(env_list, `[[`, "pred_obs_panel")
```

### Table 1: Who's in what class

```{r table 1, results='asis'}

mo_names <- c("bw_main|bw_age_wk", "gluc_main|gluc_age_wk", "fat_main|fat_age_wk")
create_t1 <- function(env, mo_names) {
  x <- env$final_models
  t1_list <- lapply(mo_names, function(name) {
    x[x$model_name == name, ]$clean_t1[[1]]
  })

  outcome_names <- lapply(mo_names, function(name) {
    x[x$model_name == name, ]$oc_name
  })

  t1_list <- lapply(seq_along(t1_list), function(i) {
    t1 <- t1_list[[i]]
    oc_name <- outcome_names[[i]]


    if (!is.na(t1[[1]])) {
      t1 <- t1
      t1 <- as.data.frame(t(t1))
      t1$row_names <- row.names(t1)
      row.names(t1) <- NULL
      t1$oc_name <- oc_name
    } else {
      t1 <- NULL
    }
    t1
  })

  t1_df <- do.call(rbind, t1_list) %>%
    select(oc_name, row_names, everything())


  env$t1_df <- t1_df

  env
}

env_list <- lapply(env_list, function(env) {
  create_t1(env, mo_names)
})

lapply(env_list, function(env) {
  uni_models <- unique(env$t1_df$oc_name)
  j <- 0
  no_rows <- list()
  for (i in seq_along(uni_models)) {
    number_rows <- sum(env$t1_df$oc_name == uni_models[[i]])
    no_rows[[i]] <- number_rows + j
    j <- number_rows
  }
  cap <- paste(env$subgroup,
    "Demographics by Metabolic Index and Class",
    sep = " "
  )
  table <- kable(env$t1_df,
    format = "html",
    caption = cap
  ) %>%
    column_spec(1:2, bold = TRUE, border_right = TRUE) %>%
    row_spec(no_rows[[1]], hline_after = T) %>%
    row_spec(no_rows[[2]], hline_after = T) %>%
    row_spec(no_rows[[3]], hline_after = T) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    kable_styling(bootstrap_options = "striped")

  print(table)
})
```


### Mortality outcomes by class

- Supplement will include same figures but for each sex strain strata

```{r figure 3, fig.height= 16, fig.width=12}

create_mortality_panel <- function(env) {
  x <- env$final_models
  bw_obs <- x[x$model_name == "bw_main|bw_age_wk", "traj_obs"][[1]]
  gluc_obs <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_obs"][[1]]
  fat_obs <- x[x$model_name == "fat_main|fat_age_wk", "traj_obs"][[1]]
  bw_kap <- x[x$model_name == "bw_main|bw_age_wk", "kap_plot"][[1]]
  gluc_kap <- x[x$model_name == "gluc_main|gluc_age_wk", "kap_plot"][[1]]
  fat_kap <- x[x$model_name == "fat_main|fat_age_wk", "kap_plot"][[1]]

  if (!is.na(bw_kap)) bw_kap <- bw_kap$plot
  if (!is.na(gluc_kap)) gluc_kap <- gluc_kap$plot
  if (!is.na(fat_kap)) fat_kap <- fat_kap$plot

  mortality_panel <- cowplot::plot_grid(bw_obs, bw_kap,
    gluc_obs, gluc_kap,
    fat_obs, fat_kap,
    nrow = 3,
    greedy = FALSE,
    labels = "AUTO"
  )

  title_theme <- ggdraw() +
    draw_label(paste(env$subgroup,
      "Observed Metabolic Life Course Trajectories and Kaplan Meiers by Class",
      sep = " "
    ))

  env$mortality_panel <- cowplot::plot_grid(title_theme, mortality_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )

  env
}
env_list <- lapply(env_list, create_mortality_panel)

lapply(env_list, `[[`, "mortality_panel")
```

### Table of Hazard Ratios, Raw and Sex/Strain Adjusted

```{r hazard ratio table}

```

### Metabolic Outcomes by Class

- Supplement will show the same figures but fore each sex strain combo

```{r figure 4, fig.height=16, fig.width=24}

create_other_met_panel <- function(env) {
  x <- env$final_models
  bw_pred <- x[x$model_name == "bw_main|bw_age_wk", "traj_pred"][[1]]
  gluc_pred <- x[x$model_name == "gluc_main|gluc_age_wk", "traj_pred"][[1]]
  fat_pred <- x[x$model_name == "fat_main|fat_age_wk", "traj_pred"][[1]]

  bw_pred <- ggdraw(bw_pred, clip = "on")
  gluc_pred <- ggdraw(gluc_pred, clip = "on")
  fat_pred <- ggdraw(fat_pred, clip = "on")

  bw_bw <- x[x$model_name == "bw_main|bw_age_wk", "plot_bw"][[1]]
  gluc_bw <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_bw"][[1]]
  fat_bw <- x[x$model_name == "fat_main|fat_age_wk", "plot_bw"][[1]]

  bw_gluc <- x[x$model_name == "bw_main|bw_age_wk", "plot_gluc"][[1]]
  gluc_gluc <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_gluc"][[1]]
  fat_gluc <- x[x$model_name == "fat_main|fat_age_wk", "plot_gluc"][[1]]

  bw_fat <- x[x$model_name == "bw_main|bw_age_wk", "plot_fat"][[1]]
  gluc_fat <- x[x$model_name == "gluc_main|gluc_age_wk", "plot_fat"][[1]]
  fat_fat <- x[x$model_name == "fat_main|fat_age_wk", "plot_fat"][[1]]

  other_met_panel <- cowplot::plot_grid(bw_pred, bw_bw, bw_gluc, bw_fat,
    gluc_pred, gluc_bw, gluc_gluc, gluc_fat,
    fat_pred, fat_bw, fat_gluc, fat_fat,
    nrow = 3,
    greedy = FALSE,
    labels = "AUTO"
  )

  title_theme <- ggdraw() +
    draw_label(paste(env$subgroup,
      "Cross-reference Other Metabolic Parameters by Each Metabolic Class",
      sep = " "
    ))

  env$other_met_panel <- cowplot::plot_grid(title_theme, other_met_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )

  env
}


env_list <- lapply(env_list, create_other_met_panel)

lapply(env_list, `[[`, "other_met_panel")
```

### Table of lme models
- Model1 (No interaction): $outcome = Age + Age^2 + Class + (Age + Age^2 | Subject)$
- Model2 : $outcome = Age + Age^2 + Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model3 (no controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2| Subject)$
- Model4 (Sex and Strain additive controlled) : $outcome = (Age + Age^2) * Class + Female * HET3 + (Age + Age^2 | Subject)$
- Model5 (Sex and Strain interacted controlled): $outcome = (Age + Age^2) * Class + (Age + Age^2) * Female * HET3 + (Age + Age^2 | Subject)$
```{r table lme}

tab_lme <- function(env, mo_names) {
  x <- env$final_models
  row <- x[x$model_name %in% mo_names, ]
  
  age_var <- row$age_var[[1]]
  age_var_name <- row$age_var_name[[1]]
  age_var_units <- row$age_var_units[[1]]
  
  age_var2_name <- paste(age_var_name, "2", sep = "^")
  age_var2_units <- paste(age_var_units, "2", sep = "^")
  
  age_var2 <- paste(age_var, 2, sep = "")
  age_var_ns <- paste(age_var, "ns", sep = "_")
  age_var2_ns <- paste(age_var2, "ns", sep = "_")
  class_names <- paste("Class", seq_len(100))
  names(class_names) <- paste0("new_class", seq_len(100))
  age_names <- c(age_var_ns = paste(age_var_name, age_var_units))
  age2_names <- c(age_var2_ns = paste(age_var2_name, age_var2_units))
  age_interaction_names <- paste(age_names, class_names, sep = ":")
  names(age_interaction_names) <- paste(names(age_names),
                                        names(class_names),
                                        sep = ":")
  age2_interaction_names <- paste(age_var2_ns, class_names, sep = ":")
  names(age2_interaction_names) <- paste(names(age2_names), names(class_names), sep = ":")
  fixcov_names <- c("female" = "Female", "HET3" = "het3")
  
  all_names <- c(class_names, age_names, age2_names, age_interaction_names, 
                 age2_interaction_names, fixcov_names)
    
  lmes_bw <- row$model_bw
  lmes_gluc <- row$model_gluc
  lmes_fat <- row$model_fat
  
  model_list <- list()
  oc_names_list <- list()
  
  for (i in seq_len(nrow(row))) {
    model_list[[i]] <- list(lmes_bw[[i]][[4]],
                            lmes_gluc[[i]][[4]],
                            lmes_fat[[i]][[4]])
    oc_names_list[[i]] <- c("Body Weight", 
                         "Body Fat", 
                         "Glucose")
    
    
  }
  model_list <- do.call(c, model_list)
  oc_names_list <- do.call(c, oc_names_list)
  
  title <- paste(env$subgroup, "Mixed Effects Models")
  lme_table <- tab_model(model_list,
                 show.ci = FALSE,
                 show.re.var = FALSE,
                 pred.labels = all_names,
                 show.reflvl = TRUE,
                 show.icc = FALSE,
                 show.r2 = FALSE,
                 show.ngroups = FALSE,
                 show.obs = FALSE,
                 dv.labels = oc_names_list,
                 title = title)
  
  env$lme_table <- lme_table
  
  env
}

env_list <-lapply(env_list, tab_lme, mo_names = mo_names)

lapply(env_list, `[[`, "lme_table")


```

### Figure 5: Class Prediction 

- Supplmenet will include same table for glucose and body fat

```{r figure 5, fig.height=16, fig.width=16}
mo_names <- c("bw_main|bw_age_wk", "gluc_main|gluc_age_wk", "fat_main|fat_age_wk")

create_class_prediction_panel <- function(env, mo_names) {
  x <- env$final_models
  rows <- list()
  for (model_name in mo_names) {
    interval <- x[x$model_name == model_name, ]$interval_boot_accuracy_plot[[1]]
    threshold <- x[x$model_name == model_name, ]$threshold_boot_accuracy_plot[[1]] +
      theme(legend.title = element_text(size = 7), 
               legend.text = element_text(size = 7))
    sample <- x[x$model_name == model_name, ]$sample_boot_accuracy_plot[[1]]

    rows[[model_name]] <- list(interval, threshold, sample)
  }
  
  rows <- do.call(c, rows)
  
  title <- paste(env$subgroup, "Classification Accuracy")
  
  title_theme <- ggdraw() +
    draw_label(title)


  class_prediction_panel <- cowplot::plot_grid(
    plotlist = rows,
    ncol = 3,
    labels = "AUTO"
  )
  env$class_prediction_panel <- cowplot::plot_grid(title_theme, 
                                                   class_prediction_panel,
    rel_heights = c(.05, 1),
    ncol = 1
  )
  
  
  env
}

env_list <- lapply(env_list, create_class_prediction_panel, mo_names = mo_names)

lapply(env_list, function(env) {
  env$class_prediction_panel
})
```
