---
title: "SLAM Metabolic Class Supplemental Tables - `r config$out_tag` Mice"
subtitle: "External Validation: `r config$external_validate`"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../06_create_figures/output/test_local.yaml"
---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)

window_legend <- TRUE
window_legend_palette <- magma(3, begin = .2, end = .7, direction = -1)
window_legend_labels <- c("2 measures", "3 measures", "4 measures")

cumulative_legend <- TRUE
cumulative_legend_palette <- viridis(3, begin = .2, end = .7, direction = -1)

max_age <- 150
max_age_km <- NA
```

## Analysis

### Load Input File

```{r read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  # params$input_path <- "../06_create_figures/output/test_local.yaml"
  # params$input_path <- "../06_create_figures/output/all_only_bw.yaml"
  # params$input_path <- "../06_create_figures/output/20250306_itp.yaml"
  # params$input_path <- "../06_create_figures/output/all.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_itp_C2010_C2011_C2013_C2016_controls_cohort.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_all_cohort.yaml"
  params$input_path <- "../06_create_figures/output/slam_c1-c10_age_fb6_bwfatgluc.yaml"

  # params$input_path <- "../06_create_figures/output/all_lean.yaml"
  # params$input_path <- "../06_create_figures/output/fb6.yaml"
  # params$input_path <- "../06_create_figures/output/external_validation.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
```

### Load Models

```{r load models}
load(input$datasets_path)
load(input$final_models_path)
config <- yaml::read_yaml(input$config_path)


# load in main_cat_surv
main_cat_surv <- read.csv(file = config$survival_dataset$path)
main_cat_surv <- as.data.frame(main_cat_surv)

final_models$dfs <- lapply(final_models$data, function(name) {
  datasets[[name]]$data
})
```

```{r initiate figure and table number}
fig_no <- 1
tab_no <- 1
tabs <- list()
```

### Supplemental Table `r as.character(tab_no)`: Descriptive Statistics of Outcomes, Age, Measure Frequency, and Age Difference Between measures for complete, training, and testing datasets

```{r describe prob class, results = 'asis'}
rest_index <- final_models$data_mod == "og"
census <- final_models[rest_index, "combined_census"][[1]]
tabs$combined_census <- kable(psych::describe(census))
tabs$combined_census
```


## Display Dataset/Model informaton 

Do this for complete dataset, training dataset, testing dataset.

- Number of observations per outcome
- Number of subjects per outcome
- Number of sujbects excluded for completeness

```{r dataset information, results = 'asis'}
make_wave <- function(data, id, age_var) {
  data <- data %>%
    dplyr::group_by(!!sym(id)) %>%
    dplyr::arrange(!!sym(age_var)) %>%
    dplyr::mutate(wave = dplyr::row_number()) %>%
    dplyr::mutate(dif_time = !!sym(age_var) - lag(!!sym(age_var))) %>%
    dplyr::ungroup() %>%
    as.data.frame()

  data
}

summarise_dataset <- function(dataset, data_name) {
  data <- dataset[[data_name]]
  if (is.null(data)) {
    return(NULL)
  }
  data_id <- dataset$data_id
  data_mod <- dataset$data_mod
  if (data_mod == "train_test" && data_name == "data") {
    data_name <- "training_data"
  }

  id <- dataset$id
  age <- paste0(dataset$age[[1]], "_ns")
  if (data_name == "missing_data") {
    age <- dataset$age[[1]]
  }
  outcome <- dataset$outcome
  unique_id <- length(unique(data[[id]]))
  data <- make_wave(data, id, age)
  sum <- round(psych::describe(data[, c(outcome, age, "wave", "dif_time")]), 2)
  rownames(sum) <- c(outcome, age, "wave", "dif_time")
  sum <- cbind(rownames(sum), data.frame(sum, row.names = NULL))
  names(sum)[1] <- "variables"
  sum <- data.frame(unique_id = rep(unique_id, nrow(sum)), sum)
  sum <- data.frame(outcome = rep(outcome, nrow(sum)), sum)
  sum <- data.frame(data_id = rep(data_id, nrow(sum)), sum)
  sum <- data.frame(data_name = rep(data_name, nrow(sum)), sum)
  sum
}


sum_data <- do.call(rbind, lapply(datasets, summarise_dataset, "data"))
rownames(sum_data) <- NULL
sum_missing <- do.call(
  rbind,
  lapply(datasets, summarise_dataset, "missing_data")
)
rownames(sum_missing) <- NULL
sum_test <- do.call(rbind, lapply(datasets, summarise_dataset, "test_data"))
rownames(sum_test) <- NULL
```

#### Complete and Training Data

```{r dataset info complete and training, results = 'asis'}
kable(sum_data)
```

#### Missing Data from Complete and Training Data

```{r dataset info missing complete and training, results = 'asis'}
kable(sum_missing)
```

#### Testing Data

```{r dataset info testing data, results = 'asis'}
kable(sum_test)
```

### Census information

- Study Demographics
- Median Life Expectancy

```{r census information, results = 'asis'}
describe_le <- function(data, age_death, event) {
  surv_object <- survival::Surv(time = data[[age_death]], event = data[[event]])
  surv_fit <- survival::survfit(surv_object ~ 1)
  quant_surv <- quantile(surv_fit)
  names(quant_surv$quantile) <- c("Q1", "Median", "Q3")
  names(quant_surv$lower) <- c("Lower Q1", "Lower Median", "Lower Q3")
  names(quant_surv$upper) <- c("Upper Q1", "Upper Median", "Upper Q3")
  quant_surv <- lapply(quant_surv, t)
  quant_surv <- lapply(quant_surv, as.data.frame)
  names(quant_surv) <- NULL
  quant_surv <- do.call(cbind, quant_surv)
  quant_surv <- data.frame(n = surv_fit$n, quant_surv)
  quant_surv$conf_int <- surv_fit$conf.int
  quant_surv$conf_type <- surv_fit$conf.type
  quant_surv$type <- surv_fit$type
  quant_surv
}
census_complete <- final_models$combined_census[[1]]
age_death <- config$survival_dataset$age_death
event <- config$survival_dataset$event
filter_vars <- config$meta_dataset$covariates
filtered_dfs <- lapply(filter_vars, function(filter) {
  print(filter)
  unique_values <- unique(census_complete[[filter]])

  filtered_dfs <- lapply(unique_values, function(value) {
    print(value)
    census_complete[census_complete[[filter]] == value, ]
  })
  names(filtered_dfs) <- unique_values
  filtered_dfs
})
names(filtered_dfs) <- filter_vars
filtered_dfs <- unlist(filtered_dfs, recursive = FALSE)

data <- c(list(all = census_complete), filtered_dfs)
# peek(data)

table_mle <- lapply(data, describe_le, age_death, event)
table_mle <- do.call(rbind, table_mle)

kable(table_mle)
```

### LCMM information 

Include latent class model fit statistics

#### LCMM - Concise Table

```{r lcmm concise, results = 'asis'}
table <- helphlme::compare_models(final_models$final_model)$table
table <- data.frame(
  outcome = final_models$oc,
  data_type = final_models$data_mod,
  table
)
kable(table)
```

#### LCMM - Verbose Table

```{r lcmm model information}
for (i in seq_len(nrow(final_models))) {
  label <- paste(final_models$oc[[i]], final_models$data_mod[[i]])
  cat("#############################################################\n")
  cat(label)
  cat("\n#############################################################\n")
  model <- final_models$final_model[[i]]
  summary(model)
  lcmm::postprob(model)
}
```

#### LCMM - Average Posterior Classifcation Probabilites

```{r lcmm model average posterior probs, results = 'asis'}
post_prob_table <- lapply(seq_len(nrow(final_models)), function(i) {
  label <- paste(final_models$oc[[i]], final_models$data_mod[[i]])
  outcome <- final_models$oc[[i]]
  data_type <- final_models$data_mod[[i]]
  model <- final_models$final_model[[i]]
  post_prob <- lcmm::postprob(model)
  if (isTRUE(all.equal(post_prob, list()))) {
    warning("Cannot call lcmm::postprob on model where ng = 1, returning NAs")
    post_prob_table <- data.frame(
      outcome = outcome,
      data_type = data_type,
      label = label,
      class = NA,
      check.names = FALSE,
      row.names = NULL
    )
    return(post_prob_table)
  }
  post_prob_table <- data.frame(
    outcome = outcome,
    data_type = data_type,
    label = label,
    class = row.names(post_prob[[2]]),
    t(post_prob[[1]]),
    post_prob[[2]],
    t(post_prob[[3]]),
    check.names = FALSE,
    row.names = NULL
  )
  post_prob_table
})
post_prob_table <- data.table::rbindlist(post_prob_table, fill = TRUE)
post_prob_table <- as.data.frame(post_prob_table)
kable(post_prob_table)
```

#### LCMM - Model residual plots

```{r plot model residuals, fig.height = 12, fig.width = 12}
for (i in seq_len(nrow(final_models))) {
  title <- paste(final_models$data_mod[[i]], final_models$oc[[i]])
  model <- final_models$final_model[[i]]
  cat("##### ")
  cat(title)
  cat("\n")

  plot(model, which = "residuals", var.time = "age_wk_ns", title = title)
}
```

#### LCMM - Compare all possible models

```{r compare all models, results = 'asis'}
load(input$models_path)
load(input$cf_path)
all_models <- helphlme::compare_models(models)$table
all_models <- data.frame(
  model_number = all_models[, 1],
  data = cf$data,
  fixed_effects = cf$fixed,
  mixture = cf$mixture,
  all_models[, 2:length(all_models)],
  check.names = FALSE
)
all_models$`Smallest Class (%)` <- round(all_models$`Smallest Class (%)`, 2)
kable(all_models)
```

### Chi-squared information

```{r chi squared information }
for (i in seq_len(nrow(final_models))) {
  title <- paste(final_models$data_mod[[i]], final_models$oc[[i]])
  chi_test <- final_models$chi_test[[i]]
  cat("#############################################################\n")
  cat(title)
  cat("\n#############################################################\n")
  print(chi_test)
}
```

### LME information

- Selected model - `r model_display`

```{r lme information}
for (models in model_list) {
  for (model in models) {
    cat("#############################################################\n")
    print(model@call)
    cat("\n#############################################################\n")
    print(model)
    print(summary(model))
    print(plot(model))
  }
}
```

### LME table

- Selected model - `r model_display`

```{r lme table}
table_list <- lapply(model_list, function(models) {
  lapply(models, function(model) {
    model_sum <- summary(model)
    data.frame(
      call = as.character(model@call)[2],
      title = model_sum$methTitle,
      loglik = model_sum$logLik,
      n_groups = model_sum$n,
      sigma = model_sum$sigma,
      AICtab = model_sum$AICtab,
      optimizer = model@optinfo$optimizer,
      conv = model@optinfo$conv$opt
    )
  })
})

table_list <- unlist(table_list, recursive = FALSE)
lme_table <- do.call(rbind, table_list)
kable(lme_table)
```

### Individual Cox Model

#### print model output and print cox.zph output

```{r individual cox model}
rest_index <- which(final_models$data_mod == "og")
for (i in rest_index) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]
  title <- paste(outcome, data_mod)
  cat("#############################################################\n")
  print(title)
  cat("\n#############################################################\n")

  for (model in models) {
    if (is.null(model)) {
      message("model is NA, will not print anything")
    } else {
      print(model)
      print(cox.zph(model))
    }
  }
}
```

#### Table coxzph Individual Cox Model Fits

```{r table coxzph individual cox model, results = 'asis'}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  zph_tables <- lapply(seq_along(models), function(i) {
    formula <- Reduce(paste, deparse(models[[i]]$formula))

    if (is.null(models[[i]])) {
      message("model is NULL, returning NA values in table")
      table <- data.frame(chisq = NA, df = NA, p = NA)
      table <- data.frame(
        outcome = outcome, data_type = data_mod,
        Model = i, formula = formula, variables = rownames(table), table
      )
      rownames(table) <- NULL
      return(table)
    }

    table <- cox.zph(models[[i]])$table
    table <- data.frame(
      outcome = outcome, data_type = data_mod, Model = i, formula = formula,
      variables = rownames(table), table
    )
    rownames(table) <- NULL
    table$chisq <- round(table$chisq, 4)
    table$p <- round(table$p, 4)

    table
  })
  zph_tables <- do.call(rbind, zph_tables)

  zph_tables
})

coxzph_tables <- do.call(rbind, coxzph_tables)
kable(coxzph_tables)
```

#### Plot Shoenfeld Residuals

```{r plot coxzph individual cox model residuals}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  lapply(seq_along(models), function(i) {
    model <- models[[i]]
    if (is.null(model)) {
      warning("Model is NULL, returning NULL")
      return(NULL)
    }
    formula <- Reduce(paste, deparse(model$formula))

    coxzph <- cox.zph(model)
    plot(coxzph, main = formula)
  })
})
```

### Combined Cox Model

#### print model output and print cox.zph output - Combined Cox

```{r combined cox model print output}
rest_index <- which(final_models$data_mod == "og")
for (i in rest_index) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]
  title <- paste(outcome, data_mod)
  cat("#############################################################\n")
  print(title)
  cat("\n#############################################################\n")

  for (model in models) {
    tt_terms <- sum(grepl("tt\\(", names(attr(model$terms, "dataClasses"))))
    print(model)
    if (!tt_terms > 0) {
      print(cox.zph(model))
    }
  }
}
```

#### Table coxzph Individual Cox Model Fits - Combined Cox

```{r table coxzph combined cox model, results = 'asis'}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  zph_tables <- lapply(seq_along(models), function(i) {
    formula <- Reduce(paste, deparse(models[[i]]$formula))
    model <- models[[i]]

    tt_terms <- sum(grepl("tt\\(", names(attr(model$terms, "dataClasses"))))
    if (tt_terms > 0) {
      table <- data.frame(
        outcome = outcome, data_type = data_mod, Model = i, formula = formula,
        variables = NA, chisq = NA, df = NA, p = NA
      )
      return(table)
    }

    table <- cox.zph(models[[i]])$table
    table <- data.frame(
      outcome = outcome, data_type = data_mod, Model = i, formula = formula,
      variables = rownames(table), table
    )
    rownames(table) <- NULL
    table$chisq <- round(table$chisq, 4)
    table$p <- round(table$p, 4)

    table
  })
  zph_tables <- do.call(rbind, zph_tables)

  zph_tables
})

coxzph_tables <- do.call(rbind, coxzph_tables)
kable(coxzph_tables)
```

#### Plot Shoenfeld Residuals - Combined Cox

```{r plot coxzph combined cox model residuals}
rest_index <- which(final_models$data_mod == "og")
coxzph_tables <- lapply(rest_index, function(i) {
  models <- final_models[i, "cox_combined_models"][[1]]
  outcome <- final_models[i, "oc"]
  data_mod <- final_models[i, "data_mod"]

  lapply(seq_along(models), function(i) {
    model <- models[[i]]
    formula <- Reduce(paste, deparse(model$formula))

    tt_terms <- sum(grepl("tt\\(", names(attr(model$terms, "dataClasses"))))
    if (tt_terms > 0) {
      message("Cannot plot residuals for models with tt terms, returning NA")
      return(NA)
    }
    coxzph <- cox.zph(model)
    plot(coxzph, main = formula)
  })
})
```


#### KM plots of covariates - complete data

```{r source surv analysis for covs, file = 'R/surv_all.R'}
source("R/surv_all.R")
```

```{r plot km covs}
census <- final_models[rest_index, "combined_census"][[1]]

cov_dummy <- datasets[[1]]$covariates_dummy
lapply(cov_dummy, function(cov) {
  census[[cov]] <- factor(census[[cov]])
  kap_plot_all(census, cov,
    age_death = config$survival_dataset$age_death,
    event = config$survival_dataset$event
  )
})
```

### Figure `r as.character(fig_no)`: Histogram of Outcome Variables - Complete and Training Data

```{r disply hist outcome panel, fig.height=12, fig.width=12}
hist_outcome_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Histogram of log transform of outcome Variables - Complete and Training Data

```{r disply hist log outcome panel, fig.height=12, fig.width=12}
hist_log_outcome_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Histogram of probability of class membership - Complete Data

```{r disply hist class prob panel, fig.height=12, fig.width=12}
hist_plot_panel
fig_no <- fig_no + 1
```

### Figure `r as.character(fig_no)`: Display Testing Data

This is a check that we provided the correct validation data

#### Cumulative

```{r cumulative validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[1]]
plot_grid(plotlist = plots)
```

#### Window

```{r window validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[2]]
plot_grid(plotlist = plots)
```

#### Interval

```{r interval validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[3]]
plot_grid(plotlist = plots)
```

#### Proportion Resampled

```{r resampled validation plots, fig.width = 36, fig.height = 36}
plots <- datasets[[1]]$validation_plots[[4]]
plot_grid(plotlist = plots)

fig_no <- fig_no + 1
```

