---
title: "SLAM Metabolic Class Supplemental Tables - `r config$out_tag` Mice"
subtitle: "External Validation: `r config$external_validate`"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../07_display_figures/output/slam_c1-c10_age_all_bwfatgluc.yaml"
---

```{r sup_tables setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```

```{r sup_tables read input yaml}
# for the case when we are not running with RStudio
if (!exists("params")) {
  params <- list()
  # params$input_path <- "../06_create_figures/output/test_local.yaml"
  # params$input_path <- "../06_create_figures/output/all_only_bw.yaml"
  # params$input_path <- "../06_create_figures/output/20250306_itp.yaml"
  # params$input_path <- "../06_create_figures/output/all.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_itp_C2010_C2011_C2013_C2016_controls_cohort.yaml"
  # params$input_path <- "../06_create_figures/output/20250426_all_cohort.yaml"
  params$input_path <- "../06_create_figures/output/slam_c1-c10_age_fb6_bwfatgluc.yaml"

  # params$input_path <- "../06_create_figures/output/all_lean.yaml"
  # params$input_path <- "../06_create_figures/output/fb6.yaml"
  # params$input_path <- "../06_create_figures/output/external_validation.yaml"
}

input <- yaml::read_yaml(file = params$input_path)
config <- yaml::read_yaml(input$config_path)
```

```{r sup_tables load models, echo = FALSE}
input_07 <- normalizePath(
  paste0(file.path("..", "06_create_figures/output", config$out_tag), ".yaml")
)

render_env <- new.env()
rmarkdown::render("../07_display_figures/07_display_figures.Rmd",
  envir = render_env,
  params = list(input_path = input_07),
  quiet = TRUE
)
```

```{r sup_tables initiate figure and table number}
fig_no <- 1
tab_no <- 1
```

### Supplemental Table `r as.character(tab_no)`: Descriptive Statistics of Outcomes, Age, Measure Frequency, and Age Difference Between measures for complete, training, and testing datasets

```{r sup_tables describe prob class, results = 'asis'}
kable(render_env$save_figtabs$describe_census)
```


## Display Dataset/Model informaton 

Do this for complete dataset, training dataset, testing dataset.

- Number of observations per outcome
- Number of subjects per outcome
- Number of sujbects excluded for completeness

#### Complete and Training Data

```{r sup_tables dataset info complete and training, results = 'asis'}
kable(render_env$save_figtabs$sum_data)
```

#### Missing Data from Complete and Training Data

```{r sup_tables dataset info missing complete and training, results = 'asis'}
kable(render_env$save_figtabs$sum_missing)
```

#### Testing Data

```{r sup_tables dataset info testing data, results = 'asis'}
kable(render_env$save_figtabs$sum_test)
```

### Census information

- Study Demographics
- Median Life Expectancy

```{r sup_tables census information, results = 'asis'}
kable(render_env$save_figtabs$table_mle)
```

### LCMM information 

Include latent class model fit statistics

#### LCMM - Concise Table

```{r sup_tables lcmm concise, results = 'asis'}
kable(render_env$save_figtabs$table_lcmm)
```


#### LCMM - Average Posterior Classifcation Probabilites

```{r sup_tables lcmm model average posterior probs, results = 'asis'}
kable(render_env$save_figtabs$post_prob_table)
```

#### LCMM - Compare all possible models

```{r sup_tables compare all models, results = 'asis'}
kable(render_env$save_figtabs$all_models_table)
```


### LME information

```{r sup_tables model_display}
model_display <- 4
```

- Selected model - `r model_display`

### LME table

- Selected model - `r model_display`

```{r sup_tables lme table}
kable(render_env$save_figtabs$lme_table)
```

### Individual Cox Model

#### Table coxzph Individual Cox Model Fits

```{r sup_tables table coxzph individual cox model, results = 'asis'}
kable(render_env$save_figtabs$individual_coxzph_tables)
```

#### Table coxzph Individual Cox Model Fits - Combined Cox

```{r sup_tables table coxzph combined cox model, results = 'asis'}
kable(render_env$save_figtabs$combined_coxzph_tables)
```

