---
title: "SLAM Metabolic Class Supplemental Tables - `r config$out_tag` Mice"
subtitle: "External Validation: `r config$external_validate`"
author: "William Mueller"
data: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
params:
  input_path: "../07_display_figures/output/slam_c1-c10_age_all_bwfatgluc.yaml"
---

```{r sup_tables setup, warning=FALSE, message=FALSE, echo=FALSE}
library(helphlme)
library(tidyverse)
library(kableExtra)
library(survival)
library(survminer)
library(ggplotify)
library(sjPlot)
library(lme4)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(consoler)
library(flextable)
set.seed(365)

# setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = FALSE)
```

```{r sup_tables read input yaml}
# for the case when we are not running with RStudio
input <- yaml::read_yaml(file = params$input_path)
config <- yaml::read_yaml(input$config_path)
```

```{r sup_tables load models, echo = FALSE}
input_07 <- normalizePath(
  paste0(file.path("..", "06_create_figures/output", config$out_tag), ".yaml")
)

render_env <- new.env()
rmarkdown::render("../07_display_figures/07_display_figures.Rmd",
  envir = render_env,
  params = list(input_path = input_07),
  quiet = TRUE
)
```

```{r sup_tables initiate figure and table number}
fig_no <- 1
tab_no <- 1
```

### Supplemental Table `r as.character(tab_no)`: Descriptive Census Information

```{r sup_tables describe prob class, results = 'asis'}
flextable(round(render_env$save_figtabs$describe_census, digits = 3)) %>%
  autofit() %>%
  fontsize(size = 7.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_vanilla()


tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Descriptive Statistics of Outcomes, Age, Measure Frequency, and Age Difference Between measures for complete, training, and testing datasets

```{r sup_tables dataset info complete and training, results = 'asis'}
flextable(rbind(
  render_env$save_figtabs$sum_data,
  render_env$save_figtabs$sum_test
)) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 5.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Missing Data Descriptive Statistics of Outcomes, Age, Measure Frequency, and Age Difference Between measures for complete, training, and testing datasets

```{r sup_tables dataset info missing complete and training, results = 'asis'}
flextable(render_env$save_figtabs$sum_missing) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 5.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)


tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Median Life Expectancy Data

```{r sup_tables census information, results = 'asis'}
flextable(render_env$save_figtabs$table_mle) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 6.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: LCM Model Information

```{r sup_tables lcmm concise, results = 'asis'}
render_env$save_figtabs$lcmm_table %>%
  dplyr::rename(`Smallest Class %` = `Smallest.Class....`) %>%
  dplyr::mutate(`Smallest Class %` = round(`Smallest Class %`, digits = 3)) %>%
  flextable() %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 6.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: LCM Posterior Probabilities

```{r sup_tables lcmm model average posterior probs, results = 'asis'}
flextable(render_env$save_figtabs$post_prob_table) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 7.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Candidate LCMs

```{r sup_tables compare all models, results = 'asis'}
flextable(render_env$save_figtabs$all_models_table) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 3, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: LME Information

```{r sup_tables lme table}
render_env$save_figtabs$lme_table %>%
  mutate(sigma = round(sigma, digits = 3)) %>%
  mutate(loglik = round(loglik, digits = 0)) %>%
  mutate(AICtab = round(AICtab, digits = 0)) %>%
  flextable() %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 7.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Individual Cox Information

```{r sup_tables table coxzph individual cox model, results = 'asis'}
flextable(render_env$save_figtabs$individual_coxzph_tables) %>%
  autofit() %>%
  theme_vanilla() %>%
  fontsize(size = 7.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0)

tab_no <- 1 + tab_no
```

\newpage

### Supplemental Table `r as.character(tab_no)`: Combined Cox Information

```{r sup_tables table coxzph combined cox model, results = 'asis'}
render_env$save_figtabs$combined_coxzph_tables %>%
  mutate(df = round(df, digits = 2)) %>%
  select(-outcome) %>%
  flextable() %>%
  autofit() %>%
  fontsize(size = 7.5, part = "all") %>%
  set_table_properties(layout = "autofit", width = 0) %>%
  theme_vanilla()
tab_no <- 1 + tab_no
```

